---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

 This code is for comparison of cluster specificity across different integration methods and gene homology
 

Upload all 4 methods with 3 homology options = 12 datasets

```{r setup, echo=FALSE, include=FALSE}
suppressPackageStartupMessages({
  library(cluster)
  library(Seurat)
  library(SeuratObject)
  library(cowplot)
  library(stats)
  library(data.table)
  library(tidyverse)
  library(ggsci)
  library(scCustomize)
  library(patchwork)
  library(sctransform)
  library(MatrixGenerics)
  library(openxlsx)
  library(RColorBrewer)
  library(MetBrewer)
  library(pheatmap)
  library(corrplot)
  library(DESeq2)
  library(viridis)
  library(scales)
  library(ggpubr)
  library(SCpubr)
  library(ggplotify)
  library(bluster)
  library(stringi)
  library(gridExtra)
  library(limma)
  library(muscat)
  library(scran)

  options(scipen = 999)
  options(repr.matrix.max.rows = 20)
  options(repr.matrix.max.cols = 50)
})
set.seed(8899)
my_pal1 <- MetBrewer::met.brewer('Egypt', n = 4, type = 'discrete')
my_pal2 <- MetBrewer::met.brewer('Signac', n = 14, type = 'discrete')
options(width = 200)


setwd(
  "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity"
)
knitr::opts_knit$set(root.dir = "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
```

# RPCA

```{r include=FALSE}
# List of .rds files
files <- list.files(path = '~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_less_mice_5/results/results/seuratRPCA/cross_species/integrated_h5ad/', pattern = '\\.rds', full.names = TRUE)

# Directory to work in
work_dir <- "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity"

# Final species genes
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98), final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
```


```{r include=FALSE}
# Loop through each file
for (file in files) {
  my_data <- readRDS(file)

  my_counts <- GetAssayData(object = my_data, assay = 'RNA', slot = 'data')
  rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["RNA_genes"]] <- CreateAssayObject(data = my_counts)
  DefaultAssay(my_data) <- "RNA_genes"
  
  # Save the modified Seurat object
  saveRDS(my_data, file = paste0(work_dir, '/', basename(file)))
  
  # Clean up memory
  rm(my_counts, my_data, df)
}

print("All files processed.")
```

# CCA

```{r include=FALSE}
# List of .rds files
files <- list.files(path = '~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_less_mice_5/results/results/seuratCCA/cross_species/integrated_h5ad/', pattern = '\\.rds', full.names = TRUE)

# Directory to work in
work_dir <- "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity"

# Final species genes
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98), final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)

# Loop through each file
for (file in files) {
  my_data <- readRDS(file)

  my_counts <- GetAssayData(object = my_data, assay = 'RNA', slot = 'data')
  rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["RNA_genes"]] <- CreateAssayObject(data = my_counts)
  DefaultAssay(my_data) <- "RNA_genes"
  
  # Save the modified Seurat object
  saveRDS(my_data, file = paste0(work_dir, '/', basename(file)))
  
  # Clean up memory
  rm(my_counts, my_data, df)
}

print("All files processed.")
```

# scVI and scANVI: create a list

```{r}
work_dir <- '/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_less_mice_5/results/results/seurat_for_py'
files <- list.files(path = '/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_less_mice_5/results/results/seurat_for_py', pattern = '\\.rds', full.names = TRUE)

py_based <- list()
for(i in  seq_along(files) ){
  name_f <-str_split(str_split(files[i], pattern = '/')[[1]][15], pattern = '_')[[1]]
  name_f <- name_f[-grep(pattern = 'matrix', x = name_f)]
  name_f <- unique(name_f)
  x <- match(1, str_detect(string = name_f, pattern = 'nf')) +1
  y <- match(1, str_detect(string = name_f, pattern = 'Seurat'))-1
  name_f <- paste0(name_f[  x: y ], collapse = '_')
  print(name_f)
  py_based[[name_f]] <- readRDS(files[i])
}

```

# CCA and RPCA: create a list

```{r}
work_dir <- "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity"
files <- list.files(path = "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity", pattern = '\\.rds', full.names = TRUE)

r_based <- list()
for(i in  seq_along(files) ){
  name_f <- str_split(str_split(files[i], pattern = '/')[[1]][13], pattern = '_')[[1]]
  # name_f <- name_f[-grep(pattern = 'matrix', x = name_f)]
  # name_f <- unique(name_f)
  x <- match(1, str_detect(string = name_f, pattern = 'nf')) +1
  y <- match(1, str_detect(string = name_f, pattern = 'integrated'))-1
  name_f <- paste0(name_f[  x: y ], collapse = '_')
  print(name_f)
  r_based[[name_f]] <- readRDS(files[i])
}
```

# Choose the right clusterisation for R-based mathods
__Silhouette does not make sense__

```{r include=FALSE}
#from https://github.com/MillerLab-CPHG/Human_athero_scRNA_meta/blob/main/integration_benchmarking/additional_integration_benchmark_metrics.R
# Write a function to calculate clusters silhouette scores from a seurat object
# Arg seurat obj: A seurat obj with reduced dims embedding like pca
# Arg distance_matrix: distance matrix between cells in reduced dims embedding 
# Value: A data.frame with silhouette scores for a given clustering resolution  
calc_sil_scores = function(seurat_obj, distance_matrix) {
  # Calculate distance matrix from seurat obj reduced dims embedding 
  #dist_matrix = dist(x = Seurat::Embeddings(object = seurat_obj[[reduced_dims]])[, 1:30])
  clusters = seurat_obj$seurat_clusters
  sil = silhouette(x = as.numeric(x = as.factor(x = clusters)), 
                   dist = distance_matrix)
  # Add silhouette scores back into seurat obj metadata
  #seurat_obj$silhouette_score = sil[, 3]
  sil_df = data.frame(cluster=sil[, 1],
                      neighbor=sil[, 2],
                      sil_width=sil[, 3])
  return(sil_df)
}


# Directory for saving output
output_dir <- "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity"

# Define a range of resolutions to test
res <- seq(0.1, 1, by = 0.1)

# Loop through each Seurat object
for (name in names(r_based)) {
  my_data <- r_based[[name]]

  # rPCA
  rpca_pca_embeddings <- Embeddings(my_data, "pca")[, 1:30]
  DefaultAssay(my_data) <- 'integrated'

  # Calculate distance matrix
  dist_matrix <- dist(x = rpca_pca_embeddings)
  saveRDS(dist_matrix, paste0(output_dir, "/", name, "_distance_matrix.rds"))
  
  rpca_scores_list <- list()

  for (i in seq_along(res)) {
    # Cluster data
    my_data <- FindClusters(my_data, resolution = res[i])
  
    # Calculate silhouette scores
    rpca_scores_list[[i]] <- calc_sil_scores(my_data, dist_matrix)
    names(rpca_scores_list)[[i]] = paste("res", 
                                       as.character(res[i]), 
                                       sep = "_")
  }

  # Create a single df with all the sil scores for the tested resolutions
  rpca_resolutions_sil_df <- rbindlist(rpca_scores_list, idcol = "method")
  rpca_resolutions_sil_df$method_name <- name
  saveRDS(rpca_resolutions_sil_df, paste0(output_dir, "/", name, "_sil_scores_per_clustering_resolution.rds"))
}
```

# For py based methods

```{r include=FALSE}
# Directory for saving output
output_dir <- "/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity"

# Define a range of resolutions to test
res <- seq(0.1, 1, by = 0.1)

# Loop through each Seurat object
for (name in names(py_based)) {
  my_data <- py_based[[name]]
  
if(str_detect(string = name, pattern = 'scANVI')){
  rpca_pca_embeddings <- Embeddings(my_data, "X_scANVI")[, 1:10]
} else rpca_pca_embeddings <- Embeddings(my_data, "X_scVI")[, 1:40] 
  
  # DefaultAssay(my_data) <- 'integrated'

  # Calculate distance matrix
  dist_matrix <- dist(x = rpca_pca_embeddings)
  saveRDS(dist_matrix, paste0(output_dir, "/", name, "_distance_matrix.rds"))
  
  rpca_scores_list <- list()

  for (i in seq_along(res)) {
    # Cluster data
    if(str_detect(string = name, pattern = 'scANVI')){
      my_data <- RunUMAP(my_data, dims = 1:10, reduction = "X_scANVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:10, reduction = "X_scANVI")
    } else {
      my_data <- RunUMAP(my_data, dims = 1:40, reduction = "X_scVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:40, reduction = "X_scVI")
    }
      
    
    my_data <- FindClusters(my_data, resolution = res[i])
  
    # Calculate silhouette scores
    rpca_scores_list[[i]] <- calc_sil_scores(my_data, dist_matrix)
    names(rpca_scores_list)[[i]] = paste("res", 
                                       as.character(res[i]), 
                                       sep = "_")
  }

  # Create a single df with all the sil scores for the tested resolutions
  rpca_resolutions_sil_df <- rbindlist(rpca_scores_list, idcol = "method")
  rpca_resolutions_sil_df$method_name <- name
  saveRDS(rpca_resolutions_sil_df, paste0(output_dir, "/", name, "_sil_scores_per_clustering_resolution.rds"))
}
```

```{r}
DimPlot(my_data, reduction = "umap",  group.by = 'smc_phenotypes', split.by = 'species')
```

```{r}
res_sil_list <- list()
files <- list.files(path = '/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity', pattern = '_sil_scores_per_clustering_resolution.rds')
for(i in seq_along(files)){
  res_sil_list[[i]] <- readRDS(files[i])
}
res_sil_scores_df <- bind_rows(res_sil_list)
```

```{r}
res_sil_scores_df$method_name = factor(res_sil_scores_df$method_name)
colnames(res_sil_scores_df)[1] <- '.id'
res_sil_scores_df <- res_sil_scores_df[res_sil_scores_df$.id %in% c('res_0.3', 'res_0.4', 'res_0.5', 'res_0.6', 'res_0.7',  'res_0.9',   'res_1')]
res_sil_scores_plot = res_sil_scores_df %>%
  group_by(.id, method_name) %>%
  summarize(mean_sil = mean(sil_width)) %>%
  ggplot(aes(x=.id, y=mean_sil, fill=method_name)) + geom_col(position="dodge", width = 0.7) +
  xlab("Integration Methods / Homology") +
  ylab("Mean Silhouette score") +
  # ggtitle("rPCA mean sil scores per clustering resolution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") + theme_bw() 

ggsave(file="Silh_resolutions.pdf",
       plot = res_sil_scores_plot, width = 16, height = 4.5)
```

```{r include=FALSE}
saveRDS(py_based, file = 'py_based_integration_methods.rds')
saveRDS(r_based, file = 'r_based_integration_methods.rds')
```

```{r}
py_based <- readRDS('py_based_integration_methods.rds')
r_based <- readRDS('r_based_integration_methods.rds')
```

# Do Clutserisation, add assays with gene names istead od Ensembl IDs

```{r include=FALSE}
# Clusterize
r_based_next <- list()
# Loop through each Seurat object
for (name in names(r_based)) {
  my_data <- r_based[[name]]
  DefaultAssay(my_data) <- 'integrated'
  my_data <- FindNeighbors(my_data, reduction = "pca", dims = 1:30)
  my_data <- FindClusters(my_data, resolution = c(0.3, 0.4, 0.5, 0.6))
  
  my_counts <- GetAssayData(object = my_data, assay = 'integrated', slot = 'data')
  rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) <- gsub(pattern = '-1-1', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["integrated_genes"]] <- CreateAssayObject(data = my_counts)
  
  r_based_next[[name]] <- my_data
}
```

```{r include=FALSE}
py_based_next <- list()
for(name in names(py_based)){
  my_data <- py_based[[name]]
  
  my_counts <- GetAssayData(object = my_data, assay = 'RNA', slot = 'data')
  # rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["RNA_genes"]] <- CreateAssayObject(data = my_counts)
  
  # Cluster data
    if(str_detect(string = name, pattern = 'scANVI')){
      my_data <- RunUMAP(my_data, dims = 1:10, reduction = "X_scANVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:10, reduction = "X_scANVI")
    } else {
      my_data <- RunUMAP(my_data, dims = 1:40, reduction = "X_scVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:40, reduction = "X_scVI")
    }
  my_data <- FindClusters(my_data, resolution = c(0.3, 0.4, 0.5, 0.6))
  py_based_next[[name]] <- my_data
  
    # Clean up memory
  rm(my_counts, df)
}
```


# Choose right clusterization - so all human clusters are distinguished in the integrated datasets. Clusters should be stable through bootsraping

```{r}
# finding the right clusterization for r_based methods: iterating through resolutions, by bootstrap identify stability of clusters
# then keep only those resolutions that allow to identify more than 7 clusters (6 from human and 1 from others)
my_k <- seq(0.1, 1, 0.1)
df_stable_clusters <- data.frame(k = numeric(),
                                 dataset = character(),
                                 mean_ratio = numeric(),
                                 n_clusters = numeric())
n=1
for(name in names(r_based_next)) {
  for (j in my_k) {
    set.seed(8899)
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = j)$membership
    }
    
    my_data <- r_based_next[[name]]
    DefaultAssay(my_data) <- 'integrated'
    
    pcs <- reducedDim(as.SingleCellExperiment(my_data), "PCA")
    originals <- myClusterFUN(pcs)
    df_stable_clusters[n, 'k'] <- j
    df_stable_clusters[n, 'dataset'] <- name
    
    df_stable_clusters[n, 'mean_ratio'] <-
      mean(diag(
        bootstrapStability(pcs, FUN = myClusterFUN, clusters = originals, iterations = 20)
      ))
    df_stable_clusters[n, 'n_clusters'] <- length(unique(originals))
    n <- n + 1
  }
}
# py_based methods
df_stable_clusters_2 <- data.frame(k = numeric(),
                                 dataset = character(),
                                 mean_ratio = numeric(),
                                 n_clusters = numeric())
n=1
for(name in names(py_based_next)) {
  for (j in my_k) {
    set.seed(8899)
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = j)$membership
    }
    
    my_data <- py_based_next[[name]]
    DefaultAssay(my_data) <- 'RNA'
    if(str_detect(string = name, pattern =  'scANVI')){
      pcs <- reducedDim(as.SingleCellExperiment(my_data), "X_SCANVI")
    } else  pcs <- reducedDim(as.SingleCellExperiment(my_data), "X_SCVI")
    
    originals <- myClusterFUN(pcs)
    df_stable_clusters_2[n, 'k'] <- j
    df_stable_clusters_2[n, 'dataset'] <- name
    
    df_stable_clusters_2[n, 'mean_ratio'] <-
      mean(diag(
        bootstrapStability(pcs, FUN = myClusterFUN, clusters = originals, iterations = 20)
      ))
    df_stable_clusters_2[n, 'n_clusters'] <- length(unique(originals))
    n <- n + 1
  }
}
df_stable_clusters_fin <- rbind(df_stable_clusters, df_stable_clusters_2)
saveRDS(df_stable_clusters_fin, file = 'cluster_stability_bootstrap_res.rds')
```


```{r}
df_stable_clusters_fin <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/cluster_stability_bootstrap_res.rds")
df_stable_clusters_fin <- df_stable_clusters_fin[df_stable_clusters_fin$n_clusters > 7,]

p <- ggplot(df_stable_clusters_fin, aes(x = factor(k), y = mean_ratio, group = dataset, color = dataset)) +
  geom_line() +
  geom_point() +  # Optional: Adds points to the line
  labs(x = "Resolution", y = "Mean Ratio", color = "Dataset") +
  theme_minimal()
print(p)
ggsave(filename = 'Cluster_stability_over_7.pdf', plot = p, width = 6,height = 4, units = 'cm', scale = 3)
```

# Clustering is done not by Seurat so that's why cluster numbers are not aligned to cluster size

```{r}
# Filter rows to keep only the maximum mean_ratio for each dataset
df_max <- df_stable_clusters_fin %>%
  group_by(dataset) %>%
  filter(mean_ratio == max(mean_ratio))
df_max <- as.data.frame(df_max)

# Best resolution that gives more than 7 clusters for r_based

for(i in seq_along(df_max$dataset[1:6])){
  name <- df_max[i, 'dataset']
  
   set.seed(8899)
   
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = df_max[i, 'k'])$membership
    }
    
    
    DefaultAssay(r_based_next[[name]]) <- 'integrated'
    
    pcs <- reducedDim(as.SingleCellExperiment(r_based_next[[name]]), "PCA")
    originals <- myClusterFUN(pcs)
    
    r_based_next[[name]]$stab_clusters <- originals
 
}

# Best resolution that gives more than 7 clusters for py_based

for(i in seq_along(df_max$dataset)[7:12]){
  name <- df_max[i, 'dataset']
  
   set.seed(8899)
   
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = df_max[i, 'k'])$membership
    }
    
    
    DefaultAssay(py_based_next[[name]]) <- 'RNA'
    
     if(str_detect(string = name, pattern =  'scANVI')){
      pcs <- reducedDim(as.SingleCellExperiment(py_based_next[[name]]), "X_SCANVI")
    } else  pcs <- reducedDim(as.SingleCellExperiment(py_based_next[[name]]), "X_SCVI")
    
    originals <- myClusterFUN(pcs)
    
    py_based_next[[name]]$stab_clusters <- originals
 
}

# Aggregate results for r_based and py_based methods and save

all_datasets <- c(r_based_next, py_based_next)
saveRDS(all_datasets, file = 'r_py_all_datasets_processed.rds')
```

# Name clusters (using majority for a cluster)
(stability of clusters)

```{r}
# make new metadata variable 'smc_phenotypes_short'
for (name in names(all_datasets)) {
  Idents(all_datasets[[name]]) <- 'smc_phenotypes'
  
  new_names <-
    c("Trans",
      "SMC1", # called 'Contratile' in the paper's plots
      "Term",
      "PC",
      "FB",
      'SMC2', # Medial
      'SMC3', # IFN_SMC
      'DLX5',
      'CRTAC1',
      'SMC6') #  Contractile_Rgs5
  names(new_names) <- levels(Idents(all_datasets[[name]]))
  
  all_datasets[[name]] <-
    Seurat::RenameIdents(object = all_datasets[[name]], new_names)
  new_names <- Idents(all_datasets[[name]])
  all_datasets[[name]]$smc_phenotypes_short <-
    new_names
  
  rm(new_names)
}

for (name in names(all_datasets)) {
  my_table <-
    table(all_datasets[[name]]$smc_phenotypes_short,
          all_datasets[[name]]$stab_clusters)
  
  # Function to get the row name with the max value for each column
  get_max_rowname <- function(column) {
    max_index <- which.max(column)
    rownames(my_table)[max_index]
  }
  
  # Apply this function to each column of the table
  max_rownames <- apply(my_table, 2, get_max_rowname)
  # Get the cluster numbers as they appear in the table
  cluster_numbers <- colnames(my_table)
  
  # Create a named vector where names are cluster numbers and values are the phenotypes
  named_max_rownames <- setNames(max_rownames, cluster_numbers)
  
  # Create the 'specific_clusters' metadata variable
  all_datasets[[name]]$specific_clusters <-
    sapply(all_datasets[[name]]$stab_clusters, function(cluster) {
      paste(cluster, paste('(', named_max_rownames[as.character(cluster)], ')', sep = ''), sep = " ")
    })
  
}
```

# Plots

```{r}
text_sizes <- theme(axis.text.x=element_text(size=9,colour="black"),
                      axis.text.y=element_text(size=9,colour="black"),
                      axis.title.y=element_text(size=9,colour="black", margin = margin(t = 2, l = 1, r = 1, b =2, unit = "pt")),
                      axis.title.x=element_text(size=9,colour="black", margin = margin(t = 1, l = 2, r = 2, b = 1, unit = "pt")),
                      legend.text = element_text(size=9,colour="black"),
                      legend.title = element_text(size=9,colour="black", margin = margin(t = 5, l = 0, r = 0, b = 5, unit = "pt")),
                      legend.key = element_rect(colour="transparent", fill = "transparent"),
                      strip.text.x = element_text(size=6,color = 'black',face="bold", angle=0),
                      strip.text.y = element_text(size=6,color = 'black', face="bold", angle=0, vjust=0.5, hjust=0),
                      axis.ticks= element_line(color = 'black', size=0.2),
                      axis.line = element_line(colour = "gray50", size = 0.2, linetype = "solid"),
                      plot.margin=unit(c(0,0,0,0),"pt"),
                      plot.title=element_text(size=7, face='plain', colour="black"))


common_minimal <- text_sizes + theme(
        plot.background = element_rect(fill = NA,colour = NA),
        strip.background = element_rect(fill = NA,colour = NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) + theme(
    legend.spacing = unit(0.15, 'cm'), 
    legend.key.size = unit(0.2, "cm"))


# commonly used, x axis text 45 degree
common_0x <- common_minimal + theme(axis.text.x = element_text(angle=0))

# pal_species <- pal_npg("nrc", alpha = 0.7)(9)[c(1,4,3)]
#   # MetBrewer::met.brewer(name = 'Thomas')[c(1,5,7)]
# names(pal_species) <- c('sscrofa', 'hsapiens', 'mmusculus')
# saveRDS(pal_species, file = 'pal_species.rds')
# 
# pal_celltypes <- c(pal_nejm(palette = 'default')(8)[-6], pal_jama(palette = 'default')(7) )
#   # sample(c(MetBrewer::met.brewer(name = 'Thomas'), MetBrewer::met.brewer(name = 'Cross')))
# saveRDS(pal_celltypes, file = 'pal_celltypes.rds')
# 
# show_col(pal_jama(palette = 'default')(9))
```

```{r}
pal_celltypes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_celltypes.rds")
pal_species <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_species.rds")
names(pal_species) <- c('pig', 'human', 'mouse')
```


__Stacked barplots for celltype composition in species for the stable named clusters__

__First for the purpose of illustration, the contribution of species to the final dataset should be balanced. I will take 1500 cell from each species and exclude GFP and TDT mouse sorted cells__

```{r}
for (name in names(all_datasets)) {
  all_datasets[[name]]$species_2 <- all_datasets[[name]]$species
  all_datasets[[name]]$species[all_datasets[[name]]$species == 'hsapiens'] <-
    'human'
  all_datasets[[name]]$species[all_datasets[[name]]$species == 'mmusculus'] <-
    'mouse'
  all_datasets[[name]]$species[all_datasets[[name]]$species == 'sscrofa'] <-
    'pig'
}
```


```{r}

# creating balanced datasets
set.seed(8899)

my_data <- all_datasets$many_higher_expr_seuratRPCA
mice <-
    my_data[, my_data$orig.ident %in% c('Klf4_WT_sorted_SMCs', 'SMC_WT_eYFP_sorted')]

mouse_cells <- colnames( mice[, sample(colnames(mice), size = 1500, replace = F)])

human_cells <- colnames( my_data[, sample(colnames(my_data[, my_data$species %in% c('human')]), size =
                       1500, replace = F)] )

pig_cells <- colnames(my_data[, sample(colnames(my_data[, my_data$species %in% c('pig')]), size =
                       1500, replace = F)])


# my_seu_balanced <- list()
# for (name in names(all_datasets)) {
#   my_data <- all_datasets[[name]]
#   mice <-
#     my_data[, my_data$orig.ident %in% c('Klf4_WT_sorted_SMCs', 'SMC_WT_eYFP_sorted')]
#   mice <- mice[, sample(colnames(mice), size = 1500, replace = F)]
#   human <-
#     my_data[, sample(colnames(my_data[, my_data$species %in% c('human')]), size =
#                        1500, replace = F)]
#   pig <-
#     my_data[, sample(colnames(my_data[, my_data$species %in% c('pig')]), size =
#                        1500, replace = F)]
#   
#   balanced <-
#     Merge_Seurat_List(list_seurat = list(mice, human, pig),
#                       merge.data = T)
#   my_seu_balanced[[name]] <- balanced
#   rm(mice, human, pig, balanced)
# }

my_seu_balanced <- list()
for (name in names(all_datasets)) {
  my_data <- all_datasets[[name]]
  # mice <-
  #   my_data[, my_data$orig.ident %in% c('Klf4_WT_sorted_SMCs', 'SMC_WT_eYFP_sorted')]
  mice <- my_data[, mouse_cells]
  human <-
    my_data[, human_cells]
  pig <-
    my_data[, pig_cells]
  
  balanced <- merge(mice, y = c(human, pig))
  
    # Merge_Seurat_List(list_seurat = list(mice, human, pig),
    #                   merge.data = T)
  my_seu_balanced[[name]] <- balanced
  rm(mice, human, pig, balanced)
}

```

```{r}
new_names <- c("HE_CCA","HE_RPCA","HHC_CCA",  "HHC_RPCA","O2O_CCA", "O2O_RPCA", "HE_scANVI","HE_scVI",   "HHC_scANVI" ,    "HHC_scVI", "O2O_scANVI"  ,"O2O_scVI" )
names(new_names) <- names(all_datasets)

new_names <- new_names[c(2,4,6,
                         1,3,5,
                         7,9,11,
                         8,10,12)]
```


```{r}
cluster_prop <- list()
for (name in names(new_names)) {
  balanced <- my_seu_balanced[[name]]
  # res <- colnames(balanced@meta.data)[grep(pattern = '0.6', x = colnames(balanced@meta.data))]
  # contribution
  df <-
    as.data.frame.array(table(balanced@meta.data[, 'specific_clusters'], balanced$species))
  df$sum <- rowSums(df)
  df$pct_sum <- round(df$sum / sum(df$sum) * 100, 1)
  
  #  creating a table with proportions
  # Assuming df is your original dataframe
  df_prop <- df
  
  # Calculate proportions for each column except 'sum'
  df_prop$human <- df$human / df$sum
  df_prop$mouse <- df$mouse / df$sum
  df_prop$pig <- df$pig / df$sum
  
  # The 'sum' column remains unchanged or can be removed if not needed
  df_prop$sum <- NULL  # Uncomment to remove the 'sum' column
  
  # View the new dataframe with proportions
  df_prop[, 1:3] <- df_prop[, 1:3] * 100
  df_prop$cluster <- rownames(df_prop)
  # df_prop$cluster <- as.numeric(df_prop$cluster)
  df_prop <- df_prop[order(x = df_prop$pct_sum, decreasing = F), ]
  
  
  # Transform data to long format
  df_long <- gather(df_prop, species, percent,-cluster,-pct_sum)
  
  
  
  # Calculate cumulative proportions for text placement correctly within each cluster
  df_long <- df_long %>%
    group_by(cluster) %>%
    mutate(cumulative = 100 - (cumsum(percent) - percent / 2))
  
  df_long$cluster <-
    factor(df_long$cluster, levels = unique(df_long$cluster))
  
  # p1 <-
  #   ggplot(df_long, aes(x = cluster, y = percent, fill = species)) +
  #   geom_bar(
  #     # aes(width = rescale(pct_sum, c(0.1, 1))),
  #     stat = "identity", position = "stack") +
  #   geom_text(aes(label = ifelse(percent >= 5, round(percent, 1), ""),
  #                 y = cumulative),
  #             size = 3) +
  #   geom_text(aes(label = pct_sum), y = 108) +
  #   scale_y_continuous(limits = c(0, 110)) +
  #   scale_fill_manual(values = pal_species) +
  #   theme_bw() +
  #   ggtitle(new_names[name]) +
  #   coord_flip()
  #   # rotate_x_text(angle = 45)
  
  p1 <-
    ggplot(df_long, aes(x = cluster, y = percent, fill = species)) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(aes(label = ifelse(percent >= 5, round(percent, 1), ""),
                  y = cumulative),
              size = 3,
              family = "Helvetica") +
    geom_text(aes(label = pct_sum), y = 108, family = "Helvetica") +
    scale_y_continuous(limits = c(0, 110)) +
    scale_fill_manual(values = pal_species) +
    theme_bw() +
    ggtitle(new_names[name]) +
    coord_flip() +
    theme(
      text = element_text(family = "Helvetica"),
      # Set global text family to Arial
      plot.title = element_text(family = "Helvetica"),
      # Ensure title is also Arial
      axis.text = element_text(family = "Helvetica"),
      # Axis texts
      axis.title = element_text(family = "Helvetica")
    )  # Axis titles
  
  
  cluster_prop[[name]] <- p1
}
saveRDS(cluster_prop, file = 'cluster_proportions_list_plots.rds')
n <- length(cluster_prop)
nCol <- floor(sqrt(n))

p1 <- do.call("grid.arrange", c(cluster_prop, ncol = nCol))

# p1 <- plot_grid(plotlist = cluster_prop, nrow = 3, ncol = 4, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Helvetica', label_x = -0.1)
# 
# ggsave(filename = 'proportions_stable_cluster.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 2.5)

```


```{r}
pdf(file = 'proportions_stable_cluster.pdf', width = 16, height = 14, family = 'Helvetica')
plot(p1)
dev.off()
```


__UMAPs with the stable clusters + species__


```{r}
# for purpose of visualisation, subset human part to 2500 cells randomly

myplots <- list()


for(name in names(new_names)) {
  my_vec <- sample(names(all_datasets[[name]]$species[all_datasets[[name]]$species == 'human']), size = 2500, replace = F) 
  my_species = DimPlot(
  object = all_datasets[[name]][,all_datasets[[name]]$species %in% c('pig', 'mouse')| colnames(all_datasets[[name]]) %in% my_vec],
    reduction = 'umap',
    group.by = c('species'),
    pt.size = 0.5,
    shuffle = TRUE
  ) + theme(legend.position = 'none') + text_sizes + theme(plot.title = element_blank()) + 
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
    scale_color_manual(values = pal_species) + theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  my_cell_type = DimPlot_scCustom(seurat_object = all_datasets[[name]][,all_datasets[[name]]$species %in% c('pig', 'mouse')| colnames(all_datasets[[name]]) %in% my_vec],
    reduction = 'umap',
    group.by = 'specific_clusters',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes, repel = T, label.size = 2.3
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  
fig = ggdraw() +
  draw_plot(my_species, x = 0.0, y = .54, width = .99, height = .45) +
  draw_plot(my_cell_type, x = 0.0, y = 0.00005, width = .99, height = .55)  +
  draw_plot_label(label = new_names[name], size = 12,
                  x = c(0), y = c(1.005), family = 'Helvetica')

  
  # fig = plot_grid(my_species,
  #                 my_cell_type,
  #                 ncol = 1, rel_widths = c(0.5, 1),
  #                 rel_heights  = c(0.5, 1))
  
  myplots[[name]] <- fig
  
}

saveRDS(myplots, 'succesful_integration_specific_clusters_species.rds')

# library(gridExtra)
# 
# do.call("grid.arrange", c(myplots[c(1,2)], ncol=2))
```


```{r}

# p <- print(plot_grid(myplots[1][[1]], myplots[2][[1]], myplots[3][[1]], myplots[4][[1]], myplots[5][[1]], myplots[6][[1]], 
#                 myplots[7][[1]], myplots[8][[1]], myplots[9][[1]], myplots[10][[1]], myplots[11][[1]], myplots[12][[1]]
#                 , ncol = 6, axis = 'lb'))

p <- plot_grid(plotlist = myplots, nrow = 2, ncol = 6, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Helvetica', label_x = -0.1)

ggsave(filename = 'UMAP_succesful_integration_specific_clusters_species.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 2)

# save_plot(filename = 'test.pdf',  plot = p, ncol = 6, nrow = 4, )

```

# To find consistent species-specific and shared clusters, we assumed that a cluster is species-specific if: either one species contributes more than 75% of its composition, or one species contributes less than 10% of a cluster composition. In contrast, a clusters was considered as shared if all species contribute to it from 10.1 to 74.9%. 

Here I also introduce a limitation for a size of clusters - it is worth considering only big enough clusters, let's say that their contribution should be more than 2 percent out of all integrated dataset cells.

Therefore, here we work with balanced datasets, where each species contribute 1500 cells. The total number of cells is 4500, and the threshold is 49 cells (3.2% out of 1500 cells)

```{r}
# Adjusted function to categorize each cluster
categorize_cluster <- function(human, mouse, pig) {
  contributions <- c(human = human, mouse = mouse, pig = pig)
  high_contributions <- contributions > 80
  low_contributions <- contributions < 10
  
  if (sum(high_contributions) == 1) {
    # Single species contributes more than 80%
    return(paste(names(contributions)[high_contributions], "specific", sep = "-"))
  } else if (all(contributions >= 10.1 & contributions <= 79.9)) {
    # All species contribute between 11% and 79%
    return("shared")
  } else if (sum(low_contributions) == 1) {
    # Exactly one species contributes less than 10%, making the cluster specific for the other two species
    # Filter out the species with less than 10% contribution
    specific_species <- names(contributions)[!low_contributions]
    # Create a label indicating the cluster is specific to the remaining species
    specific_label <- paste(specific_species, collapse = "-")
    return(paste0(specific_label, "-specific"))
  } else {
    return("mixed")
  }
}

my_spec_clusters <- data.frame(
  Method = character(),
  Mouse_specific = character(),
  Human_specific = character(),
  Pig_specific = character(),
  Mouse_pig_specific = character(),
  Mouse_human_specific = character(),
  Human_pig_specific = character(),
  Shared = character(),
  Too_small_cluster = character()
)

for (name in names(new_names)) {
  balanced <- my_seu_balanced[[name]]
  # res <- colnames(balanced@meta.data)[grep(pattern = '0.6', x = colnames(balanced@meta.data))]
  # contribution
  df <-
    as.data.frame.array(table(balanced@meta.data[, 'specific_clusters'], balanced$species))
  df$sum <- rowSums(df)
  df$pct_sum <- round(df$sum / sum(df$sum) * 100, 1)
  
  #  creating a table with proportions
  # Assuming df is your original dataframe
  df_prop <- df
  
  # Calculate proportions for each column except 'sum'
  df_prop$human <- df$human / df$sum
  df_prop$mouse <- df$mouse / df$sum
  df_prop$pig <- df$pig / df$sum
  
  # The 'sum' column remains unchanged or can be removed if not needed
  # df_prop$sum <- NULL  # Uncomment to remove the 'sum' column
  
  # View the new dataframe with proportions
  df_prop[, 1:3] <- df_prop[, 1:3] * 100
  df_prop$cluster <- rownames(df_prop)
  # df_prop$cluster <- as.numeric(df_prop$cluster)
  df_prop <- df_prop[order(x = df_prop$pct_sum, decreasing = T), ]
  df_prop$Big <- df_prop$sum > 49
  
  # # filter out small clusters, that are less than 0.6%
  #  df_prop <- df_prop[df_prop$sum > 49,]
  
  df_prop$category <- apply(df_prop[, c("human", "mouse", "pig")], 1, function(x) categorize_cluster(x['human'], x['mouse'], x['pig']))
  
  df_prop$category <- gsub(pattern = 'human.human', replacement = 'human', x = df_prop$category)
  df_prop$category <- gsub(pattern = 'pig.pig', replacement = 'pig', x = df_prop$category)
  df_prop$category <- gsub(pattern = 'mouse.mouse', replacement = 'mouse', x = df_prop$category)
  
  my_spec_clusters[name, 'Method'] <-  new_names[name]
  
  my_spec_clusters[name, 'Mouse_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'mouse-specific' & df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  my_spec_clusters[name, 'Human_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'human-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  my_spec_clusters[name, 'Pig_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'pig-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ' ')
  
  my_spec_clusters[name, 'Mouse_pig_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'mouse-pig-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  my_spec_clusters[name, 'Mouse_human_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'human-mouse-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  my_spec_clusters[name, 'Human_pig_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'human-pig-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  
  my_spec_clusters[name, 'Shared'] <- paste(c(unlist(df_prop[df_prop$category == 'shared'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  
  my_spec_clusters[name, 'Too_small_cluster'] <- paste(c(unlist(df_prop[df_prop$Big == FALSE, 'cluster'])), collapse = ', ')

}

saveRDS(my_spec_clusters, file = 'specific_clusters_df.rds')

print(my_spec_clusters)
```

```{r}
set.seed(8899)
# pal_cat <- sample(pal_jco(palette = 'default')(8))
# saveRDS(pal_cat, file = 'pal_cat.rds')
pal_cat <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_cat.rds")
pal_celltypes_all <- c("#0072B5FF",  "#79AF97FF",  "#374E55FF","#80796BFF", "#FFDC91FF","#EE4C97FF", "#7876B1FF","#20854EFF", "#BC3C29FF",  "#E18727FF")
```


```{r}
names(pal_cat) <- c('shared','too_small',  'human-pig-specific','human-specific','human-mouse-specific' , 'mouse-specific',   'pig-specific', 'mouse-pig-specific' )
myplots <- list()
my_prop <- list()
for (name in names(new_names)) {
  my_data <- my_seu_balanced[[name]]
  # res <- colnames(balanced@meta.data)[grep(pattern = '0.6', x = colnames(balanced@meta.data))]
  # contribution
  df <-
    as.data.frame.array(table(my_data@meta.data[, 'specific_clusters'], my_data$species))
  df$sum <- rowSums(df)
  df$pct_sum <- round(df$sum / sum(df$sum) * 100, 1)
  
  #  creating a table with proportions
  # Assuming df is your original dataframe
  df_prop <- df
  
  # Calculate proportions for each column except 'sum'
  df_prop$human <- df$human / df$sum
  df_prop$mouse <- df$mouse / df$sum
  df_prop$pig <- df$pig / df$sum
  
  # The 'sum' column remains unchanged or can be removed if not needed
  # df_prop$sum <- NULL  # Uncomment to remove the 'sum' column
  
  # View the new dataframe with proportions
  df_prop[, 1:3] <- df_prop[, 1:3] * 100
  df_prop$cluster <- rownames(df_prop)
  # df_prop$cluster <- as.numeric(df_prop$cluster)
  df_prop <- df_prop[order(x = df_prop$pct_sum, decreasing = T), ]
  # # filter out small clusters, that are less than 0.6%
  df_prop$Big <- df_prop$sum > 49
  
  #  df_prop <- df_prop[df_prop$sum > 49,]
  
  df_prop$category <- apply(df_prop[, c("human", "mouse", "pig")], 1, function(x) categorize_cluster(x['human'], x['mouse'], x['pig']))
  
  df_prop$category <- gsub(pattern = 'human.human', replacement = 'human', x = df_prop$category)
  df_prop$category <- gsub(pattern = 'pig.pig', replacement = 'pig', x = df_prop$category)
  df_prop$category <- gsub(pattern = 'mouse.mouse', replacement = 'mouse', x = df_prop$category) 
  my_prop[[name]] <- df_prop
  
  df_prop[df_prop$Big == FALSE, 'category'] <- 'too_small'
  my_vec <- df_prop$category
  names(my_vec) <- df_prop$cluster
  
  my_data <- all_datasets[[name]]
  Idents(my_data) <- 'specific_clusters'
  my_data$specific_clusters <- factor(my_data$specific_clusters)
  
  my_data$specific_clusters_2 <- my_data$specific_clusters
    
  non_pres <- levels(my_data$specific_clusters)[!levels(my_data$specific_clusters) %in% names(my_vec)]  
  
  my_data$specific_clusters_2[my_data$specific_clusters_2 %in% non_pres] <- NA
  
  Idents(my_data) <- 'specific_clusters_2'
  
  my_data <- RenameIdents(my_data, my_vec)
  my_vec_2 <- Idents(my_data)
  my_data$specificity_category <- my_vec_2
  
  all_datasets[[name]]$specificity_category <- my_vec_2
  
  my_data$smc_phenotypes <- factor(my_data$smc_phenotypes)
  
  my_species = DimPlot_scCustom(
  my_data[,!is.na(my_data$specificity_category)],
    reduction = 'umap',
    group.by = c('specificity_category'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_cat, repel = T, label.size = 2.7
  ) + 
    scale_color_manual(values = pal_cat) +
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
    scale_color_manual(values = pal_cat) + theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  my_cell_type = DimPlot_scCustom(my_data,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_all, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_all) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
# fig = ggdraw() +
#   draw_plot(my_cell_type, x = 0.0, y = .5001, width = .6, height = .55) +
#   draw_plot(my_species , x = 0.0, y = 0.00005, width = .6, height = .55)  +
#   draw_plot_label(label = new_names[name], size = 12,
#                   x = c(-0.01), y = c(1.005), family = 'Helvetica')
fig = ggdraw() +
  draw_plot(my_cell_type, x = 0.0, y = .54, width = .99, height = .45) +
  draw_plot(my_species, x = 0.0, y = 0.00005, width = .99, height = .55)  +
  draw_plot_label(label = new_names[name], size = 12,
                  x = c(0), y = c(1.005), family = 'Helvetica')


  
  # fig = plot_grid(my_species,
  #                 my_cell_type,
  #                 ncol = 1, rel_widths = c(0.5, 1),
  #                 rel_heights  = c(0.5, 1))
  
  myplots[[name]] <- fig
  
}

saveRDS(myplots, file = 'species_specific_clusters_category.rds')
saveRDS(my_prop, file = 'species_specific_clusters_categoryproportions_df.rds')
p <- plot_grid(plotlist = myplots, nrow = 2, ncol = 6, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Arial', label_x = -0.1)

ggsave(filename = 'UMAP_specific_clusters_species_categories.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 3)

```

# Compute mean and sd for composition of shared and specific clusters
```{r}
my_prop <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/species_specific_clusters_categoryproportions_df.rds")
 


# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% filter(sum > 50) %>% 
    filter(category == "shared") %>%
    summarise(sum_pct = sum(pct_sum, na.rm = TRUE)) %>%
    pull(sum_pct)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df, na.rm = TRUE)
sd_sums   <- sd(sums_each_df, na.rm = TRUE)
cat("Mean of sums for shared category:", mean_sums, "\n") # 75.575
cat("SD of sums for shared category:  ", sd_sums,   "\n") # 20.13956 


median_sums <- median(sums_each_df, na.rm = TRUE)
IQR_value <- IQR(sums_each_df, na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df, 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df, 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 77.05 
cat("1st Quartile (Q1): ", Q1, "\n") # 65
cat("3rd Quartile (Q3): ", Q3, "\n") # 89.825
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 24.825


# Mouse-specific 'Terminal'
# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 50) %>% 
    filter(category == "mouse-specific") %>%
    summarise(mouse = sum(mouse, na.rm = TRUE)) %>%
    pull(mouse)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0][-3], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0][-3], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") # 90.98468 
cat("SD of sums for the category:  ", sd_sums,   "\n") #  7.291969 

median_sums <- median(sums_each_df[sums_each_df != 0][-3], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0][-3], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0][-3], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0][-3], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 92.3445 
cat("1st Quartile (Q1): ", Q1, "\n") # 84.03756
cat("3rd Quartile (Q3): ", Q3, "\n") # 97.67442
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 13.63686

# # Human-pig-specific contractile
# # For each data frame, sum pct_sum where category == "shared"
# sums_each_df <- sapply(my_prop, function(df) {
#   df %>% 
#     filter(sum > 50) %>% 
#     filter(category == "human-pig-specific") %>%
#     filter(grepl( 'SMC', cluster)) %>%
#     summarise(mouse = sum(mouse, na.rm = TRUE)) %>%
#     pull(mouse)
# })
# 
# #  Now sums_each_df is a numeric vector (one sum per data frame)
# # Calculate mean and sd of those sums across the data frames
# mean_sums <- mean(sums_each_df, na.rm = TRUE)
# sd_sums   <- sd(sums_each_df, na.rm = TRUE)
# 
# cat("Mean of sums for the category:", mean_sums, "\n") #  10.09659
# cat("SD of sums for the category:  ", sd_sums,   "\n") # 5.57072

# DLX5
# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 50) %>% 
    # filter(category == "human-pig-specific") %>%
    filter(grepl('DLX5', cluster)) %>%
    summarise(sum_pct = sum(pct_sum, na.rm = TRUE)) %>% 
    pull(sum_pct)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df, na.rm = TRUE)
sd_sums   <- sd(sums_each_df, na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  3.191667
cat("SD of sums for the category:  ", sd_sums,   "\n") # 1.440618 



sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 50) %>% 
    # filter(category == "human-pig-specific") %>%
    filter(grepl( 'DLX5', cluster)) %>%
    summarise(human = sum(human, na.rm = TRUE)) %>%
    pull(human)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  91.31739
cat("SD of sums for the category:  ", sd_sums,   "\n") #  10.43497

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 95.94595 
cat("1st Quartile (Q1): ", Q1, "\n") # 83.62283
cat("3rd Quartile (Q3): ", Q3, "\n") # 99.6988 
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 16.07597

# Pericytes
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 50) %>% 
    # filter(category == "human-pig-specific") %>%
    filter(grepl( 'PC', cluster)) %>%
    summarise(human = sum(human, na.rm = TRUE)) %>%
    pull(human)
})

sums_each_df_2 <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 50) %>% 
    # filter(category == "human-pig-specific") %>%
    filter(grepl( 'PC', cluster)) %>%
    summarise(pig = sum(pig, na.rm = TRUE)) %>%
    pull(pig)
})

sums_each_df <- sums_each_df + sums_each_df_2
#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  96.65557 
cat("SD of sums for the category:  ", sd_sums,   "\n") #  4.560509

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 99.02624 
cat("1st Quartile (Q1): ", Q1, "\n") # 94.88636
cat("3rd Quartile (Q3): ", Q3, "\n") # 100
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 5.113636



# Contractile pig-human-enriched
for (i in names(my_prop)) {
  my_prop[[i]]$pig_human <- my_prop[[i]]$human + my_prop[[i]]$pig
}


sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 50) %>% 
    filter(pig_human > 90) %>% 
    # filter(category == "human-pig-specific") %>%
    filter(grepl( 'SMC1', cluster)) %>%
    summarise(pig_human = mean(pig_human, na.rm = TRUE)) %>%
    pull(pig_human)
})

# sums_each_df_2 <- sapply(my_prop, function(df) {
#   df %>% 
#     filter(sum > 50) %>% 
#     # filter(category == "human-pig-specific") %>%
#     filter(grepl( 'SMC1', cluster)) %>%
#     summarise(pig = sum(pig, na.rm = TRUE)) %>%
#     pull(pig)
# })

# sums_each_df <- sums_each_df + sums_each_df_2
#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  94.17904
cat("SD of sums for the category:  ", sd_sums,   "\n") #  0.244814 


median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 94.13562 
cat("1st Quartile (Q1): ", Q1, "\n") # 94.88636
cat("3rd Quartile (Q3): ", Q3, "\n") # 100
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 0.2609978
```


```{r}
new_names <- c("HE_CCA","HE_RPCA","HHC_CCA",  "HHC_RPCA","O2O_CCA", "O2O_RPCA", "HE_scANVI","HE_scVI",   "HHC_scANVI" ,    "HHC_scVI", "O2O_scANVI"  ,"O2O_scVI" )
names(new_names) <- names(all_datasets)

new_names <- new_names[c(2,4,6,
                         1,3,5,
                         7,9,11,
                         8,10,12)]

specific_clusters_df <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/specific_clusters_df.rds")

my_data <- all_datasets$many_higher_expr_seuratRPCA

my_vec <- c(rep(NA, length(colnames(my_data))))
names(my_vec) <- colnames(my_data)
```

```{r}
### 1) Precompute cluster membership sets

# Let's say we build a list to store membership for each (name, k):
# E.g.: membership_list[[name]][[k]] = character vector of cell names
membership_list <- list()

for (name in names(new_names)) {
  membership_list[[name]] <- list()
  for (k in colnames(specific_clusters_df)[-1]) {
    # gather all cell names that are in specific_clusters_df[name, k]
    target_cluster <- strsplit(specific_clusters_df[name, k], ", ")[[1]]
    # subset the $specific_clusters from all_datasets[[name]]
    # fetch the names of those cells
    # We can store as a character vector
    membership_list[[name]][[k]] <-
      names(all_datasets[[name]]$specific_clusters[
        all_datasets[[name]]$specific_clusters %in% target_cluster
      ])
  }
}

### 2) For each k, combine membership across 'names' if that’s needed
# Because you do: "all_spec <- c(all_spec, ...)" across all "name in new_names"
# we can do that once per k:

combined_membership <- list()  # combined_membership[[k]] = all_spec for that k

for (k in colnames(specific_clusters_df)[-1]) {
  # We'll do a union or concatenation across all name
  all_spec_for_k <- c()
  for (name in names(new_names)) {
    all_spec_for_k <- c(all_spec_for_k, membership_list[[name]][[k]])
  }
  # store it
  combined_membership[[k]] <- all_spec_for_k
}

### 3) Now, loop over i in names(my_vec) and check membership count
# We can define a function is i in all_spec_for_k more than 5 times?

for (i in names(my_vec)) {
  # optional: track if we found a match
  assigned <- FALSE

  for (k in colnames(specific_clusters_df)[-1]) {
    # we have combined_membership[[k]] which is a vector of cell names
    # We do sum(...) or we can do table etc.

    count_i <- sum(combined_membership[[k]] == i)
    if (count_i > 3) {
      my_vec[i] <- k
      assigned <- TRUE
      break  # we found a match, so let's stop checking further k
    }
  }

  if (!assigned) {
    my_vec[i] <- "NS"
  }
}
table(my_vec) # Mouse_human_specific 2 cells, Too_small_cluster 128 cells
my_vec[my_vec %in% c('Mouse_human_specific', 'NS')] <- 'Too_small_cluster'

my_data$consensus <- my_vec[colnames(my_data)]
```

```{r}
p <- DimPlot_scCustom(my_data, group.by = 'consensus', reduction = 'umap') + DimPlot_scCustom(my_data, group.by = 'smc_phenotypes', reduction = 'umap') & theme(aspect.ratio = 1)
ggsave(filename = 'UMAP_HERPCA_consensus_at_least_4int.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 8, height = 3, scale = 3)
```


```{r}
for (name in names(all_datasets)) {
  all_datasets[[name]]$consensus <-  my_vec[colnames(all_datasets[[name]])]
  all_datasets[[name]]$specificity_category_original <-
    all_datasets[[name]]$specificity_category
  all_datasets[[name]]$specificity_category <- all_datasets[[name]]$consensus
  all_datasets[[name]]$specificity_category <- factor(all_datasets[[name]]$specificity_category)
}

```

```{r}
saveRDS(all_datasets, file = 'r_py_all_datasets_processed_with_specific_clusters.rds')
```

# Shared phenotypes
```{r}
all_datasets <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/r_py_all_datasets_processed_with_specific_clusters.rds")
```


```{r}
shared_data <- list()
for (name in names(all_datasets)) {
  shared_data[[name]] <- all_datasets[[name]][, all_datasets[[name]]$specificity_category == 'Shared' ]
}


best_resolution <- function(dataset, resolutions, co2, type_integration) {
  best_res <- NULL
  best_clusters <- NULL
  type_integration <- type_integration
  dataset <- RunUMAP(
      object = dataset,
      reduction = type_integration,
      dims = 1:co2,
      verbose = FALSE
    ) %>%
      FindNeighbors(reduction = type_integration,
                    dims = 1:co2,
                    verbose = FALSE)
  
  # Iterate over resolutions in decreasing order
  for (res in rev(resolutions)) {
    # Run UMAP and clustering for current resolution
    dataset <-
      FindClusters(object = dataset,
        resolution = res,
        verbose = FALSE
      )
    
    # Extract cluster assignments
    clusters <- dataset$seurat_clusters
    
    # Check if the number of clusters is between 3 and 5
    num_clusters <- length(unique(clusters))
    if (num_clusters < 3 || num_clusters > 5) {
      next  # Skip if the number of clusters is not between 3 and 5
    }
    
    # Check if there are species-specific clusters (splitting by 'species')
    species <- dataset$species
    cluster_table <- table(clusters, species)
    
    # Check if any cluster is dominated by a single species
    species_specific_clusters <- apply(cluster_table, 1, function(x) {
      max(x) / sum(x) > 0.8  # A cluster is considered species-specific if one species dominates > 90%
    })
    
    if (any(species_specific_clusters)) {
      next  # Skip if there are species-specific clusters
    }
    
    # If it passed the checks, save the resolution and clusters
    best_res <- res
    best_clusters <- clusters
    break  # Stop once we find the highest valid resolution
  }
  
  return(list(best_res = best_res, best_clusters = best_clusters, dataset = dataset))
}



# For RPCA and CCA
for (name in names(all_datasets)[1:6]) {
  DefaultAssay(shared_data[[name]]) <- 'integrated'
  
  shared_data[[name]] <- RunPCA(object = shared_data[[name]], npcs = 40)
 
  # Find significant PCs
  stdv <- shared_data[[name]][["pca"]]@stdev
  sum.stdv <- sum(shared_data[[name]][["pca"]]@stdev)
  percent.stdv <- (stdv / sum.stdv) * 100
  cumulative <- cumsum(percent.stdv)
  # co1 <- which(cumulative > 85 & percent.stdv < 5)[1]
  co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] -
                       percent.stdv[2:length(percent.stdv)]) > 0.1),
              decreasing = T)[1] + 1

  result <- best_resolution(shared_data[[name]], resolutions = c(0.1, 0.15,0.16, 0.17, 0.18, 0.19, 0.2,0.25, 0.3, 0.35, 0.4), co2 = 13, type_integration = 'pca')
  
   # If a valid resolution and clusters were found, update the dataset
  if (!is.null(result$best_res)) {
    shared_data[[name]]$seurat_clusters <- result$best_clusters
    shared_data[[name]]$umap <- result$dataset[["umap"]]  # Save UMAP to the dataset
    message(paste("Dataset", name, "selected best resolution:", result$best_res))
  } else {
    warning(paste("Dataset", name, "did not meet clustering criteria. Skipping..."))
  }
}

# For scANVI
for (name in names(all_datasets)[c(7,9,11)]) {
  DefaultAssay(shared_data[[name]]) <- 'RNA'

   # Call the function to find the best clustering for each dataset
  result <- best_resolution(shared_data[[name]], resolutions =  c(0.1, 0.15,0.16, 0.17, 0.18, 0.19, 0.2,0.25, 0.3, 0.35, 0.4), co2 = 10, type_integration = 'X_scANVI')
  
   # If a valid resolution and clusters were found, update the dataset
  if (!is.null(result$best_res)) {
    shared_data[[name]]$seurat_clusters <- result$best_clusters
    shared_data[[name]]$umap <- result$dataset[["umap"]]  # Save UMAP to the dataset
    message(paste("Dataset", name, "selected best resolution:", result$best_res))
  } else {
    warning(paste("Dataset", name, "did not meet clustering criteria. Skipping..."))
  }
}



# For scVI
for (name in names(all_datasets)[c(8, 10, 12)]) {
  DefaultAssay(shared_data[[name]]) <- 'RNA'
  
  # Call the function to find the best clustering for each dataset
  result <- best_resolution(shared_data[[name]], resolutions =  c(0.1, 0.15,0.16, 0.17, 0.18, 0.19, 0.2,0.25, 0.3, 0.35, 0.4), co2 = 30, type_integration = 'X_scVI')
  
   # If a valid resolution and clusters were found, update the dataset
  if (!is.null(result$best_res)) {
    shared_data[[name]]$seurat_clusters <- result$best_clusters
    shared_data[[name]]$umap <- result$dataset[["umap"]]  # Save UMAP to the dataset
    message(paste("Dataset", name, "selected best resolution:", result$best_res))
  } else {
    warning(paste("Dataset", name, "did not meet clustering criteria. Skipping..."))
  }
}


```

```{r}
# Create a plot for all datasets with UMAP and colored by 'seurat_clusters'
plot_list <- list()

for (name in names(shared_data)) {
  # Retrieve UMAP coordinates and clusters
  umap_coords <- Embeddings(shared_data[[name]], reduction = "umap")
  clusters <- shared_data[[name]]$seurat_clusters
  
  # Create a data frame for plotting
  umap_data <- data.frame(UMAP1 = umap_coords[, 1], UMAP2 = umap_coords[, 2], Cluster = as.factor(clusters))
  
  # Create UMAP plot for this dataset
  p <- ggplot(umap_data, aes(x = UMAP1, y = UMAP2, color = Cluster)) +
    geom_point(size = 1) +
    labs(title = paste("UMAP -", name)) +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Add the plot to the list
  plot_list[[name]] <- p
}

# Combine all plots into a single grid
grid.arrange(grobs = plot_list, ncol = 4)  # Adjust ncol for the number of columns

```

```{r}
markers <- list()
for (name in names(shared_data)) {
  Idents(shared_data[[name]]) <- 'seurat_clusters'
  tmp <- FindAllMarkers(shared_data[[name]], assay = 'RNA_genes', only.pos = T, max.cells.per.ident = 1000)
  markers[[name]] <- tmp
}
saveRDS(markers, 'markers_shared_phenotypes.12.rds')
```



```{r}
# Identify corresponding clusters based on top 100 markers. Cluster with highest ACTA2 - Contractile, cluster with highest LUM - Terminal, others - leave the original name


# Initialize the list to store top markers for each dataset and cluster combination
top_markers_list <- list()

# Loop through the selected datasets
for (name in names(shared_data)[c(1, 2, 3, 4, 5, 6, 7, 8, 10)]) {
  # Find markers for each cluster
  markers <- FindAllMarkers(shared_data[[name]], only.pos = TRUE, assay = 'RNA_genes', max.cells.per.ident = 1000)
  
  # Loop through each cluster and find the top 100 genes based on avg_log2FC
  for (cluster in unique(markers$cluster)) {
    cluster_markers <- markers[markers$cluster == cluster, ]
    top_cluster_markers <- cluster_markers[order(cluster_markers$avg_log2FC, decreasing = TRUE), ][1:100, ]
    
    # Create a name for the list entry using both dataset name and cluster name
    list_name <- paste(name, as.character(cluster), sep = ".")
    
    # Store the top markers for each dataset and cluster in the list
    top_markers_list[[list_name]] <- top_cluster_markers
  }
}


# Function to calculate overlap of top 100 markers between two datasets
calculate_overlap <- function(dataset1_markers, dataset2_markers) {
  # Extract gene names from both datasets
  genes1 <- rownames(dataset1_markers)
  genes2 <- rownames(dataset2_markers)
  
  # Calculate the overlap of gene names
  overlap_genes <- intersect(genes1, genes2)
  overlap_count <- length(overlap_genes)
  
  return(overlap_count)
}

# Initialize an empty matrix to store the overlap counts
dataset_names <- names(top_markers_list)
overlap_matrix <- matrix(0, nrow = length(dataset_names), ncol = length(dataset_names), 
                         dimnames = list(dataset_names, dataset_names))

# Loop through each pair of datasets and clusters to calculate the overlap
for (name1 in dataset_names) {
  for (name2 in dataset_names) {
    if (name1 != name2) {
      # Extract the dataset and cluster names from the list names
      dataset1 <- strsplit(name1, "\\.")[[1]][1]
      cluster1 <- strsplit(name1, "\\.")[[1]][2]
      dataset2 <- strsplit(name2, "\\.")[[1]][1]
      cluster2 <- strsplit(name2, "\\.")[[1]][2]
      
      # Get the top 100 markers for both clusters
      markers1 <- top_markers_list[[name1]]
      markers2 <- top_markers_list[[name2]]
      
      # Calculate the overlap between the top 100 markers of the two clusters
      overlap <- calculate_overlap(markers1, markers2)
      
      # Store the overlap count in the matrix
      overlap_matrix[name1, name2] <- overlap
    }
  }
}

# Convert matrix to dataframe for easier visualization
overlap_df <- as.data.frame(overlap_matrix)

for(i in 1:length(rownames(overlap_df))) {
  overlap_df[i, 'Integration'] <-
    strsplit(rownames(overlap_df)[i], "\\.")[[1]][1]
}

# Display the overlap matrix
print(overlap_df)

pheatmap(overlap_df[,1:34])



####

my_celltypes_clusters <- data.frame(
  Integration_1 = character(),
  Integration_2 = character(),
  Cluster = character(),
  Correspond_Cluster = character()
)

my_celltypes_clusters_subset <- data.frame(
  Integration_1 = character(),
  Integration_2 = character(),
  Cluster = character(),
  Correspond_Cluster = character()
)



for (j in unique(overlap_df$Integration)) {
  overlap_df_subset <- overlap_df[overlap_df$Integration == j, 1:34]
  for (i in 1:length(colnames(overlap_df_subset))) {
    my_celltypes_clusters_subset[i, 'Integration_2'] <- j
    
    my_celltypes_clusters_subset[i, 'Cluster'] <-
      colnames(overlap_df_subset)[i]
    
    
    my_celltypes_clusters_subset[i, 'Correspond_Cluster'] <-
      ifelse(is.null(rownames(overlap_df_subset)[which.max(overlap_df_subset[, i])]) == TRUE,
             0,
             rownames(overlap_df_subset)[which.max(overlap_df_subset[, i])])
    
    my_celltypes_clusters_subset[i, 'Integration_1'] <-
      strsplit(my_celltypes_clusters_subset[i, 'Cluster'], "\\.")[[1]][1]
    
    my_celltypes_clusters <-
      rbind(my_celltypes_clusters, my_celltypes_clusters_subset)
  }
}

my_celltypes_clusters <- my_celltypes_clusters %>%
  filter(Cluster != Correspond_Cluster) %>%
  filter(Integration_1 != Integration_2) %>% 
  distinct()


# So far we have identified pairs of corresponding clusters. Now we need to name them to Contractile, Intermediate, and Terminal
res <- pheatmap(overlap_df[, 1:34])
mtcars.clust <- cbind(overlap_df[, 1:34], 
                      cluster = cutree(res$tree_row, 
                                       k = 5))


for (i in 1:length(rownames(mtcars.clust))) {
  x <- strsplit(rownames(mtcars.clust)[i], "\\.")[[1]][1]
  z <- strsplit(rownames(mtcars.clust)[i], "\\.")[[1]][2]
  
  
  shared_data[[x]]$cluster_shared[shared_data[[x]]$seurat_clusters == z] <-
    mtcars.clust[rownames(mtcars.clust)[i], 'cluster']
  
}


markers <- list()
for (name in names(shared_data)) {
  Idents(shared_data[[name]]) <- 'cluster_shared'
  tmp <- FindAllMarkers(shared_data[[name]], assay = 'RNA_genes', only.pos = T, max.cells.per.ident = 1000)
  markers[[name]] <- tmp
}

table(shared_data$many_higher_expr_scVI$seurat_clusters, shared_data$many_higher_expr_scVI$cluster_shared)
saveRDS(markers, 'markers_shared_phenotypes.12_corresponding_groups.rds')



for(x in names(shared_data)[c(1, 2, 3, 4, 5, 6, 7, 8, 10)]) {
  
     
  # Initialize the cluster_shared_pheno column as NA
  shared_data[[x]]$cluster_shared_pheno <- 0
  
  # Check if each cluster value exists before assigning phenotype labels
  if (any(shared_data[[x]]$cluster_shared == 1)) {
    shared_data[[x]]$cluster_shared_pheno[shared_data[[x]]$cluster_shared == 1] <-  'Intermediate'
  }
  
  if (any(shared_data[[x]]$cluster_shared == 2)) {
    shared_data[[x]]$cluster_shared_pheno[shared_data[[x]]$cluster_shared == 2] <- 'Terminal'
  }
  
  if (any(shared_data[[x]]$cluster_shared == 3)) {
    shared_data[[x]]$cluster_shared_pheno[shared_data[[x]]$cluster_shared == 3] <- 'Contractile'
  }
  
  if (any(shared_data[[x]]$cluster_shared == 4)) {
    shared_data[[x]]$cluster_shared_pheno[shared_data[[x]]$cluster_shared == 4] <- 'Proliferating'
  }
  
  if (any(shared_data[[x]]$cluster_shared == 5)) {
    shared_data[[x]]$cluster_shared_pheno[shared_data[[x]]$cluster_shared == 5] <- 'Other'
  }
}
```

```{r}
myplots <- list()


for(name in names(shared_data)[c(1, 2, 3, 4, 5, 6, 7, 8, 10)]) {
 
  my_plot = DimPlot(
  object = shared_data[[name]],
    reduction = 'umap',
    group.by = c('cluster_shared_pheno'),
    pt.size = 0.5,
    shuffle = TRUE
  ) + theme(legend.position = 'bottom') + text_sizes + theme(plot.title = element_blank()) + 
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
    scale_color_manual(values = pal_celltypes) + theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  myplots[[name]] <- my_plot
}

p <- plot_grid(plotlist = myplots, nrow = 3, ncol = 3, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Helvetica', label_x = -0.1)
print(p)
```


```{r}
saveRDS(shared_data, 'shared_datasets_clusered.rds')
```


```{r}
my_data <- all_datasets$many_higher_expr_seuratRPCA
my_data <- my_data[,my_data$consensus == 'Shared']
DefaultAssay(my_data) <- 'integrated'

my_data <- RunPCA(object = my_data, npcs = 40)
ElbowPlot(my_data, ndims = 40)
# Find significant PCs
stdv <- my_data[["pca"]]@stdev
sum.stdv <- sum(my_data[["pca"]]@stdev)
percent.stdv <- (stdv / sum.stdv) * 100
cumulative <- cumsum(percent.stdv)
co1 <- which(cumulative > 85 & percent.stdv < 5)[1]
co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] -
                     percent.stdv[2:length(percent.stdv)]) > 0.1),
            decreasing = T)[1] + 1
co1
co2
```


```{r}
my_data <- 
  RunUMAP(object = my_data, reduction = "pca",
          dims = 1:13,
          verbose = FALSE) %>%
  FindNeighbors(reduction = "pca",
                dims = 1:13,
                verbose = FALSE) %>%
  FindClusters(resolution = c(0.1, 0.15,0.17, 0.18,0.19, 0.2, 0.3, 0.35, 0.4),
               verbose = FALSE)

DimPlot_scCustom(my_data, group.by = 'integrated_snn_res.0.17', split.by = 'species')
table(my_data$integrated_snn_res.0.17, my_data$species)
Idents(my_data) <- 'integrated_snn_res.0.17'
tmp <- FindAllMarkers(my_data, assay = 'RNA_genes', only.pos = T, max.cells.per.ident = 2000)
saveRDS(tmp, file = 'shared_clusters_markers_cell_categorization.rds')

my_data$smc_phenotypes <- as.factor(as.character(my_data$smc_phenotypes))

DimPlot_scCustom(my_data, group.by = 'integrated_snn_res.0.17', reduction = 'umap') + DimPlot_scCustom(my_data, group.by = 'smc_phenotypes', reduction = 'umap') 
DimPlot_scCustom(my_data, group.by = 'integrated_snn_res.0.17', split.by = 'species', reduction = 'umap')
Idents(my_data) <- 'integrated_snn_res.0.17'
new_names <-
  c("Terminal", "Transitional",
    "Contractile" , 'Proliferating')
names(new_names) <- levels(Idents(my_data))

my_data <-
  Seurat::RenameIdents(object = my_data, new_names)
new_names <- Idents(my_data)
my_data$shared_clusters <-
  new_names

p1 <- DimPlot_scCustom(my_data, group.by = 'shared_clusters', split.by = 'species', colors_use = pal_celltypes_pig[c(3,1,2,5)], aspect_ratio = 1, split_seurat = T, reduction = 'umap')
p1
ggsave(filename = 'reclustered_shared_SMCs_split_cell_categorization.pdf', plot = p1, width = 6, height = 6, units = 'cm', scale = 3)
saveRDS(my_data, 'HE_RPCA_cell_categorization.rds')
##############
##############
# Perform differential expression analysis
Idents(my_data) <- 'shared_clusters'

# #scVI does not have all genes, only variable. Transfer labels for shared clusters from scVI to HE_RPCA
# my_data_scVI <- my_data
# my_data <- all_datasets_el$many_higher_expr_seuratRPCA
# my_data <- my_data[,colnames(my_data) %in% colnames(my_data_scVI)]
# my_data$shared_clusters <- my_data_scVI$shared_clusters[colnames(my_data)]
# my_data$RNA_snn_res.0.3 <- my_data_scVI$RNA_snn_res.0.3[colnames(my_data)]

DefaultAssay(my_data) <- 'RNA_genes'
markers_contractile_vs_terminal <- FindMarkers(my_data, ident.1 = "Contractile", ident.2 = "Terminal")
markers_contractile_vs_transitional <- FindMarkers(my_data, ident.1 = "Contractile", ident.2 = "Transitional")
markers_terminal_vs_transitional <- FindMarkers(my_data, ident.1 = "Terminal", ident.2 = "Transitional")

# Identify genes uniquely high in contractile vs terminal but also present in transitional
gene1_candidates <- subset(markers_contractile_vs_terminal, avg_log2FC > 0.1 & p_val_adj < 0.05)
gene1_candidates <- intersect(rownames(gene1_candidates), rownames(subset(markers_contractile_vs_transitional, avg_log2FC > 1)))

# Identify genes uniquely high in terminal vs contractile but also present in transitional
gene2_candidates <- subset(markers_terminal_vs_transitional, avg_log2FC > 1 & p_val_adj < 0.05)
gene2_candidates <- intersect(rownames(gene2_candidates), rownames(subset(markers_contractile_vs_transitional, avg_log2FC < 0.7)))

# Plotting expression profiles for selected candidate genes
if(length(gene1_candidates) > 0 && length(gene2_candidates) > 0) {
  FeaturePlot(my_data, features = c(gene1_candidates[1], gene2_candidates[1]), cols = c("blue", "red"), reduction = 'umap')
}

rownames(subset(markers_contractile_vs_transitional, avg_log2FC < 0))[rownames(subset(markers_contractile_vs_transitional, avg_log2FC < 0)) %in% rownames(subset(markers_terminal_vs_transitional, avg_log2FC < 0))]

FeaturePlot(object = my_data, features = c('PSMB1', 'THBS2'), blend = T, blend.threshold = 0.1, pt.size = 1)

#########################################################
# Assuming data is normalized and ready for analysis
# Get the expression matrix
Idents(my_data) <- 'shared_clusters'
data_mat <- GetAssayData(my_data, slot = "data", assay = 'RNA_genes')

# Binary matrix: 1 if expressed, 0 if not
binary_mat <- data_mat > 1  # Adjust this threshold as needed

# Get cluster identities
clusters <- Idents(my_data)

# Prepare an empty data frame to store the results
results_df <- data.frame(
  Gene1 = character(),
  Gene2 = character(),
  Cluster = character(),
  PercExprGene1 = numeric(),
  PercExprGene2 = numeric(),
  PercBothExpr = numeric(),
  stringsAsFactors = FALSE
)



evaluate_and_store_results <- function(gene1, gene2, binary_mat, clusters) {
  expr1 <- binary_mat[gene1, ]
  expr2 <- binary_mat[gene2, ]
  
  clusters_unique <- unique(clusters)
  
  for (cluster in clusters_unique) {
    cluster_cells <- which(clusters == cluster)
    num_cells <- length(cluster_cells)
    
    if (num_cells > 0) {  # Ensure division by zero is not performed
      # Calculate percentages
      perc_expr1 <- sum(expr1[cluster_cells] & !expr2[cluster_cells]) / num_cells * 100
      perc_expr2 <- sum(!expr1[cluster_cells] & expr2[cluster_cells]) / num_cells * 100
      perc_both <- sum(expr1[cluster_cells] & expr2[cluster_cells]) / num_cells * 100
      
      # Append results to the dataframe
      results_df <<- rbind(results_df, data.frame(
        Gene1 = gene1,
        Gene2 = gene2,
        Cluster = cluster,
        PercExprGene1 = perc_expr1,
        PercExprGene2 = perc_expr2,
        PercBothExpr = perc_both,
        stringsAsFactors = FALSE
      ))
    }
  }
}


tmp1 <- FindAllMarkers(object = my_data[,my_data$species == 'human'], assay = 'RNA_genes', only.pos = T)
tmp2 <- FindAllMarkers(object = my_data[,my_data$species == 'mouse'], assay = 'RNA_genes', only.pos = T)
tmp3 <- FindAllMarkers(object = my_data[,my_data$species == 'pig'], assay = 'RNA_genes', only.pos = T)

contr_g <- intersect(intersect(tmp1[tmp1$cluster == 'Contractile' & tmp1$avg_log2FC > 0.5, 'gene'] , tmp2[tmp2$cluster == 'Contractile' & tmp2$avg_log2FC > 0.5, 'gene'] ) ,  tmp3[tmp3$cluster == 'Contractile' & tmp3$avg_log2FC > 0.5, 'gene'] )
trans_g <- intersect(intersect(tmp1[tmp1$cluster == 'Transitional' & tmp1$avg_log2FC > 0, 'gene'] , tmp2[tmp2$cluster == 'Transitional' & tmp2$avg_log2FC > 0, 'gene'] ) ,  tmp3[tmp3$cluster == 'Transitional' & tmp3$avg_log2FC > 0, 'gene'] )
term_g <- intersect(intersect(tmp1[tmp1$cluster == 'Terminal' & tmp1$avg_log2FC > 0.5, 'gene'] , tmp2[tmp2$cluster == 'Terminal' & tmp2$avg_log2FC > 0.5, 'gene'] ) ,  tmp3[tmp3$cluster == 'Terminal' & tmp3$avg_log2FC > 0.5, 'gene'] )

gn <- unique(c(contr_g, trans_g, term_g))

for (i in 1:(length(gn) - 1)) {
  for (j in (i + 1):length(gn)) {
    evaluate_and_store_results(gn[i], gn[j], binary_mat, clusters)
  }
}
saveRDS(results_df, 'best_pair_shareds.rds')

# Ensure all relevant columns are in correct format, particularly not lists unless intended
results_df$PercExprGene1 <- as.numeric(as.character(results_df$PercExprGene1))
results_df$PercExprGene2 <- as.numeric(as.character(results_df$PercExprGene2))
results_df$PercBothExpr <- as.numeric(as.character(results_df$PercBothExpr))


# Assuming results_df is already populated and contains the columns:
# Gene1, Gene2, Cluster, PercExprGene1, PercExprGene2, PercBothExpr

# Base R approach using reshape()
wide_df_base <- reshape(results_df, 
                        timevar = "Cluster", 
                        idvar = c("Gene1", "Gene2"),
                        direction = "wide",
                        sep = "_")

# Check the output
print(wide_df_base)


# wide_df <- results_df %>%
#   pivot_wider(
#     names_from = Cluster, 
#     values_from = c(PercExprGene1, PercExprGene2, PercBothExpr),
#     names_glue = "{Cluster}_{.value}"  # Naming pattern to distinguish columns per cluster and metric
#   )
# 
# wide_df$Terminal_PercExprGene1 <- as.numeric(as.character(results_df$PercExprGene1))
# wide_df$PercExprGene2 <- as.numeric(as.character(results_df$PercExprGene2))
# wide_df$PercBothExpr <- as.numeric(as.character(results_df$PercBothExpr))


# Check the newly formatted dataframe
print(wide_df_base)
saveRDS(wide_df_base, 'best_pair_shared_wide_scVI.rds')
openxlsx::write.xlsx(x = wide_df_base, file = 'best_pair_shared_wide_cell_categorization.xlsx')

DefaultAssay(my_data) <- 'RNA_genes'
FeaturePlot(my_data, features = c('LTBP2', 'CNN1'),reduction = 'umap',pt.size = 0.7, order = T,  blend = T, cols = c("darkblue", "green", "magenta"), blend.threshold = 0.5) &DarkTheme() &NoAxes() & plot_annotation(title = 'Carotid HS')


##############
##############
# Make lists of genes that are conservatively expressed across species in the shared SMC phenotypes
gene_list <- list()
tmp <- list()

for (i in unique(my_data$shared_clusters)) {
  for (j in unique(my_data$species)) {
    exrMat <-
      GetAssayData(my_data[, my_data$shared_clusters == i &
                             my_data$species == j], slot = 'data', assay = 'RNA_genes')
    
    # Step 1: Determine if a gene is expressed in a cell
    # Assuming any value greater than 0 indicates expression
    is_expressed <- exrMat > 0
    
    # Step 2: Calculate the percentage of cells in which each gene is expressed
    percent_expressed <- rowSums(is_expressed) / ncol(exrMat) * 100
    
    # Step 3: Identify genes expressed in at least 10% of cells
    keep_genes <-
      names(percent_expressed)[percent_expressed >= 10]
    tmp[[j]] <- keep_genes
  }
   tmp_2 <-
   as.data.frame(sort(Reduce(intersect,list( tmp$pig, tmp$human, tmp$mouse)), decreasing = F) )
   colnames(tmp_2) <- paste('conservative_genes', i)
  gene_list[[i]] <- tmp_2
}

openxlsx::write.xlsx(gene_list, file = 'conservative_genes_expressed_min10pct_per_phenotype_cell_categorization.xlsx', overwrite = T)

table(my_data$integrated_snn_res.0.17, my_data$species)

# Make balanced dataset
set.seed(1984)
my_data_balanced <- my_data[,colnames(my_data) %in% c(
  sample(names(my_data$species[my_data$species == 'human' & my_data$integrated_snn_res.0.17 == 0]), size = 600, replace = F),
  sample(names(my_data$species[my_data$species == 'human' & my_data$integrated_snn_res.0.17 == 1]), size = 500, replace = F),
  sample(names(my_data$species[my_data$species == 'human' & my_data$integrated_snn_res.0.17 == 2]), size = 500, replace = F),
  sample(names(my_data$species[my_data$species == 'human' & my_data$integrated_snn_res.0.17 == 3]), size = 6, replace = F),
  
  
  sample(names(my_data$species[my_data$species == 'pig' & my_data$integrated_snn_res.0.17 == 0]), size = 600, replace = F) ,
  sample(names(my_data$species[my_data$species == 'pig' & my_data$integrated_snn_res.0.17 == 1]), size = 500, replace = F) ,
  sample(names(my_data$species[my_data$species == 'pig' & my_data$integrated_snn_res.0.17 == 2]), size = 500, replace = F) ,
  sample(names(my_data$species[my_data$species == 'pig' & my_data$integrated_snn_res.0.17 == 3]), size = 6, replace = F) ,
  
  
  sample(names(my_data$species[my_data$species == 'mouse' & my_data$integrated_snn_res.0.17 == 0]), size = 600, replace = F),
  sample(names(my_data$species[my_data$species == 'mouse' & my_data$integrated_snn_res.0.17 == 1]), size = 500, replace = F),
  sample(names(my_data$species[my_data$species == 'mouse' & my_data$integrated_snn_res.0.17 == 2]), size = 180, replace = F),
  sample(names(my_data$species[my_data$species == 'mouse' & my_data$integrated_snn_res.0.17 == 3]), size = 6, replace = F)
  )]

table(my_data_balanced$species, my_data_balanced$integrated_snn_res.0.17)

DimPlot_scCustom(my_data_balanced, group.by = 'integrated_snn_res.0.17', split.by = 'species', reduction = 'umap')

Idents(my_data_balanced) <- 'integrated_snn_res.0.17'
DefaultAssay(my_data_balanced) <- 'RNA_genes'

tmp <- FindAllMarkers(my_data_balanced, only.pos = T, min.pct = 0.3)
# tmp <-  tmp[tmp$p_val_adj < 0.05,]
FeaturePlot_scCustom(my_data_balanced, features = c('THBS1', 'SERPINE2'), split.by = 'species', reduction = 'umap')


#### DEseq2
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98) == TRUE, final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)

CreatePseudoBulkData <- function(raw.data,normalized.data,sample.id,fun="sum")
{
    # fun can be "sum" or "mean"
    library(muscat)
    library(SingleCellExperiment)
    
    if (fun=="sum")
    {
        sce <- SingleCellExperiment(assays=list(counts=raw.data),colData=DataFrame(sample_id=sample.id))
    } else if (fun=="mean") {
        sce <- SingleCellExperiment(assays=list(counts=normalized.data),colData=DataFrame(sample_id=sample.id))
    }
    pb <- aggregateData(sce, assay = "counts", fun=fun, by=c("sample_id"))
    return(pb)
}
    
# Make counts slot with gene names
my_counts <-
  GetAssayData(object = my_data_balanced,
               assay = 'RNA',
               slot = 'counts')
rownames(my_counts) <-
  gsub(pattern = '-0',
       replacement = '',
       x = rownames(my_counts))
rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

my_counts <-
  my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

rownames(final_species_genes) <-
  final_species_genes$`Pig Ensembl ID`
df <-
  data.frame(
    pig_ens_data = rownames(my_counts),
    transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'],
    transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID']
  )

rownames(df) <- df$pig_ens_data
rownames(my_counts) <-
  df[rownames(my_counts), 'transl_human_gene']

# Run DESeq2 across species within each shared SMC cluster (pairwise)
deseq2_res <- list()
for(i in unique(my_data_balanced$shared_clusters)) {
  my_data_balanced_subset <-
    my_data_balanced[, my_data_balanced$shared_clusters == i]
  pb <-
    CreatePseudoBulkData(
      raw.data = my_counts[, colnames(my_counts) %in% colnames(my_data_balanced_subset)],
      sample.id = my_data_balanced_subset$orig.ident,
      fun = 'sum'
    )
  
  bca <-
    as.data.frame.matrix(table(my_data_balanced_subset$orig.ident, my_data_balanced_subset$species))
  group <- apply(bca, 1, function(x)
    colnames(bca)[which.max(x)])
  
  
  #Make the model and contrast for statistical testing
  df_con <- DataFrame(condition = factor(group))
  pb_counts <- assay(pb)
  mode(pb_counts) <- "integer"
  dge <- DESeqDataSetFromMatrix(pb_counts,
                                colData = df_con,
                                design = ~ 0 + condition)
  dds <- DESeq(dge, betaPrior = FALSE)
  
  hp_res <-
    results(
      dds,
      contrast = c("condition", "human", "pig"),
      independentFiltering = TRUE,
      alpha = 0.05,
      pAdjustMethod = "BH",
      parallel = FALSE
    )
  hp_res <-
    lfcShrink(
      dds,
      contrast = c("condition", "human", "pig"),
      res = hp_res,
      type = 'ashr'
    )
  hp_res <-
    as.data.frame(hp_res@listData, row.names = hp_res@rownames)
  
  
  colnames(hp_res) <- paste('HP', colnames(hp_res), sep = '_')
  
  hm_res <-
    results(
      dds,
      contrast = c("condition", "human", "mouse"),
      independentFiltering = TRUE,
      alpha = 0.05,
      pAdjustMethod = "BH",
      parallel = FALSE
    )
  hm_res <-
    lfcShrink(
      dds,
      contrast = c("condition", "human", "mouse"),
      res = hm_res,
      type = 'ashr'
    )
  hm_res <-
    as.data.frame(hm_res@listData, row.names = hm_res@rownames)
  colnames(hm_res) <- paste('HM', colnames(hm_res), sep = '_')
  
  pm_res <-
    results(
      dds,
      contrast = c("condition", "pig", "mouse"),
      independentFiltering = TRUE,
      alpha = 0.05,
      pAdjustMethod = "BH",
      parallel = FALSE
    )
  pm_res <-
    lfcShrink(
      dds,
      contrast = c("condition", "pig", "mouse"),
      res = pm_res,
      type = 'ashr'
    )
  pm_res <-
    as.data.frame(pm_res@listData, row.names = pm_res@rownames)
  colnames(pm_res) <- paste('PM', colnames(pm_res), sep = '_')
  
  res_all_org <- cbind(hp_res, hm_res, pm_res)
  res_all_org$gene <- rownames(res_all_org)
  res_all_org$exp_PIG <- rowMeans2(expm1(subset(my_data_balanced_subset, species == 'pig')[["RNA_genes"]]@data[rownames(res_all_org), ]))
  res_all_org$exp_HUMAN <- rowMeans2(expm1(subset(my_data_balanced_subset, species == 'human')[["RNA_genes"]]@data[rownames(res_all_org), ]))
  res_all_org$exp_MOUSE <- rowMeans2(expm1(subset(my_data_balanced_subset, species == 'mouse')[["RNA_genes"]]@data[rownames(res_all_org), ]))
  
  res_all_org$pct_PIG <- rowSums(subset(my_data_balanced_subset, species == 'pig')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) / ncol(subset(my_data_balanced_subset, species == 'pig')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) * 100
  res_all_org$pct_HUMAN <- rowSums(subset(my_data_balanced_subset, species == 'human')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) / ncol(subset(my_data_balanced_subset, species == 'human')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) * 100
  res_all_org$pct_MOUSE <- rowSums(subset(my_data_balanced_subset, species == 'mouse')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) / ncol(subset(my_data_balanced_subset, species == 'mouse')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) * 100

  deseq2_res[[i]] <- res_all_org
}
saveRDS(deseq2_res, file = 'deseq2.pairwise.shared.SMC.clusters_Elastic_cell_categorization.rds')
write.xlsx(deseq2_res, 'deseq2.pairwise.shared.SMC.clusters_Elastic_cell_categorization.xlsx', overwrite = T)

#### Shared coronary SMC clusters
all_datasets_cor <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/r_py_all_datasets_processed_with_specific_clusters_coronary_2.rds")
my_data_cor <- all_datasets$many_higher_expr_seuratRPCA
my_data_cor$run <- 'coronary'

my_data <- my_data_cor[,my_data_cor$specificity_category == 'Shared']
DefaultAssay(my_data) <- 'integrated'
# my_data <- RunPCA(object = my_data, npcs = 40)
ElbowPlot(my_data, ndims = 40)
my_data <- 
  RunUMAP(object = my_data, reduction = "pca",
          dims = 1:16,
          verbose = FALSE) %>%
  FindNeighbors(reduction = "pca",
                dims = 1:16,
                verbose = FALSE) %>%
  FindClusters(resolution = c(0.1, 0.15,  0.2, 0.25, 0.3, 0.35),
               verbose = FALSE)


table(my_data$integrated_snn_res.0.3, my_data$species)
# my_data$species_run <- paste(my_data$species, my_data$run, sep = '_')
DimPlot_scCustom(my_data, group.by = 'integrated_snn_res.0.2',  reduction = 'umap') 

my_data$smc_phenotypes <- as.factor(as.character(my_data$smc_phenotypes))

DimPlot_scCustom(my_data, group.by = 'integrated_snn_res.0.2', reduction = 'umap') + DimPlot_scCustom(my_data, group.by = 'smc_phenotypes', reduction = 'umap') 
DimPlot_scCustom(my_data, group.by = 'integrated_snn_res.0.2', split.by = 'species', reduction = 'umap')
DimPlot_scCustom(my_data, group.by = 'smc_phenotypes', split.by = 'species', reduction = 'umap')
Idents(my_data) <- 'integrated_snn_res.0.2'
tmp <- FindAllMarkers(my_data, assay = 'RNA_genes', only.pos = T)

new_names <-
  c('Terminal',
    "Contractile",
    "Fibroblasts",
     "Transitional",
    'IFN_SMC'
    )
names(new_names) <- levels(Idents(my_data))

my_data <-
  Seurat::RenameIdents(object = my_data, new_names)
new_names <- Idents(my_data)
my_data$shared_clusters <-
  new_names
pal_celltypes_pig <- c("#E18727FF", "#0072B5FF",  "#BC3C29FF", "#20854EFF", "#7876B1FF" ,"#FFDC91FF", "#EE4C97FF")
p1 <- DimPlot_scCustom(my_data, group.by = 'shared_clusters', split.by = 'species', colors_use = pal_celltypes_pig[c(3,2,6,1,7)], aspect_ratio = 1, split_seurat = T)
p1
ggsave(filename = 'reclustered_shared_SMCs_split_Cor.pdf', plot = p1, width = 6, height = 6, units = 'cm', scale = 3)
# Make lists of genes that are conservatively expressed across species in the shared SMC phenotypes
gene_list <- list()
tmp <- list()

for (i in unique(my_data$shared_clusters)) {
  for (j in unique(my_data$species)) {
    exrMat <-
      GetAssayData(my_data[, my_data$shared_clusters == i &
                             my_data$species == j], slot = 'data', assay = 'RNA_genes')
    
    # Step 1: Determine if a gene is expressed in a cell
    # Assuming any value greater than 0 indicates expression
    is_expressed <- exrMat > 0
    
    # Step 2: Calculate the percentage of cells in which each gene is expressed
    percent_expressed <- rowSums(is_expressed) / ncol(exrMat) * 100
    
    # Step 3: Identify genes expressed in at least 20% of cells
    keep_genes <-
      names(percent_expressed)[percent_expressed >= 10]
    tmp[[j]] <- keep_genes
  }
   tmp_2 <-
   as.data.frame(sort(Reduce(intersect,list( tmp$sscrofa, tmp$hsapiens)), decreasing = F) )
   colnames(tmp_2) <- paste('conservative_genes', i)
  gene_list[[i]] <- tmp_2
}

openxlsx::write.xlsx(gene_list, file = 'conservative_genes_expressed_min10pct_per_phenotype_Cor_9010_cell_categorization.xlsx', overwrite = T)

table(my_data$integrated_snn_res.0.2, my_data$species)
# Make balanced dataset

my_data_balanced <- my_data[,colnames(my_data) %in% c(
  sample(names(my_data$species_2[my_data$species_2 == 'human' & my_data$shared_clusters == 'Terminal']), size = 200, replace = F),
  sample(names(my_data$species_2[my_data$species_2 == 'human' & my_data$shared_clusters == 'Contractile']), size = 200, replace = F),
  sample(names(my_data$species_2[my_data$species_2 == 'human' & my_data$shared_clusters == 'Fibroblasts']), size = 200, replace = F),
  sample(names(my_data$species_2[my_data$species_2 == 'human' & my_data$shared_clusters == 'Transitional']), size = 200, replace = F),
  sample(names(my_data$species_2[my_data$species_2 == 'human' & my_data$shared_clusters == 'IFN_SMC']), size = 10, replace = F),
  
  sample(names(my_data$species_2[my_data$species_2 == 'pig' & my_data$shared_clusters == 'Terminal']), size = 200, replace = F) ,
  sample(names(my_data$species_2[my_data$species_2 == 'pig' & my_data$shared_clusters == 'Contractile']), size = 200, replace = F) ,
  sample(names(my_data$species_2[my_data$species_2 == 'pig' & my_data$shared_clusters == 'Fibroblasts']), size = 200, replace = F) ,
  sample(names(my_data$species_2[my_data$species_2 == 'pig' & my_data$shared_clusters == 'Transitional']), size = 200, replace = F) ,
  sample(names(my_data$species_2[my_data$species_2 == 'pig' & my_data$shared_clusters == 'IFN_SMC']), size = 10, replace = F) 
  )]

table(my_data_balanced$species, my_data_balanced$shared_clusters)

DimPlot_scCustom(my_data_balanced, group.by = 'shared_clusters', split.by = 'species')

Idents(my_data_balanced) <- 'integrated_snn_res.0.2'
DefaultAssay(my_data_balanced) <- 'RNA_genes'

tmp <- FindAllMarkers(my_data_balanced, only.pos = T, min.pct = 0.3)
tmp <-  tmp[tmp$p_val_adj < 0.05,]
FeaturePlot_scCustom(my_data_balanced, features = c('THBS1', 'SERPINE2'), split.by = 'species')


#### DEseq2
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98) == TRUE, final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)

CreatePseudoBulkData <- function(raw.data,normalized.data,sample.id,fun="sum")
{
    # fun can be "sum" or "mean"
    library(muscat)
    library(SingleCellExperiment)
    
    if (fun=="sum")
    {
        sce <- SingleCellExperiment(assays=list(counts=raw.data),colData=DataFrame(sample_id=sample.id))
    } else if (fun=="mean") {
        sce <- SingleCellExperiment(assays=list(counts=normalized.data),colData=DataFrame(sample_id=sample.id))
    }
    pb <- aggregateData(sce, assay = "counts", fun=fun, by=c("sample_id"))
    return(pb)
}
    
# Make counts slot with gene names
my_counts <-
  GetAssayData(object = my_data_balanced,
               assay = 'RNA',
               slot = 'counts')
rownames(my_counts) <-
  gsub(pattern = '-0',
       replacement = '',
       x = rownames(my_counts))
rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

my_counts <-
  my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

rownames(final_species_genes) <-
  final_species_genes$`Pig Ensembl ID`
df <-
  data.frame(
    pig_ens_data = rownames(my_counts),
    transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'],
    transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID']
  )

rownames(df) <- df$pig_ens_data
rownames(my_counts) <-
  df[rownames(my_counts), 'transl_human_gene']

# Run DESeq2 across species within each shared SMC cluster (pairwise)
deseq2_res <- list()
for(i in unique(my_data_balanced$shared_clusters)) {
  my_data_balanced_subset <-
    my_data_balanced[, my_data_balanced$shared_clusters == i]
  pb <-
    CreatePseudoBulkData(
      raw.data = my_counts[, colnames(my_counts) %in% colnames(my_data_balanced_subset)],
      sample.id = my_data_balanced_subset$orig.ident,
      fun = 'sum'
    )
  
  bca <-
    as.data.frame.matrix(table(my_data_balanced_subset$orig.ident, my_data_balanced_subset$species_2))
  group <- apply(bca, 1, function(x)
    colnames(bca)[which.max(x)])
  
  
  #Make the model and contrast for statistical testing
  df_con <- DataFrame(condition = factor(group))
  pb_counts <- assay(pb)
  mode(pb_counts) <- "integer"
  dge <- DESeqDataSetFromMatrix(pb_counts,
                                colData = df_con,
                                design = ~ 0 + condition)
  dds <- DESeq(dge, betaPrior = FALSE)
  
  hp_res <-
    results(
      dds,
      contrast = c("condition", "human", "pig"),
      independentFiltering = TRUE,
      alpha = 0.05,
      pAdjustMethod = "BH",
      parallel = FALSE
    )
  hp_res <-
    lfcShrink(
      dds,
      contrast = c("condition", "human", "pig"),
      res = hp_res,
      type = 'ashr'
    )
  hp_res <-
    as.data.frame(hp_res@listData, row.names = hp_res@rownames)
  
  
  colnames(hp_res) <- paste('HP', colnames(hp_res), sep = '_')
  
  
  res_all_org <- cbind(hp_res)
  res_all_org$gene <- rownames(res_all_org)
  res_all_org$exp_PIG <- rowMeans2(expm1(subset(my_data_balanced_subset, species_2 == 'pig')[["RNA_genes"]]@data[rownames(res_all_org), ]))
  res_all_org$exp_HUMAN <- rowMeans2(expm1(subset(my_data_balanced_subset, species_2 == 'human')[["RNA_genes"]]@data[rownames(res_all_org), ]))
  
  res_all_org$pct_PIG <- rowSums(subset(my_data_balanced_subset, species_2 == 'pig')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) / ncol(subset(my_data_balanced_subset, species_2 == 'pig')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) * 100
  res_all_org$pct_HUMAN <- rowSums(subset(my_data_balanced_subset, species_2 == 'human')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) / ncol(subset(my_data_balanced_subset, species_2 == 'human')[["RNA_genes"]]@data[rownames(res_all_org), ] > 0) * 100

  deseq2_res[[i]] <- res_all_org
}
saveRDS(deseq2_res, file = 'deseq2.pairwise.shared.SMC.clusters_Coronary_9010_cell_categorization.rds')
write.xlsx(deseq2_res, 'deseq2.pairwise.shared.SMC.clusters_Coronary_9010_cell_categorization.xlsx', overwrite = T)
saveRDS(my_data, 'shared_Cor_HE_RPCA_cell_categorization.rds')
```


```{r}
## Rank comparison
# avex <- AverageExpression(my_data, assays = 'RNA_genes', group.by = 'orig.ident', slot = 'data')
# avex2 <- apply(avex$RNA_genes, 2, rank_my)
# bca <- as.data.frame.matrix(table(my_data$orig.ident, my_data$species))
# nl<- apply(bca, 1, function(x) colnames(bca)[which.max(x)])
# avex4 <- avex2[!(rowMax(avex2) < 0.33),]


minMax <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}
rank_my <- function(x) {
  rank(x) / length(x)
}


rank_test <- list()
for(i in unique(my_data_balanced$shared_clusters)) {
  my_data_subset <- my_data_balanced[,my_data_balanced$shared_clusters == i]
  avex <-
    AverageExpression(my_data_subset,
                      assays = 'RNA_genes',
                      group.by = 'orig.ident',
                      slot = 'data')
  
  
  #rank data using custom function
  avex2 <- apply(avex$RNA_genes, 2, rank_my)
  avex4 <- avex2[!(rowMax(avex2) < 0.33),]
  
  bca <-
    as.data.frame.matrix(table(my_data_subset$orig.ident, my_data_subset$species))
  nl <- apply(bca, 1, function(x)
    colnames(bca)[which.max(x)])
  
  kr.test <- apply(avex4, 1, function(x) {
  res <- kruskal.test(x ~ nl)
  return(c(res$statistic, res$p.value))
})
  kr.test <- data.frame(t(kr.test))
  colnames(kr.test) <- c("kr.chisq", "kr.p.value")
  kr.test$kr.p.fdr <- p.adjust(kr.test$kr.p.value, method = "fdr")
  table(kr.test$kr.p.fdr < 0.05)
  kr.test$gene <- rownames(kr.test)
  
  # ## Wilcox test
  # wilc.test.res <- apply(avex4, 1, function(x) {
  #   res <-
  #     pairwise.wilcox.test(x, g = nl, p.adjust.method = "bonferroni")
  #   return(c(res$p.value[1, 1], res$p.value[2, 1], res$p.value[2, 2]))
  # })
  # wilc.test.res <- data.frame(t(wilc.test.res))
  # colnames(wilc.test.res) <-
  #   c("wilcox.p.val.hm", "wilcox.p.val.hp", "wilcox.p.val.pm")
  # wilc.test.res$wilcox.p.val.hm.fdr <-
  #   p.adjust(wilc.test.res$wilcox.p.val.hm, method = "fdr")
  # wilc.test.res$wilcox.p.val.hp.fdr <-
  #   p.adjust(wilc.test.res$wilcox.p.val.hp, method = "fdr")
  # wilc.test.res$wilcox.p.val.pm.fdr <-
  #   p.adjust(wilc.test.res$wilcox.p.val.pm, method = "fdr")
  # wilc.test.res$gene <- rownames(wilc.test.res)
  
  ### t test
  t.test_res <- apply(avex4, 1, function(x) {
    res <- pairwise.t.test(x, g = nl, p.adjust.method = "bonferroni")
    return(c(res$p.value[1, 1], res$p.value[2, 1], res$p.value[2, 2]))
  })
  t.test_res <- data.frame(t(t.test_res))
  colnames(t.test_res) <-
    c("t.test.p.val.hm", "t.test.p.val.hp", "t.test.p.val.pm")
  t.test_res$t.test.p.val.hm.fdr <-
    p.adjust(t.test_res$t.test.p.val.hm, method = "fdr")
  t.test_res$t.test.p.val.hp.fdr <-
    p.adjust(t.test_res$t.test.p.val.hp, method = "fdr")
  t.test_res$t.test.p.val.pm.fdr <-
    p.adjust(t.test_res$t.test.p.val.pm, method = "fdr")
  
  table(t.test_res$t.test.p.val.hm.fdr > 0.05)
  table(t.test_res$t.test.p.val.hp.fdr > 0.05)
  table(t.test_res$t.test.p.val.pm.fdr > 0.05)
  
  t.test_res$gene <- rownames(t.test_res)
  
  avex5 <- merge.data.frame(x =  t.test_res, y= c(  kr.test), by = 'gene', all = T)
  
  avex5 <- column_to_rownames(avex5, var = 'gene')
  avex5$exp_PIG <- rowMeans2(expm1(subset(my_data_subset, species == 'pig')[["RNA_genes"]]@data[rownames(avex5), ]))
  avex5$exp_HUMAN <- rowMeans2(expm1(subset(my_data_subset, species == 'human')[["RNA_genes"]]@data[rownames(avex5), ]))
  avex5$exp_MOUSE <- rowMeans2(expm1(subset(my_data_subset, species == 'mouse')[["RNA_genes"]]@data[rownames(avex5), ]))
  
  avex5$pct_PIG <- rowSums(subset(my_data_subset, species == 'pig')[["RNA_genes"]]@data[rownames(avex5), ] > 0) / ncol(subset(my_data_subset, species == 'pig')[["RNA_genes"]]@data[rownames(avex5), ] > 0) * 100
  avex5$pct_HUMAN <- rowSums(subset(my_data_subset, species == 'human')[["RNA_genes"]]@data[rownames(avex5), ] > 0) / ncol(subset(my_data_subset, species == 'human')[["RNA_genes"]]@data[rownames(avex5), ] > 0) * 100
  avex5$pct_MOUSE <- rowSums(subset(my_data_subset, species == 'mouse')[["RNA_genes"]]@data[rownames(avex5), ] > 0) / ncol(subset(my_data_subset, species == 'mouse')[["RNA_genes"]]@data[rownames(avex5), ] > 0) * 100
  
  avex5$gene <- rownames(avex5)
  rank_test[[i]] <- avex5
}

write.xlsx(rank_test, 'perc.rank.pairwise.shared.SMC.clusters_categorization.xlsx', overwrite = T)

```

```{r}
deseq2_res <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/deseq2.pairwise.shared.SMC.clusters_Elastic_cell_categorization.rds")


# Correlation of ranked genes
avex <- AverageExpression(my_data, assays = 'RNA_genes', group.by = 'orig.ident', slot = 'data')

# Filter out house-keeping genes
HKG <- intersect(na.omit(deseq2_res$Contractile[deseq2_res$Contractile$HP_padj > 0.05 & 
             deseq2_res$Contractile$HM_padj > 0.05 &
           deseq2_res$Contractile$PM_padj > 0.05, 'gene' ]),  
           
           c( 
           
          na.omit( deseq2_res$Transitional[deseq2_res$Transitional$HP_padj > 0.05 & 
             deseq2_res$Transitional$HM_padj > 0.05 &
           deseq2_res$Transitional$PM_padj > 0.05, 'gene']),
           
          na.omit( deseq2_res$Terminal[deseq2_res$Terminal$HP_padj > 0.05 & 
             deseq2_res$Terminal$HM_padj > 0.05 &
           deseq2_res$Terminal$PM_padj > 0.05, 'gene']) ,
          
          na.omit( deseq2_res$Proliferating[deseq2_res$Terminal$HP_padj > 0.05 & 
             deseq2_res$Proliferating$HM_padj > 0.05 &
           deseq2_res$Proliferating$PM_padj > 0.05, 'gene']) ))
           





#### loop across clusters
correl_rank <- list()
for(i in unique(my_data_balanced$shared_clusters)) {
  my_data_subset <- my_data_balanced[,my_data_balanced$shared_clusters == i]
  avex <-
    AverageExpression(my_data_subset,
                      assays = 'RNA_genes',
                      group.by = 'orig.ident',
                      slot = 'data')
  
  avex <- avex[["RNA_genes"]][! rownames(avex[["RNA_genes"]]) %in% HKG, ]
  #rank data using custom function
  avex2 <- apply(avex, 2, rank_my)
  avex2 <- avex2[!(rowMax(avex2) < 0.33),]
  
  bca <-
    as.data.frame.matrix(table(my_data_subset$orig.ident, my_data_subset$species))
  nl <- apply(bca, 1, function(x)
    colnames(bca)[which.max(x)])
  
  avex_avg <- data.frame(
    human = rowMeans(avex2[, nl == "human"]),
    mouse = rowMeans(avex2[, nl == "mouse"]),
    pig = rowMeans(avex2[, nl == "pig"])
  )

  # Convert the matrix to a long format data frame

  mat_long <- as.data.frame(cor(avex_avg)) %>%
    tibble::rownames_to_column(var = "Species1") %>%
    pivot_longer(cols = -Species1,
                 names_to = "Species2",
                 values_to = "Value")
  mat_long$IsDiagonal <- mat_long$Species1 == mat_long$Species2

  p <- ggplot(mat_long, aes(x = Species1, y = Species2, fill = Value)) +
  geom_tile(color = "white", aes(alpha = IsDiagonal)) +  # Control tile appearance
     scale_alpha_manual(values = c(1, 0), guide = "none") +  # Make diagonal tiles fully transparent and remove legend
  geom_text(data = subset(mat_long, !IsDiagonal),  # Exclude diagonal from text plotting
            aes(label = sprintf("%.2f", Value)), color = "black", size = 4) +
  scale_fill_gradientn(colors = colorRampPalette(rev(brewer.pal(9, "PuBuGn")))(100))  +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1)) +
  labs(title = paste('Shared SMC cluster', i), fill = "Correlation", x = NULL, y = NULL) +
     geom_text(aes(label = sprintf("%.2f", Value)), color = "black", size = 4)   # Text labels
  
   correl_rank[[i]] <- p
}

p <- plot_grid(plotlist = correl_rank, nrow = 2, ncol = 2, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Helvetica', label_x = -0.1)
print(p)
saveRDS(p, 'correlation_shared_SMC_phenotypes_LIST_categorization.rds')
p <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/correlation_shared_SMC_phenotypes_LIST.rds")

ggsave(filename = 'correlation_shared_SMC_phenotypes_categorization.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 1.5)
```


# To find out markers for species-specific populations, I will go back to individual datasets. Integrated datasets do not have all genes, but only those that are expressed in all 3 species, while it can be important to have genes expressed solely in some species

```{r}
Idents(my_data) <- 'specificity_category'

tmp <- FindAllMarkers(object = my_data, assay = 'RNA_genes', only.pos = T)
tmp <- tmp[tmp$p_val_adj < 0.05,]

View(tmp[tmp$cluster == 'mmusculus-specific', c(2,3,4,6,7)]) ## IBSP, PTH1R, MIA, COMP, BMP2

table( my_data$smc_phenotypes ,my_data$specificity_category) 
```

# UMAPs of the individual datasets with highlighted species-specific and shared clusters (based on barcodes)


```{r}
mouse <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/mouse/smc_plaque_Alencar_seurat.rds")
human <-  readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/human/athero_comb_smc_seurat.rds")
pig <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/Pigs/Pigs_regression/Carlos_new_analysis/SC.Analysis.V4.ThirdPass.RNA.New/SMC/bengal/pig_plaque_seurat.rds")
```

# Use HE scVI integareted datasets 

```{r}
dim(mouse) # 2412 cells

mouse$specificity <- rep(NA, length(mouse$orig.ident))
human$specificity <- rep(NA, length(human$orig.ident))
pig$specificity <- rep(NA, length(pig$orig.ident))

# all_datasets_el$many_higher_expr_scVI$specificity_category <- my_data$consensus[colnames(all_datasets_el$many_higher_expr_scVI)]
# all_datasets_el$many_higher_expr_scVI$specificity_category <- factor(all_datasets_el$many_higher_expr_scVI$specificity_category)

# Transfer **shared** cluster to all individual species datasets
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Shared']) %>% na.omit()
    my_vec <- gsub('-mmusculus', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    mouse$specificity <-
      ifelse(colnames(mouse) %in% names(my_vec), 'Shared', mouse$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Shared']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Shared', human$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Shared']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Shared', pig$specificity)
        
# **hsapiens-sscrofa-specific**
    
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Human_pig_specific']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Human_pig_specific', human$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Human_pig_specific']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Human_pig_specific', pig$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Human_pig_specific']) %>% na.omit()
    my_vec <- gsub('-mmusculus', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    mouse$specificity <-
      ifelse(colnames(mouse) %in% names(my_vec), 'Human_pig_specific', mouse$specificity)
    
    
# mmusculus-specific
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Mouse_specific']) %>% na.omit()
    my_vec <- gsub('-mmusculus', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    mouse$specificity <-
      ifelse(colnames(mouse) %in% names(my_vec), 'Mouse_specific', mouse$specificity) # Mouse contractile cells contribute to hsapiens-sscrofa-specific
    
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Mouse_specific']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Mouse_specific', human$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Mouse_specific']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Mouse_specific', pig$specificity)
    
    
    
    
# hsapiens-specific
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Human_specific']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Human_specific', human$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Human_specific']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Human_specific', pig$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Human_specific']) %>% na.omit()
    my_vec <- gsub('-mmusculus', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    mouse$specificity <-
      ifelse(colnames(mouse) %in% names(my_vec), 'Human_specific', mouse$specificity) 
    
# **Mouse_human_specific**
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Mouse_human_specific']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Mouse_human_specific', human$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Mouse_human_specific']) %>% na.omit()
    my_vec <- gsub('-mmusculus', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    mouse$specificity <-
      ifelse(colnames(mouse) %in% names(my_vec), 'Mouse_human_specific', mouse$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'Mouse_human_specific']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Mouse_human_specific', pig$specificity)
    
    
# Non significant
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'NS']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Small_contribution', human$specificity)

my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'NS']) %>% na.omit()
    my_vec <- gsub('-mmusculus', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    mouse$specificity <-
      ifelse(colnames(mouse) %in% names(my_vec), 'Small_contribution', mouse$specificity)
    
my_vec <-
      names(all_datasets_el$many_higher_expr_scVI$specificity_category[all_datasets_el$many_higher_expr_scVI$specificity_category == 'NS']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Small_contribution', pig$specificity)
```

```{r}
saveRDS(list(Mouse = mouse$specificity, Human = human$specificity, Pig = pig$specificity), file = 'vector_of_specificity.rds')
```

```{r}
my_data <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/HE_RPCA_cell_categorization.rds")
```


```{r}
pal_species <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_species.rds")

# plots smc_phenotypes + specificIty on individual datasets
# pal_specificity <- met.brewer(name = 'Cross', n = 9, direction = -1)
# pal_specificity <-
#   c(
#     met.brewer(
#       name = 'Cross',
#       n = 9,
#       direction = -1
#     )[c(5, 9)],
#     "#20854EFF",
#     "#00A087FF",
#     met.brewer(
#       name = 'Demuth',
#       n = 8,
#       direction = -1
#     )[c(4, 3, 2, 7)],
#     "#7876B1FF"
#   )

# Example of a nice 9-color palette in R
# Example 9-color palette, slightly brighter
# Combine 3-step purple gradient + 6 NPG-inspired colors
pal_specificity <- c('#C1C8D9', '#8491B4df',  '#4f576c',
  # 3 purples in a gradient
  # "#CBD5E1",  # light gray-blue
  # "#94A3B8",  # medium gray-blue
  # "#64748B",  # darker gray-blue
  pal_npg("nrc")(10)[c(1,2,3,10,5,7,8,9)]
  # 6 colors adapted from an NPG-like palette
 #  "#E64B35",  # vivid red
 #  "#4DBBD5",  # teal
 #  "#00A087",  # greenish-teal
 # "#FFDC91FF", 
 #  "#F39B7F",  # peach/orange
 #  "#7E6148"   # subdued brown
)


# "#E64B35FF" "#4DBBD5FF" "#00A087FF" "#3C5488FF" "#F39B7FFF" "#8491B4FF"
# pal_specificity <- c( "#CBD5E1",  # light gray-blue
#   "#94A3B8",  # medium gray-blue
#   "#64748B", brewer.pal(n = 9, name = 'Paired' )[3:9]) 


names(pal_specificity) <- c('Contractile', 'Transitional', 'Terminal',  'Pericytes_Human_pig_specific', 'DLX5_SMC_Human_specific','Terminal_Mouse_specific', 'Proliferating', 'Contractile_Human_pig_specific',   'Too_small_cluster')

pal_celltypes_pig <- c("#E18727FF", "#0072B5FF",  "#BC3C29FF", "#20854EFF", "#7876B1FF" ,"#FFDC91FF", "#EE4C97FF")

pal_celltypes_human <- c("#0072B5FF", "#BC3C29FF" ,"#E18727FF", "#80796BFF", "#20854EFF",  "#374E55FF")

pal_celltypes_mouse <- c("#E18727FF","#BC3C29FF",  "#0072B5FF", "#79AF97FF")


mouse$smc_phenotypes_specificity <- paste(mouse$smc_phenotypes, mouse$specificity, sep = '_') # Mouse Contractile and Contractile_Rgs5 contribure to pig-human contractile, but too little in comparison with human and pig

human$smc_phenotypes_specificity <- paste(human$smc_phenotypes, human$specificity, sep = '_')
pig$smc_phenotypes_specificity <- paste(pig$smc_phenotypes, pig$specificity, sep = '_')

my_vec <- c('Contractile_Human_pig_specific', 'DLX5_SMC_Human_specific', 'Pericytes_Human_pig_specific', 'Terminal_Mouse_specific')
mouse$for_heatmap <- ifelse(mouse$smc_phenotypes_specificity %in% my_vec, mouse$smc_phenotypes_specificity, 'Shared')
pig$for_heatmap <- ifelse(pig$smc_phenotypes_specificity %in% my_vec, pig$smc_phenotypes_specificity, 'Shared')
human$for_heatmap <- ifelse(human$smc_phenotypes_specificity %in% my_vec, human$smc_phenotypes_specificity, 'Shared')

## Divide 'Shared' into sub-phenotypes
head(colnames(mouse))
head(colnames(human))
head(colnames(pig))

#Mouse
xxx <- my_data[, my_data$species == 'mouse']$shared_clusters 
names(xxx) <-  substr(names(xxx), start = 1, stop = 16)
xxx <- xxx[substr(colnames(mouse), start = 1, stop = 16)]
names(xxx) <- colnames(mouse)
mouse$shared_clusters <- xxx
mouse$shared_clusters <- as.character(mouse$shared_clusters)
mouse$shared_clusters[is.na(mouse$shared_clusters)] <- 'specific'
table(mouse$shared_clusters, mouse$smc_phenotypes)

mouse$shared_clusters <- ifelse(mouse$shared_clusters == 'specific', mouse$specificity, no = mouse$shared_clusters)
mouse$shared_clusters[mouse$shared_clusters == 'Mouse_specific'] <- 'Terminal_Mouse_specific'
mouse$shared_clusters[mouse$shared_clusters == 'Human_pig_specific'] <- 'Contractile_Human_pig_specific'

# Pig
xxx <- my_data[, my_data$species == 'pig']$shared_clusters 
names(xxx) <-  substr(names(xxx), start = 1, stop = 16)
xxx <- xxx[substr(colnames(pig), start = 1, stop = 16)]
names(xxx) <- colnames(pig)
pig$shared_clusters <- xxx
pig$shared_clusters <- as.character(pig$shared_clusters)
pig$shared_clusters[is.na(pig$shared_clusters)] <- 'specific'
table(pig$shared_clusters, pig$smc_phenotypes)

pig$shared_clusters <- ifelse(pig$shared_clusters == 'specific', pig$specificity, no = pig$shared_clusters)
pig$shared_clusters[pig$smc_phenotypes == 'Pericytes'] <- 'Pericytes_Human_pig_specific' 
pig$shared_clusters[pig$shared_clusters == 'Human_pig_specific'] <- 'Contractile_Human_pig_specific' 
pig$shared_clusters[pig$shared_clusters == 'Mouse_specific'] <- 'Terminal_Mouse_specific' 
table(pig$shared_clusters)
table(pig$shared_clusters, pig$smc_phenotypes)

# Human
xxx <- my_data[, my_data$species == 'human']$shared_clusters 
names(xxx) <-  substr(names(xxx), start = 12, stop = 27)
xxx <- xxx[substr(colnames(human), start = 12, stop = 27)]
names(xxx) <- colnames(human)
human$shared_clusters <- xxx
human$shared_clusters <- as.character(human$shared_clusters)
human$shared_clusters[is.na(human$shared_clusters)] <- 'specific'
table(human$shared_clusters, human$smc_phenotypes)

human$shared_clusters <- ifelse(human$shared_clusters == 'specific', human$specificity, no = human$shared_clusters)
human$shared_clusters[human$smc_phenotypes == 'Pericytes'] <- 'Pericytes_Human_pig_specific' 
human$shared_clusters[human$shared_clusters == 'Human_pig_specific'] <- 'Contractile_Human_pig_specific' 
human$shared_clusters[human$shared_clusters == 'Human_specific'] <- 'DLX5_SMC_Human_specific'
human$shared_clusters[is.na(human$shared_clusters)==TRUE] <- 'Too_small_cluster'
table(human$shared_clusters)
table(human$shared_clusters, human$smc_phenotypes)


##### Mouse
spec_mouse = DimPlot_scCustom(
  mouse,
    reduction = 'umap',
    group.by = c('shared_clusters'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_specificity, repel = T, label.size = 2.7
  ) + 
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
      scale_color_manual(values = pal_specificity) + 
  theme(aspect.ratio = 1,
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  smc_mouse = DimPlot_scCustom(mouse,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_mouse, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_mouse) +
    theme(aspect.ratio = 1,
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )

##### Pig
spec_pig = DimPlot_scCustom(
  pig,
    reduction = 'umap',
    group.by = c('shared_clusters'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_specificity, repel = T, label.size = 2.7
  ) + 
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
      scale_color_manual(values = pal_specificity) + 
  theme(aspect.ratio = 1,
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  smc_pig = DimPlot_scCustom(pig,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_pig, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_pig) +
    theme(aspect.ratio = 1,
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
###### Human
spec_human = DimPlot_scCustom(
  human,
    reduction = 'umap',
    group.by = c('shared_clusters'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_specificity, repel = T, label.size = 2.7
  ) + 
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
      scale_color_manual(values = pal_specificity) + 
  theme(aspect.ratio = 1,
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  smc_human = DimPlot_scCustom(human,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_human, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_human) +
    theme(aspect.ratio = 1,
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
#### Combine

fig = ggdraw() +
  draw_plot(spec_mouse, x = 0.0, y = 0, width = .33, height = .48) +
  draw_plot(smc_mouse , x = 0.0, y = 0.5, width = .33, height = .48 )  +
  draw_plot_label(label = 'Mouse SMCs', size = 12,
                  x = c(-0.01), y = c(1.005), family = 'Helvetica')  +
  draw_plot(spec_pig, x = 0.34, y = 0, width = .34, height = .48) +
  draw_plot(smc_pig , x = 0.34, y = 0.5, width = .34 , height =  .48)  +
  draw_plot_label(label = 'Pig SMCs', size = 12,
                  x = c(0.4), y = c(1.005), family = 'Helvetica')  +
  draw_plot(spec_human, x = 0.65, y = 0, width = .34, height = .48) +
  draw_plot(smc_human , x = 0.65, y = .5 , width = .34 , height = .48 )  +
  draw_plot_label(label = 'Human SMCs', size = 12,
                  x = c(0.65), y = c(1.005), family = 'Helvetica')  

fig
saveRDS(fig, file = 'UMAPS_individual_ds_species_specific_clusters_category.rds')


ggsave(filename = 'UMAPS_individual_ds_species_specific_clusters_category.pdf',  plot = fig, device = 'pdf', units = 'cm', dpi = 400, width = 8, height = 4, scale = 3)  

DimPlot(mouse, group.by = 'specificity') + DimPlot(human, group.by = 'specificity') + DimPlot(pig, group.by = 'specificity')
my_vec %in% colnames(mouse) %>% table()
```

```{r}
# Marker identification for cross-species SMC phenotypes mapping
Idents(pig) <- 'shared_clusters'
tmp_p <- FindAllMarkers(pig, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_p, file = 'pig_elastic_unified_markers.xlsx')

Idents(human) <- 'shared_clusters'
tmp_h <- FindAllMarkers(human, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_h, file = 'human_elastic_unified_markers.xlsx')

Idents(mouse) <- 'shared_clusters'
tmp_m <- FindAllMarkers(mouse, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_m, file = 'mouse_elastic_unified_markers.xlsx')

Idents(pig) <- 'smc_phenotypes'
tmp_p <- FindAllMarkers(pig, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_p, file = 'pig_elastic_ind_markers.xlsx')

Idents(human) <- 'smc_phenotypes'
tmp_h <- FindAllMarkers(human, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_h, file = 'human_elastic_ind_markers.xlsx')

Idents(mouse) <- 'smc_phenotypes'
tmp_m <- FindAllMarkers(mouse, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_m, file = 'mouse_elastic_ind_markers.xlsx')



```

```{r}
FeaturePlot_scCustom(pig, features = c('SERPINF1'))
FeaturePlot_scCustom(human, features = c('SERPINF1'))
FeaturePlot_scCustom(mouse, features = c('Serpinf1'))

FeaturePlot_scCustom(pig, features = c('TNFRSF11B'))
FeaturePlot_scCustom(human, features = c('TNFRSF11B'))
FeaturePlot_scCustom(mouse, features = c('Tnfrsf11b'))
```

# Marker identification for species-specific SMC phenotypes
To do for: mouse terminal, human DLX5, human-pig pericytes and contractile SMCs. Visualize expression to check specificity of markers for other datasets

# Mouse-specific terminal SMCs. Going back to mouse individual dataset. 

```{r}
Idents(mouse) <- 'specificity'

term_markers <- multiMarkerStats(
    wilcox=findMarkers(as.SingleCellExperiment(mouse), test="wilcox", direction="up",group = mouse$specificity, add.summary = TRUE),
    binom=findMarkers(as.SingleCellExperiment(mouse), test="binom", direction="up",group = mouse$specificity, add.summary = TRUE), sorted = TRUE
)

term_markers <- term_markers@listData[["Mouse-specific"]] %>% as.matrix() %>% as.data.frame()
term_markers <- term_markers[term_markers$wilcox.summary.AUC > 0.7,]
term_markers <- term_markers[order(term_markers$binom.summary.logFC, decreasing = T),] 
term_markers$gene <- rownames(term_markers)


# tmp_mouse <- FindAllMarkers(object = mouse, assay = 'RNA', test.use = 'roc', only.pos = T)
# tmp_mouse <- tmp_mouse[tmp_mouse$cluster == 'Mouse-specific', ]
Idents(mouse) <- 'specificity'
tmp_mouse <- FindAllMarkers(object = mouse, assay = 'RNA', test.use = 'MAST', only.pos = T)
tmp_mouse <- tmp_mouse[tmp_mouse$cluster == 'Mouse-specific', ]
```

# Human DLX5_SMCs. Majority of the cluster cells are placed specifically as unique human cells, so I will find markers for DLX5_SMCs cluster

```{r}
table(human$smc_phenotypes, human$specificity)

Idents(human) <- 'smc_phenotypes'

dlx5_markers <- multiMarkerStats(
    wilcox=findMarkers(as.SingleCellExperiment(human), test="wilcox", direction="up",group = human$smc_phenotypes, add.summary = TRUE),
    binom=findMarkers(as.SingleCellExperiment(human), test="binom", direction="up",group = human$smc_phenotypes, add.summary = TRUE), sorted = TRUE
)

dlx5_markers <- dlx5_markers@listData[["DLX5_SMC"]] %>% as.matrix() %>% as.data.frame()
dlx5_markers <- dlx5_markers[dlx5_markers$wilcox.summary.AUC > 0.7,]
dlx5_markers <- dlx5_markers[order(dlx5_markers$binom.summary.logFC, decreasing = T),] 
dlx5_markers$gene <- rownames(dlx5_markers)


# dlx5_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'roc', only.pos = T)
# dlx5_markers <- dlx5_markers[dlx5_markers$cluster == 'DLX5_SMC',]

dlx5_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'MAST', only.pos = T)
dlx5_markers <- dlx5_markers[dlx5_markers$cluster == 'DLX5_SMC',]
```

# Human-pig specific pericytes. Find common markers

```{r}
Idents(human) <- 'smc_phenotypes'

PC_markers <- multiMarkerStats(
    wilcox=findMarkers(as.SingleCellExperiment(human), test="wilcox", direction="up",group = human$smc_phenotypes, add.summary = TRUE),
    binom=findMarkers(as.SingleCellExperiment(human), test="binom", direction="up",group = human$smc_phenotypes, add.summary = TRUE), sorted = TRUE
)

PC_markers <- PC_markers@listData[["Pericytes"]] %>% as.matrix() %>% as.data.frame()
PC_markers <- PC_markers[PC_markers$wilcox.summary.AUC > 0.7,]
PC_markers <- PC_markers[order(PC_markers$binom.summary.logFC, decreasing = T),] 
colnames(PC_markers) <- paste('Human', colnames(PC_markers), sep = '_')
PC_markers$gene <- rownames(PC_markers)

# PC_markers <- FindAllMarkers(object = human, assay = 'RNA',  only.pos = T)
# PC_markers <- PC_markers[PC_markers$cluster == 'Pericytes',]
# colnames(PC_markers)[1:6] <- paste('Human', colnames(PC_markers)[1:6], sep = '_')

# PC_markers <- FindAllMarkers(object = human, assay = 'RNA',  only.pos = T, test.use = 'MAST')
# PC_markers <- PC_markers[PC_markers$cluster == 'Pericytes',]
# colnames(PC_markers)[1:6] <- paste('Human', colnames(PC_markers)[1:6], sep = '_')


Idents(pig) <- 'smc_phenotypes'
PC_markers_pig <- multiMarkerStats(
    wilcox=findMarkers(as.SingleCellExperiment(pig), test="wilcox", direction="up",group = pig$smc_phenotypes, add.summary = TRUE),
    binom=findMarkers(as.SingleCellExperiment(pig), test="binom", direction="up",group = pig$smc_phenotypes, add.summary = TRUE), sorted = TRUE
)

PC_markers_pig <- PC_markers_pig@listData[["Pericytes"]] %>% as.matrix() %>% as.data.frame()
PC_markers_pig <- PC_markers_pig[PC_markers_pig$wilcox.summary.AUC > 0.7,]
PC_markers_pig <- PC_markers_pig[order(PC_markers_pig$binom.summary.logFC, decreasing = T),] 
colnames(PC_markers_pig) <- paste('Pig', colnames(PC_markers_pig), sep = '_')
PC_markers_pig$gene <- rownames(PC_markers_pig)

# PC_markers_pig <- FindAllMarkers(object = pig, assay = 'RNA',  only.pos = T)
# PC_markers_pig <- PC_markers_pig[PC_markers_pig$cluster == 'Pericytes',]
# colnames(PC_markers_pig)[1:6] <- paste('Pig', colnames(PC_markers_pig)[1:6], sep = '_')
# PC_markers_pig <- PC_markers_pig[PC_markers_pig$Pig_p_val_adj < 0.05,]

# PC_markers_pig <- FindAllMarkers(object = pig, assay = 'RNA',  only.pos = T, test.use = 'MAST')
# PC_markers_pig <- PC_markers_pig[PC_markers_pig$cluster == 'Pericytes',]
# colnames(PC_markers_pig)[1:6] <- paste('Pig', colnames(PC_markers_pig)[1:6], sep = '_')


PC_markers_HP <- merge.data.frame(x = PC_markers, y = PC_markers_pig, by = 'gene')
```

# Human-pig specific contractile SMCs. Not all contractile SMCs are pig-human specific. Some of them are common between the species. I will split cSMCs and find markers only for human-pig specific ones


```{r}
table(human$smc_phenotypes, human$specificity)
table(pig$smc_phenotypes, pig$specificity)

human$smc_phenotypes_specificity <- paste(human$smc_phenotypes, human$specificity, sep = '_')
pig$smc_phenotypes_specificity <- paste(pig$smc_phenotypes, pig$specificity, sep = '_')

Idents(human) <- 'smc_phenotypes_specificity'
PC_markers <- multiMarkerStats(
    wilcox=findMarkers(as.SingleCellExperiment(human), test="wilcox", direction="up",group = human$smc_phenotypes_specificity, add.summary = TRUE),
    binom=findMarkers(as.SingleCellExperiment(human), test="binom", direction="up",group = human$smc_phenotypes_specificity, add.summary = TRUE), sorted = TRUE
)

PC_markers <- PC_markers@listData[["Contractile_Human-pig-specific"]] %>% as.matrix() %>% as.data.frame()
PC_markers <- PC_markers[PC_markers$wilcox.summary.AUC > 0.7,]
PC_markers <- PC_markers[order(PC_markers$binom.summary.logFC, decreasing = T),] 
colnames(PC_markers) <- paste('Human', colnames(PC_markers), sep = '_')
PC_markers$gene <- rownames(PC_markers)

# PC_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'roc', only.pos = T)
# PC_markers <- PC_markers[PC_markers$cluster == 'Contractile_Human-pig-specific',]
# colnames(PC_markers)[1:6] <- paste('Human', colnames(PC_markers)[1:6], sep = '_')


####### MAST
Idents(human) <- 'smc_phenotypes_specificity'
PC_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'MAST', only.pos = T)
PC_markers <- PC_markers[PC_markers$cluster == 'Contractile_Human-pig-specific',]
colnames(PC_markers)[1:6] <- paste('Human', colnames(PC_markers)[1:6], sep = '_')

###############################
Idents(pig) <- 'smc_phenotypes_specificity'
PC_markers_pig <- multiMarkerStats(
    wilcox=findMarkers(as.SingleCellExperiment(pig), test="wilcox", direction="up",group = pig$smc_phenotypes_specificity, add.summary = TRUE),
    binom=findMarkers(as.SingleCellExperiment(pig), test="binom", direction="up",group = pig$smc_phenotypes_specificity, add.summary = TRUE), sorted = TRUE
)

PC_markers_pig <- PC_markers_pig@listData[["Contractile_Human-pig-specific"]] %>% as.matrix() %>% as.data.frame()
PC_markers_pig <- PC_markers_pig[PC_markers_pig$wilcox.summary.AUC > 0.7,]
PC_markers_pig <- PC_markers_pig[order(PC_markers_pig$binom.summary.logFC, decreasing = T),] 
colnames(PC_markers_pig) <- paste('Pig', colnames(PC_markers_pig), sep = '_')
PC_markers_pig$gene <- rownames(PC_markers_pig)

# PC_markers_pig <- FindAllMarkers(object = pig, assay = 'RNA', test.use = 'roc', only.pos = T)
# PC_markers_pig <- PC_markers_pig[PC_markers_pig$cluster == 'Contractile_Human-pig-specific',]
# colnames(PC_markers_pig)[1:6] <- paste('Pig', colnames(PC_markers_pig)[1:6], sep = '_')


####### MAST
Idents(pig) <- 'smc_phenotypes_specificity'
PC_markers_pig <- FindAllMarkers(object = pig, assay = 'RNA', test.use = 'MAST', only.pos = T)
PC_markers_pig <- PC_markers_pig[PC_markers_pig$cluster == 'Contractile_Human-pig-specific',]
colnames(PC_markers_pig)[1:6] <- paste('Pig', colnames(PC_markers_pig)[1:6], sep = '_')

cSMC_markers_HP <- merge.data.frame(x = PC_markers, y = PC_markers_pig, by = 'gene')
```

```{r}
species_markers <- list('Mouse terminal' = term_markers, 
                        'Human DLX5' = dlx5_markers,
                        'Human-pig pericytes' = PC_markers_HP,
                        'Human-pig cSMCs' = cSMC_markers_HP)
saveRDS(species_markers, file = 'df_markers_species_specific_clusters_scran.rds')
openxlsx::write.xlsx(species_markers, file = "Markers_species_specific_clusters_scran.xlsx",
                     asTable = TRUE, overwrite = TRUE)

# species_markers <- list('Mouse terminal' = tmp_mouse, 
#                         'Human DLX5' = dlx5_markers,
#                         'Human-pig pericytes' = PC_markers_HP,
#                         'Human-pig cSMCs' = cSMC_markers_HP)
# saveRDS(species_markers, file = 'df_markers_species_specific_clusters_MAST.rds')
# openxlsx::write.xlsx(species_markers, file = "Markers_species_specific_clusters_MAST.xlsx",
#                      asTable = TRUE, overwrite = TRUE)
```

# Visualization of markers for species-specific clusters. 

Make one-to-one orthologs, but keep features that are not expressed in other species

```{r}
species_markers <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/df_markers_species_specific_clusters.rds") # or _scran with binom mode
DefaultAssay(mouse) <- 'RNA'
DefaultAssay(pig) <- 'RNA'
DefaultAssay(human) <- 'RNA'

# mouse$specificity <- ifelse(is.na(mouse$specificity) == T, 'Other', mouse$specificity)
# pig$specificity <- ifelse(is.na(pig$specificity) == T, 'Other', pig$specificity)
# human$specificity <- ifelse(is.na(human$specificity) == T, 'Other', human$specificity)
# 
# mouse$specificity <- ifelse(mouse$specificity == 'Shared', 'shared', mouse$specificity)
# pig$specificity <- ifelse(pig$specificity == 'Shared', 'shared', pig$specificity)
# human$specificity <- ifelse(human$specificity == 'Shared', 'shared', human$specificity)


mouse$smc_phenotypes_specificity <- paste(mouse$smc_phenotypes, mouse$specificity, sep = '_')
human$smc_phenotypes_specificity <- paste(human$smc_phenotypes, human$specificity, sep = '_')
pig$smc_phenotypes_specificity <- paste(pig$smc_phenotypes, pig$specificity, sep = '_')

my_vec <- c('Contractile_Human_pig_specific', 'DLX5_SMC_Human_specific', 'Pericytes_Human_pig_specific', 'Terminal_Mouse_specific')
mouse$for_heatmap <- ifelse(mouse$smc_phenotypes_specificity %in% my_vec, mouse$smc_phenotypes_specificity, 'Mouse_rest')
pig$for_heatmap <- ifelse(pig$smc_phenotypes_specificity %in% my_vec, pig$smc_phenotypes_specificity, 'Pig_rest')
human$for_heatmap <- ifelse(human$smc_phenotypes_specificity %in% my_vec, human$smc_phenotypes_specificity, 'Human_rest')
```

```{r}
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98) == TRUE, final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)

final_species_genes[nrow(final_species_genes)+1,'Human_gene_name_v98' ] <- 'CST3'
final_species_genes[nrow(final_species_genes),'Human Ensembl ID' ] <- 'CST3'
final_species_genes[nrow(final_species_genes),'mouse_origin' ] <- 'Cst3'
final_species_genes[nrow(final_species_genes),'pig_origin' ] <- 'CST3'

final_species_genes[nrow(final_species_genes)+1,'Human_gene_name_v98' ] <- 'BGLAP'
final_species_genes[nrow(final_species_genes),'Human Ensembl ID' ]  <- 'BGLAP'
final_species_genes[nrow(final_species_genes),'mouse_origin' ] <- 'Bglap'
final_species_genes[nrow(final_species_genes),'pig_origin' ] <- 'BGLAP'


final_species_genes[nrow(final_species_genes)+1,'Human_gene_name_v98' ] <- 'COL9A3'
final_species_genes[nrow(final_species_genes),'Human Ensembl ID' ] <-  'COL9A3'
final_species_genes[nrow(final_species_genes),'mouse_origin' ] <- 'Col9a3'
final_species_genes[nrow(final_species_genes),'pig_origin' ] <- 'COL9A3'
```

```{r}
DefaultAssay(mouse) <- 'RNA' 
mouse_2 <- mouse

# Counts
counts_all <- GetAssayData(object = mouse_2, slot = "counts", assay = 'RNA')
my_df <- data.frame(old=rownames(mouse_2), ##create a dataframe of correspondence
                    old_2=rownames(mouse_2),
                    original = rownames(mouse_2))
rownames(my_df) <- rownames(mouse_2)

colnames(my_df)[2] <- 'mouse_origin' # Mouse_Gene_Names_v99

table(my_df$mouse_origin %in% final_species_genes$mouse_origin)
final_species_genes$scRnaseq_mouse <- final_species_genes$mouse_origin %in% my_df$original
mouse_unique <-  my_df$original[!my_df$original %in% final_species_genes$mouse_origin]
my_df_2 <- merge(my_df,final_species_genes,by="mouse_origin")  ## merging data about mouse dataset genes and ortholog table, Mouse_Gene_Names_v99
my_df_2$human_fin <- ifelse(is.na(my_df_2$`Human Ensembl ID`) == TRUE, my_df_2$`Human Ensembl ID`, my_df_2$Human_gene_name_v98)
rownames(my_df_2) <- my_df_2$original
my_df_2['Cst3', 'human_fin'] <- 'CST3'
filtered_counts_h <- counts_all[my_df_2$original, ] ## keep only genes that have orthologous
rownames(filtered_counts_h) <- my_df_2$human_fin
mouse_2@assays$RNA@counts <- filtered_counts_h

# Data
counts_all <- GetAssayData(object = mouse_2, slot = "data", assay = 'RNA')
my_df <- data.frame(old=rownames(mouse_2), ##create a dataframe of correspondence
                    old_2=rownames(mouse_2),
                    original = rownames(mouse_2))
rownames(my_df) <- rownames(mouse_2)

colnames(my_df)[2] <- 'mouse_origin' # Mouse_Gene_Names_v99

table(my_df$mouse_origin %in% final_species_genes$mouse_origin)
final_species_genes$scRnaseq_mouse <- final_species_genes$mouse_origin %in% my_df$original
mouse_unique <-  my_df$original[!my_df$original %in% final_species_genes$mouse_origin]
my_df_2 <- merge(my_df,final_species_genes,by="mouse_origin")  ## merging data about mouse dataset genes and ortholog table, Mouse_Gene_Names_v99
my_df_2$human_fin <- ifelse(is.na(my_df_2$`Human Ensembl ID`) == TRUE, my_df_2$`Human Ensembl ID`, my_df_2$Human_gene_name_v98)
rownames(my_df_2) <- my_df_2$original
my_df_2['Cst3', 'human_fin'] <- 'CST3'
filtered_counts_h <- counts_all[my_df_2$original, ] ## keep only genes that have orthologous
rownames(filtered_counts_h) <- my_df_2$human_fin
mouse_2@assays$RNA@data <- filtered_counts_h

```


```{r}
DefaultAssay(pig) <- 'RNA' 
pig_2 <- pig

# Counts
DefaultAssay(pig_2) <- 'RNA' 
counts_all <- GetAssayData(object = pig_2, slot = "counts", assay = 'RNA')
my_df <- data.frame(old=rownames(pig_2), ##create a dataframe of correspondence
                    old_2=rownames(pig_2),
                    original = rownames(pig_2))
rownames(my_df) <- rownames(pig_2)
colnames(my_df)[2] <- 'pig_origin' # Pig gene name
final_species_genes$scRnaseq_pig_2 <- final_species_genes$pig_origin %in% my_df$pig_origin
my_df_2 <- merge(my_df,final_species_genes,by=c('pig_origin'))  ## merging data about mouse dataset genes and ortholog table, Pig gene name
my_df_2$human_fin <- ifelse(is.na(my_df_2$Human_gene_name_v98) == TRUE, my_df_2$`Human Ensembl ID`, my_df_2$Human_gene_name_v98)
rownames(my_df_2) <- my_df_2$original
my_df_2['CST3', 'human_fin'] <- 'CST3'
filtered_counts_h <- counts_all[my_df_2$original, ] ## keep only genes that have orthologous
rownames(filtered_counts_h) <- my_df_2$human_fin
pig_2@assays$RNA@counts <- filtered_counts_h

# Data
DefaultAssay(pig_2) <- 'RNA' 
counts_all <- GetAssayData(object = pig, slot = "data", assay = 'RNA')
my_df <- data.frame(old=rownames(pig_2), ##create a dataframe of correspondence
                    old_2=rownames(pig_2),
                    original = rownames(pig_2))
rownames(my_df) <- rownames(pig_2)
colnames(my_df)[2] <- 'pig_origin' # Pig gene name
final_species_genes$scRnaseq_pig_2 <- final_species_genes$pig_origin %in% my_df$pig_origin
my_df_2 <- merge(my_df,final_species_genes,by=c('pig_origin'))  ## merging data about mouse dataset genes and ortholog table, Pig gene name
my_df_2$human_fin <- ifelse(is.na(my_df_2$Human_gene_name_v98) == TRUE, my_df_2$`Human Ensembl ID`, my_df_2$Human_gene_name_v98)
rownames(my_df_2) <- my_df_2$original
my_df_2['CST3', 'human_fin'] <- 'CST3'
filtered_counts_h <- counts_all[my_df_2$original, ] ## keep only genes that have orthologous
rownames(filtered_counts_h) <- my_df_2$human_fin
pig_2@assays$RNA@data <- filtered_counts_h
```



```{r}
human_2 <-  human[, sample(colnames(human), size = 2500, replace = F)]
pig_2 <-  pig_2[, sample(colnames(pig_2), size = 2500, replace = F)]
merged <- Merge_Seurat_List(list_seurat = list(mouse_2, pig_2, human_2), merge.data = T)
saveRDS(merged, file = 'merged_pig_human_mouse_smcs.rds')


merged <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/merged_pig_human_mouse_smcs.rds")
my_data <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/HE_RPCA_cell_categorization.rds") # reclustered shared clusters


# We need to devide 'rest categories' to shared phenotypes
merged$for_heatmap <- factor(merged$for_heatmap)
xxx <- my_data$shared_clusters 

names(xxx) <-  substr(names(xxx), start = 1, stop = 16)
xxx <- xxx[substr(colnames(merged), start = 2, stop = 17)]
names(xxx) <- colnames(merged)
merged$shared_clusters <- xxx
merged$shared_clusters <- as.character(merged$shared_clusters)
merged$shared_clusters[is.na(merged$shared_clusters)] <- 'specific'
table(merged$shared_clusters, merged$for_heatmap)

merged$for_heatmap <-
  factor(merged$for_heatmap, levels = 
    c(
      "DLX5_SMC_Human_specific",
      "Pericytes_Human_pig_specific",
      'Contractile_Human_pig_specific',
       "Terminal_Mouse_specific",
      'Human_rest',
      'Pig_rest',
      'Mouse_rest'
    
  ))

merged$xxx <- paste(merged$shared_clusters, merged$for_heatmap, sep = '_')

merged$xxx <- ifelse(merged$xxx %in% c('Contractile_Human_rest', 'Contractile_Pig_rest', 'Contractile_Mouse_rest', 'specific_Mouse_rest'), 'Contractile_shared', merged$xxx) # because mouse contractile cells contribute to pig-uman contractile cluster, but too little, technically theu are just shared contractile
merged$xxx <- ifelse(merged$xxx %in% c('Transitional_Human_rest', 'Transitional_Pig_rest', 'Transitional_Mouse_rest'), 'Transitional_shared', merged$xxx )
merged$xxx <- ifelse(merged$xxx %in% c('Terminal_Human_rest', 'Terminal_Pig_rest', 'Terminal_Mouse_rest'), 'Terminal_shared', merged$xxx )
merged$xxx <- ifelse(merged$xxx %in% c('Proliferating_Human_rest', 'Proliferating_Pig_rest', 'Proliferating_Mouse_rest'), 'Proliferating_shared', merged$xxx )
table(merged$xxx)

# names(merged$xxx[merged$xxx == 'Terminal_shared'])
merged$for_heatmap_2 <- merged$for_heatmap
merged$for_heatmap_2 <- as.character(merged$for_heatmap_2)
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Terminal_shared'])] <- 'Terminal_shared'
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Contractile_shared'])] <- 'Contractile_shared'
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Transitional_shared'])] <- 'Transitional_shared'
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Proliferating_shared'])] <- 'Proliferating_shared'

table(merged$for_heatmap_2, merged$smc_phenotypes) # 
table(merged$for_heatmap_2)



merged <- merged[,merged$for_heatmap_2 %in%  c(
      "DLX5_SMC_Human_specific",
      "Pericytes_Human_pig_specific",
      'Contractile_Human_pig_specific',
       "Terminal_Mouse_specific",
      'Contractile_shared',
      'Transitional_shared',
      'Terminal_shared',
      'Proliferating_shared')]

#### Make heatmap
Idents(merged) <- 'for_heatmap_2'


merged$for_heatmap_2 <-
  factor(merged$for_heatmap_2, levels = 
    c(
      "DLX5_SMC_Human_specific",
      "Pericytes_Human_pig_specific",
      'Contractile_Human_pig_specific',
       "Terminal_Mouse_specific",
      'Contractile_shared',
      'Transitional_shared',
      'Terminal_shared',
     'Proliferating_shared' ))


saveRDS(merged, file = 'merged_for_heatmap.rds')

# tmp_mouse_1 <- FindAllMarkers(object = merged, assay = 'RNA', test.use = 'roc',  only.pos = T, max.cells.per.ident = 3000, logfc.threshold = 0.1, min.diff.pct = 0.01, min.pct = 0.1)
merged <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/merged_for_heatmap.rds")
tmp_mouse <- FindAllMarkers(object = merged, assay = 'RNA',  min.pct = 0.3, only.pos = T, max.cells.per.ident = 3000, min.diff.pct = 0.2)

top_markers <- c()
for (i in unique(tmp_mouse$cluster)[c(8,7, 5, 4, 3,2,1,6)]){
  top_markers <- c(top_markers, head((subset(tmp_mouse, subset = tmp_mouse$cluster == i)[order(subset(tmp_mouse, subset = tmp_mouse$cluster == i)$avg_log2FC, decreasing = T),])$gene, 7))
}
top_markers[6] <- 'MYH10'
top_markers[22] <- 'COL2A1'
top_markers[42] <- 'LUM'
top_markers[5] <- 'MSX2'
top_markers[21] <- 'ACTG2'
top_markers[13] <- 'GJA4'
top_markers[14] <- 'RGS5'
top_markers[29] <- 'ACTA2'
top_markers[30] <- 'ITGA8'
top_markers[45] <- 'COL1A2'
# species_markers <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/df_markers_species_specific_clusters_scran.rds")
# 
# top_markers <- c()
# for (i in names(species_markers)[c(2,3,4,1)]){
#   my_col <- colnames(species_markers[[i]])[ grep('Top', colnames(species_markers[[i]]))][1]
#   top_markers <- c(top_markers, species_markers[[i]][order(species_markers[[i]][,my_col], decreasing = F), 'gene'][1:10])
# }

p1 <- DotPlot(
  merged,
  features = toupper(unique(top_markers)),
  assay = 'RNA',
  scale = T,
  group.by = 'for_heatmap_2'
) + 
  # ggtitle('Top markers for SMC phenotypes pig aorta plaque') + 
  # coord_flip() +
  # RotatedAxis() + 
  scale_colour_gradientn(name = 'log2 (count + 1)',colours = rev(brewer.pal(n = 11, name = "RdBu")))  +
  theme_bw() + theme(legend.position = 'bottom') +
  theme(axis.line  = element_blank()) +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ylab('') +
  theme(axis.ticks = element_blank()) 
print(p1)


ggsave(filename = 'dotplot_merged_species-specific_clusters_6.pdf', plot = p1, device = 'pdf', width = 9.5, height = 2.9, units = 'cm', scale = 3)
```

```{r}
# For enrichment analysis
table(merged$for_heatmap_2)
Idents(merged) <- 'for_heatmap_2'
tmp <- FindAllMarkers(merged, assay = 'RNA', only.pos = T)
saveRDS(tmp, 'markers_for_enrichment.rds')


```


```{r}
# terminal
feats <- species_markers$`Mouse terminal`[1:12, 'gene']

FeaturePlot_scCustom(mouse, features = feats) + plot_annotation(title = 'Mouse SMCs') 

FeaturePlot_scCustom(pig, features = str_to_upper(feats) )+ plot_annotation(title = 'Pig SMCs') 

FeaturePlot_scCustom(human, features = str_to_upper(feats)) + plot_annotation(title = 'Human SMCs') 

# DLX5 human specific
feats <- c(species_markers$`Human DLX5`[1:12, 'gene'], 'MECOM',  'ID3')

# Pericytes
feats <- c(species_markers$`Human-pig pericytes`[species_markers$`Human-pig pericytes`$Human_pct.2 < 0.5 & species_markers$`Human-pig pericytes`$Pig_pct.2 < 0.5, 'gene'], 'Gucy1b3', 'Filip1')

# cSMCs
feats <- c(species_markers$`Human-pig cSMCs`[, 'gene'], )

FeaturePlot_scCustom(mouse, features = str_to_title(feats)  ) + plot_annotation(title = 'Mouse SMCs') 

FeaturePlot_scCustom(pig, features = str_to_upper(feats) )+ plot_annotation(title = 'Pig SMCs') 

FeaturePlot_scCustom(human, features = str_to_upper(feats)) + plot_annotation(title = 'Human SMCs') 
```

```{r}
my_data <- all_datasets$many_higher_expr_seuratRPCA
DefaultAssay(my_data) <- 'RNA_genes'

ddr <- c("ABL1", "ABRAXAS1", "ACTB", "ACTL6A", "ACTL6B", "ACTR2", "ACTR5", "ACTR8", "ANKLE1", "AP5S1", "AP5Z1", "APLF", "ARID1A", "ARID1B", "ARID2", "ATM", "ATP23", "ATR", "ATRIP", "AUNIP", "BABAM1", "BABAM2", "BCL7A", "BCL7B", "BCL7C", "BLM", "BRCA1", "BRCA2", "BRCC3", "BRD7", "BRD8", "BRIP1", "BRME1", "C14orf39", "C1QBP", "CDC45", "CDC7", "CDCA5", "CGAS", "CHD4", "CHEK1", "CHEK2", "CIB1", "CREBBP", "CSNK2A1", "CSNK2A2", "CYREN", "DCLRE1A", "DCLRE1B", "DCLRE1C", "DDX1", "DDX11", "DEK", "DMAP1", "DMC1", "DNA2", "DNTT", "DPF1", "DPF2", "DPF3", "DTX3L", "EID3", "EME1", "EME2", "EP400", "EPC1", "EPC2", "ERCC1", "ERCC4", "ERCC5", "ERCC6", "ERCC8", "ESCO2", "EXD2", "EYA1", "EYA3", "FAN1", "FANCB", "FANCD2", "FANCM", "FBH1", "FEN1", "FIGNL1", "FMN2", "FOXM1", "FUS", "GEN1", "GGN", "GINS2", "GINS4", "H2AX", "HDGFL2", "HELB", "HELQ", "HMCES", "HMGA2", "HMGB1", "HMGB2", "HPF1", "HSF1", "HSF2BP", "HTATSF1", "HUS1", "HUS1B", "IFFO1", "ING3", "INIP", "INO80", "INTS3", "KASH5", "KAT5", "KDM1A", "KDM2A", "KDM4D", "KHDC3L", "KLHL15", "KMT5A", "KMT5B", "KMT5C", "LIG3", "LIG4", "MAD2L2", "MAGEF1", "MAJIN", "MARF1", "MBTD1", "MCM2", "MCM3", "MCM4", "MCM5", "MCM6", "MCM7", "MCM8", "MCM9", "MCMDC2", "MEAF6", "MEIOB", "MEIOC", "MGMT", "MIR221", "MLH1", "MMS22L", "MORF4L1", "MORF4L2", "MPND", "MRE11", "MRGBP", "MRNIP", "MSH2", "MTA1", "MUS81", "NABP1", "NABP2", "NBN", "NHEJ1", "NIPBL", "NSD2", "NSMCE1", "NSMCE2", "NSMCE3", "NSMCE4A", "NUCKS1", "NUDT16L1", "OGG1", "OOEP", "OTUB1", "OTUB2", "PALB2", "PARP1", "PARP2", "PARP3", "PARP9", "PARPBP", "PAXX", "PBRM1", "PELI1", "PHF10", "PIAS4", "PLK1", "PML", "PNKP", "POGZ", "POLA1", "POLB", "POLL", "POLM", "POLN", "POLQ", "POT1", "PPP4C", "PPP4R2", "PPP4R3B", "PPP5C", "PRDM9", "PRKDC", "PRMT1", "PRPF19", "PSMD14", "RAD21", "RAD21L1", "RAD50", "RAD51", "RAD51AP1", "RAD51B", "RAD51C", "RAD51D", "RAD52", "RAD54B", "RAD54L", "RADX", "RBBP8", "REC8", "RECQL", "RECQL4", "RECQL5", "REV3L", "RFWD3", "RHNO1", "RIF1", "RMI1", "RMI2", "RNF126", "RNF138", "RNF168", "RNF169", "RNF8", "RPA1", "RPA2", "RPA3", "RPA4", "RTEL1", "RUVBL1", "RUVBL2", "SAMHD1", "SENP3", "SETD2", "SETMAR", "SETX", "SFPQ", "SFR1", "SHLD1", "SHLD2", "SHLD3", "SIRT1", "SIRT6", "SKP2", "SLF1", "SLF2", "SLX1A", "SLX1B", "SLX4", "SMARCA2", "SMARCA4", "SMARCAD1", "SMARCAL1", "SMARCB1", "SMARCC1", "SMARCC2", "SMARCD1", "SMARCD2", "SMARCD3", "SMARCE1", "SMC5", "SMC6", "SMCHD1", "SPIDR", "SPIRE1", "SPIRE2", "SPO11", "SWI5", "SWSAP1", "TDP1", "TDP2", "TENT4A", "TERB1", "TERB2", "TERF2IP", "TEX15", "TFIP11", "TIMELESS", "TNKS1BP1", "TONSL", "TOP2B", "TOP3A", "TOPBP1", "TP53", "TP53BP1", "TRIP13", "TRRAP", "TWIST1", "UBE2N", "UBE2V2", "UBQLN4", "UHRF1", "UIMC1", "USP51", "UVRAG", "VCP", "VPS72", "WAS", "WDR48", "WRAP53", "WRN", "XRCC1", "XRCC2", "XRCC3", "XRCC4", "XRCC5", "XRCC6", "YEATS4", "YY1", "ZBTB7A", "ZCWPW1", "ZFYVE26", "ZGRF1", "ZMYND8", "ZNF365", 
         "ZSWIM7")

ros <- c("ABCC1", "ATOX1", "CAT", "CDKN2D", "EGLN2", "ERCC2", "FES", "FTL", "G6PD", "GCLC", "GCLM", "GLRX", "GLRX2", "GPX3", "GPX4", "GSR", "HHEX", "HMOX2", "IPCEF1", "JUNB", "LAMTOR5", "LSP1", "MBP", "MGST1", "MPO", "MSRA", "NDUFA6", "NDUFB4", "NDUFS2", "NQO1", "OXSR1", "PDLIM1", "PFKP", "PRDX1", "PRDX2", "PRDX4", "PRDX6", "PRNP", "PTPA", "SBNO2", "SCAF4", "SELENOS", "SOD1", "SOD2", "SRXN1", "STK25", "TXN", "TXNRD1", "TXNRD2")



my_data <- AddModuleScore(
  object = my_data,
  features = list(ros),
  ctrl = 200,assay = 'RNA_genes', seed = 1984, name = 'ROS')
  

p <- FeaturePlot_scCustom(
  seurat_object = my_data,
  features = 'ROS1',
  order = F,
  split.by = 'species', pt.size = 1, na_cutoff = 0, colors_use =  viridis_magma_dark_high) & NoAxes() & theme(aspect.ratio = 1)



p
ggsave(filename = 'ROS_species_split.pdf', device = 'pdf',plot = p, width = 10, height = 
         3, scale = 1.2)

p <- DimPlot_scCustom(
    seurat_object = my_data,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 1,
    shuffle = TRUE,
    # label = T,
    # label.box = T,
    colors_use = pal_celltypes_all,
    # repel = T,
    # label.size = 5
  ) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_all) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      text =  element_text(family = "Helvetica")
    )& theme(legend.background = element_rect(fill = "transparent"),
             legend.box.background = element_rect(fill = "transparent"),
             panel.background = element_rect(fill = "transparent"),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             plot.background = element_rect(fill = "transparent",
                                            color = NA)) & theme(aspect.ratio = 1)



p
ggsave(filename = 'HE_RPCA_SMC_phenotypes.pdf', device = 'pdf',plot = p, width = 3, height = 
         2, scale = 2)
```


```{r}
human$is_pericytes <- ifelse(human$smc_phenotypes == 'Pericytes', 'Pericytes', 'Non-pericytes')
Idents(human) <- 'is_pericytes'
PC_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'roc')
PC_markers <- PC_markers[PC_markers$cluster == 'Pericytes',]

sum(mouse@assays$RNA@data['Cd200',] > 0) / dim(mouse)[2] 

sum(human@assays$RNA@data['CD200',] > 0) / dim(human)[2] 

sum(pig@assays$RNA@data['CD200',] > 0) / dim(pig)[2] 

p1 <- FeaturePlot_scCustom(mouse, features = 'Cd200',order = F) 
p2 <- FeaturePlot_scCustom(pig, features = 'CD200',order = F) 
p3 <- FeaturePlot_scCustom(human, features = 'CD200',order = F) 

ggdraw() +
  draw_plot(p1, x = 0.0, y = 0, width = .33, height = .48) +
  draw_plot_label(label = 'Mouse SMCs, 86%', size = 12,
                  x = c(-0.01), y = c(0.5), family = 'Helvetica')  +
  
  draw_plot(p2, x = 0.34, y = 0, width = .34, height = .48) +
 
  draw_plot_label(label = 'Pig SMCs, 33%', size = 12,
                  x = c(0.3), y = c(0.5), family = 'Helvetica')  +
  draw_plot(p3, x = 0.65, y = 0, width = .34, height = .48) +
 
  draw_plot_label(label = 'Human SMCs, 28%', size = 12,
                  x = c(0.6), y = c(0.5), family = 'Helvetica')  
```