---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

 This code is for comparison of cluster specificity across different integration methods and gene homology
 

Upload all 4 methods with 3 homology options = 12 datasets

```{r setup, echo=FALSE, include=FALSE}
suppressPackageStartupMessages({
  library(cluster)
  library(Seurat)
  library(SeuratObject)
  library(cowplot)
  library(stats)
  library(data.table)
  library(tidyverse)
  library(ggsci)
  library(scCustomize)
  library(patchwork)
  library(sctransform)
  library(MatrixGenerics)
  library(openxlsx)
  library(RColorBrewer)
  library(psych)
  library(MetBrewer)
  library(pheatmap)
  library(corrplot)
  library(ggpubr)
  library(DESeq2)
  library(viridis)
  library(scales)
  library(ggplotify)
  library(bluster)
  library(stringi)
  library(gridExtra)
  library(scran)

  options(scipen = 999)
  options(repr.matrix.max.rows = 20)
  options(repr.matrix.max.cols = 50)
})

set.seed(8899)
my_pal1 <- MetBrewer::met.brewer('Egypt', n = 4, type = 'discrete')
my_pal2 <- MetBrewer::met.brewer('Signac', n = 14, type = 'discrete')
options(width = 200)


setwd("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal")
knitr::opts_knit$set(root.dir = "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
```

# RPCA

```{r include=FALSE}
# List of .rds files
files <- list.files(path = '~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_coronaries/results/results/seuratRPCA/cross_species/integrated_h5ad/', pattern = '\\.rds', full.names = TRUE)

# Directory to work in
work_dir <- "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal"

# Final species genes
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98), final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
```


```{r include=FALSE}
# Loop through each file
for (file in files) {
  my_data <- readRDS(file)

  my_counts <- GetAssayData(object = my_data, assay = 'RNA', slot = 'data')
  rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["RNA_genes"]] <- CreateAssayObject(data = my_counts)
  DefaultAssay(my_data) <- "RNA_genes"
  
  # Save the modified Seurat object
  saveRDS(my_data, file = paste0(work_dir, '/', basename(file)))
  
  # Clean up memory
  rm(my_counts, my_data, df)
}

print("All files processed.")
```

# CCA

```{r include=FALSE}
# List of .rds files
files <- list.files(path = '~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_coronaries/results/results/seuratCCA/cross_species/integrated_h5ad/', pattern = '\\.rds', full.names = TRUE)

# Directory to work in
work_dir <-  "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal"

# Final species genes
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98), final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)

# Loop through each file
for (file in files) {
  my_data <- readRDS(file)

  my_counts <- GetAssayData(object = my_data, assay = 'RNA', slot = 'data')
  rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["RNA_genes"]] <- CreateAssayObject(data = my_counts)
  DefaultAssay(my_data) <- "RNA_genes"
  
  # Save the modified Seurat object
  saveRDS(my_data, file = paste0(work_dir, '/', basename(file)))
  
  # Clean up memory
  rm(my_counts, my_data, df)
}

print("All files processed.")
```

# scVI and scANVI: create a list

```{r}
work_dir <- '~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_coronaries/results/results/seurat_for_py'
files <- list.files(path = '~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/bengal_coronaries/results/results/seurat_for_py', pattern = '\\.rds', full.names = TRUE)

py_based <- list()
for(i in  seq_along(files) ){
  name_f <-str_split(str_split(files[i], pattern = '/')[[1]][15], pattern = '_')[[1]]
  name_f <- name_f[-grep(pattern = 'matrix', x = name_f)]
  name_f <- unique(name_f)
  x <- match(1, str_detect(string = name_f, pattern = 'coronaries')) +1
  y <- match(1, str_detect(string = name_f, pattern = 'Seurat'))-1
  name_f <- paste0(name_f[  x: y ], collapse = '_')
  print(name_f)
  py_based[[name_f]] <- readRDS(files[i])
}

```

# CCA and RPCA: create a list

```{r}
work_dir <-  "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal"
files <- list.files(path =  "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal", pattern = '\\.rds', full.names = TRUE)

r_based <- list()
for(i in  seq_along(files) ){
  name_f <- str_split(str_split(files[i], pattern = '/')[[1]][13], pattern = '_')[[1]]
  # name_f <- name_f[-grep(pattern = 'matrix', x = name_f)]
  # name_f <- unique(name_f)
  x <- match(1, str_detect(string = name_f, pattern = 'coronaries')) +1
  y <- match(1, str_detect(string = name_f, pattern = 'integrated'))-1
  name_f <- paste0(name_f[  x: y ], collapse = '_')
  print(name_f)
  r_based[[name_f]] <- readRDS(files[i])
}
```

# Choose the right clusterisation for R-based mathods
__Silhouette does not make sense__

```{r include=FALSE}
#from https://github.com/MillerLab-CPHG/Human_athero_scRNA_meta/blob/main/integration_benchmarking/additional_integration_benchmark_metrics.R
# Write a function to calculate clusters silhouette scores from a seurat object
# Arg seurat obj: A seurat obj with reduced dims embedding like pca
# Arg distance_matrix: distance matrix between cells in reduced dims embedding 
# Value: A data.frame with silhouette scores for a given clustering resolution  
calc_sil_scores = function(seurat_obj, distance_matrix) {
  # Calculate distance matrix from seurat obj reduced dims embedding 
  #dist_matrix = dist(x = Seurat::Embeddings(object = seurat_obj[[reduced_dims]])[, 1:30])
  clusters = seurat_obj$seurat_clusters
  sil = silhouette(x = as.numeric(x = as.factor(x = clusters)), 
                   dist = distance_matrix)
  # Add silhouette scores back into seurat obj metadata
  #seurat_obj$silhouette_score = sil[, 3]
  sil_df = data.frame(cluster=sil[, 1],
                      neighbor=sil[, 2],
                      sil_width=sil[, 3])
  return(sil_df)
}


# Directory for saving output
output_dir <- "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal"

# Define a range of resolutions to test
res <- seq(0.1, 1, by = 0.1)

# Loop through each Seurat object
for (name in names(r_based)) {
  my_data <- r_based[[name]]

  # rPCA
  rpca_pca_embeddings <- Embeddings(my_data, "pca")[, 1:30]
  DefaultAssay(my_data) <- 'integrated'

  # Calculate distance matrix
  dist_matrix <- dist(x = rpca_pca_embeddings)
  saveRDS(dist_matrix, paste0(output_dir, "/", name, "_distance_matrix.rds"))
  
  rpca_scores_list <- list()

  for (i in seq_along(res)) {
    # Cluster data
    my_data <- FindClusters(my_data, resolution = res[i])
  
    # Calculate silhouette scores
    rpca_scores_list[[i]] <- calc_sil_scores(my_data, dist_matrix)
    names(rpca_scores_list)[[i]] = paste("res", 
                                       as.character(res[i]), 
                                       sep = "_")
  }

  # Create a single df with all the sil scores for the tested resolutions
  rpca_resolutions_sil_df <- rbindlist(rpca_scores_list, idcol = "method")
  rpca_resolutions_sil_df$method_name <- name
  saveRDS(rpca_resolutions_sil_df, paste0(output_dir, "/", name, "_sil_scores_per_clustering_resolution.rds"))
}
```

# For py based methods

```{r include=FALSE}
# Directory for saving output
output_dir <- "~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal"

# Define a range of resolutions to test
res <- seq(0.1, 1, by = 0.1)

# Loop through each Seurat object
for (name in names(py_based)) {
  my_data <- py_based[[name]]
  
if(str_detect(string = name, pattern = 'scANVI')){
  rpca_pca_embeddings <- Embeddings(my_data, "X_scANVI")[, 1:10]
} else rpca_pca_embeddings <- Embeddings(my_data, "X_scVI")[, 1:40] 
  
  # DefaultAssay(my_data) <- 'integrated'

  # Calculate distance matrix
  dist_matrix <- dist(x = rpca_pca_embeddings)
  saveRDS(dist_matrix, paste0(output_dir, "/", name, "_distance_matrix.rds"))
  
  rpca_scores_list <- list()

  for (i in seq_along(res)) {
    # Cluster data
    if(str_detect(string = name, pattern = 'scANVI')){
      my_data <- RunUMAP(my_data, dims = 1:10, reduction = "X_scANVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:10, reduction = "X_scANVI")
    } else {
      my_data <- RunUMAP(my_data, dims = 1:40, reduction = "X_scVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:40, reduction = "X_scVI")
    }
      
    
    my_data <- FindClusters(my_data, resolution = res[i])
  
    # Calculate silhouette scores
    rpca_scores_list[[i]] <- calc_sil_scores(my_data, dist_matrix)
    names(rpca_scores_list)[[i]] = paste("res", 
                                       as.character(res[i]), 
                                       sep = "_")
  }

  # Create a single df with all the sil scores for the tested resolutions
  rpca_resolutions_sil_df <- rbindlist(rpca_scores_list, idcol = "method")
  rpca_resolutions_sil_df$method_name <- name
  saveRDS(rpca_resolutions_sil_df, paste0(output_dir, "/", name, "_sil_scores_per_clustering_resolution.rds"))
}
```

```{r}
DimPlot(my_data, reduction = "umap",  group.by = 'smc_phenotypes', split.by = 'species')
```

```{r}
res_sil_list <- list()
files <- list.files(path = '/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity', pattern = '_sil_scores_per_clustering_resolution.rds')
for(i in seq_along(files)){
  res_sil_list[[i]] <- readRDS(files[i])
}
res_sil_scores_df <- bind_rows(res_sil_list)
```

```{r}
res_sil_scores_df$method_name = factor(res_sil_scores_df$method_name)
colnames(res_sil_scores_df)[1] <- '.id'
res_sil_scores_df <- res_sil_scores_df[res_sil_scores_df$.id %in% c('res_0.3', 'res_0.4', 'res_0.5', 'res_0.6', 'res_0.7',  'res_0.9',   'res_1')]
res_sil_scores_plot = res_sil_scores_df %>%
  group_by(.id, method_name) %>%
  summarize(mean_sil = mean(sil_width)) %>%
  ggplot(aes(x=.id, y=mean_sil, fill=method_name)) + geom_col(position="dodge", width = 0.7) +
  xlab("Integration Methods / Homology") +
  ylab("Mean Silhouette score") +
  # ggtitle("rPCA mean sil scores per clustering resolution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") + theme_bw() 

ggsave(file="Silh_resolutions.pdf",
       plot = res_sil_scores_plot, width = 16, height = 4.5)
```

```{r include=FALSE}
saveRDS(py_based, file = 'py_based_integration_methods.rds')
saveRDS(r_based, file = 'r_based_integration_methods.rds')
```

```{r}
py_based <- readRDS('py_based_integration_methods.rds')
r_based <- readRDS('r_based_integration_methods.rds')
```

# Do Clutserisation, add assays with gene names istead od Ensembl IDs

```{r include=FALSE}
# Clusterize
r_based_next <- list()
# Loop through each Seurat object
for (name in names(r_based)) {
  my_data <- r_based[[name]]
  DefaultAssay(my_data) <- 'integrated'
  my_data <- FindNeighbors(my_data, reduction = "pca", dims = 1:30)
  my_data <- FindClusters(my_data, resolution = c(0.3, 0.4, 0.5, 0.6))
  
  my_counts <- GetAssayData(object = my_data, assay = 'integrated', slot = 'data')
  rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) <- gsub(pattern = '-1-1', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["integrated_genes"]] <- CreateAssayObject(data = my_counts)
  
  r_based_next[[name]] <- my_data
}
```

```{r include=FALSE}
py_based_next <- list()
for(name in names(py_based)){
  my_data <- py_based[[name]]
  
  my_counts <- GetAssayData(object = my_data, assay = 'RNA', slot = 'data')
  # rownames(my_counts) <- gsub(pattern = '-0', replacement = '', x = rownames(my_counts))
  rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID` %>% table()
  rownames(my_counts)[!rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`]

  my_counts <- my_counts[rownames(my_counts)[rownames(my_counts) %in% final_species_genes$`Pig Ensembl ID`],]

  rownames(final_species_genes) <- final_species_genes$`Pig Ensembl ID`
  df <- data.frame(pig_ens_data = rownames(my_counts), transl_human_gene = final_species_genes[rownames(my_counts), 'Human_gene_name_v98'], transl_hum_ens = final_species_genes[rownames(my_counts), 'Human Ensembl ID'])

  rownames(df) <- df$pig_ens_data
  rownames(my_counts) <- df[rownames(my_counts), 'transl_human_gene']

  my_data[["RNA_genes"]] <- CreateAssayObject(data = my_counts)
  
  # Cluster data
    if(str_detect(string = name, pattern = 'scANVI')){
      my_data <- RunUMAP(my_data, dims = 1:10, reduction = "X_scANVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:10, reduction = "X_scANVI")
    } else {
      my_data <- RunUMAP(my_data, dims = 1:40, reduction = "X_scVI", n.components = 2)
    my_data <- FindNeighbors(my_data, dims = 1:40, reduction = "X_scVI")
    }
  my_data <- FindClusters(my_data, resolution = c(0.3, 0.4, 0.5, 0.6))
  py_based_next[[name]] <- my_data
  
    # Clean up memory
  rm(my_counts, df)
}
```


# Choose right clusterization - so all human clusters are distinguished in the integrated datasets. Clusters should be stable through bootsraping

```{r}
# finding the right clusterization for r_based methods: iterating through resolutions, by bootstrap identify stability of clusters
# then keep only those resolutions that allow to identify more than 7 clusters (6 from human and 1 from others)
my_k <- seq(0.1, 1, 0.1)
df_stable_clusters <- data.frame(k = numeric(),
                                 dataset = character(),
                                 mean_ratio = numeric(),
                                 n_clusters = numeric())
n=1
for(name in names(r_based_next)) {
  for (j in my_k) {
    set.seed(8899)
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = j)$membership
    }
    
    my_data <- r_based_next[[name]]
    DefaultAssay(my_data) <- 'integrated'
    
    pcs <- reducedDim(as.SingleCellExperiment(my_data), "PCA")
    originals <- myClusterFUN(pcs)
    df_stable_clusters[n, 'k'] <- j
    df_stable_clusters[n, 'dataset'] <- name
    
    df_stable_clusters[n, 'mean_ratio'] <-
      mean(diag(
        bootstrapStability(pcs, FUN = myClusterFUN, clusters = originals, iterations = 20)
      ))
    df_stable_clusters[n, 'n_clusters'] <- length(unique(originals))
    n <- n + 1
  }
}
# py_based methods
df_stable_clusters_2 <- data.frame(k = numeric(),
                                 dataset = character(),
                                 mean_ratio = numeric(),
                                 n_clusters = numeric())
n=1
for(name in names(py_based_next)) {
  for (j in my_k) {
    set.seed(8899)
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = j)$membership
    }
    
    my_data <- py_based_next[[name]]
    DefaultAssay(my_data) <- 'RNA'
    if(str_detect(string = name, pattern =  'scANVI')){
      pcs <- reducedDim(as.SingleCellExperiment(my_data), "X_SCANVI")
    } else  pcs <- reducedDim(as.SingleCellExperiment(my_data), "X_SCVI")
    
    originals <- myClusterFUN(pcs)
    df_stable_clusters_2[n, 'k'] <- j
    df_stable_clusters_2[n, 'dataset'] <- name
    
    df_stable_clusters_2[n, 'mean_ratio'] <-
      mean(diag(
        bootstrapStability(pcs, FUN = myClusterFUN, clusters = originals, iterations = 20)
      ))
    df_stable_clusters_2[n, 'n_clusters'] <- length(unique(originals))
    n <- n + 1
  }
}
df_stable_clusters_fin <- rbind(df_stable_clusters, df_stable_clusters_2)
saveRDS(df_stable_clusters_fin, file = 'cluster_stability_bootstrap_res.rds')
```


```{r}
df_stable_clusters_fin <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/cluster_stability_bootstrap_res.rds")
df_stable_clusters_fin <- df_stable_clusters_fin[df_stable_clusters_fin$n_clusters > 6,]

p <- ggplot(df_stable_clusters_fin, aes(x = factor(k), y = mean_ratio, group = dataset, color = dataset)) +
  geom_line() +
  geom_point() +  # Optional: Adds points to the line
  labs(x = "Resolution", y = "Mean Ratio", color = "Dataset") +
  theme_minimal()
print(p)
ggsave(filename = 'Cluster_stability_over_6.pdf', plot = p, width = 6,height = 4, units = 'cm', scale = 3)
```


```{r}
# Filter rows to keep only the maximum mean_ratio for each dataset
df_max <- df_stable_clusters_fin %>%
  group_by(dataset) %>%
  filter(mean_ratio == max(mean_ratio))
df_max <- as.data.frame(df_max)

# Best resolution that gives more than 7 clusters for r_based

for(i in seq_along(df_max$dataset[1:6])){
  name <- df_max[i, 'dataset']
  
   set.seed(8899)
   
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = df_max[i, 'k'])$membership
    }
    
    
    DefaultAssay(r_based_next[[name]]) <- 'integrated'
    
    pcs <- reducedDim(as.SingleCellExperiment(r_based_next[[name]]), "PCA")
    originals <- myClusterFUN(pcs)
    
    r_based_next[[name]]$stab_clusters <- originals
 
}

# Best resolution that gives more than 7 clusters for py_based

for(i in seq_along(df_max$dataset)[7:12]){
  name <- df_max[i, 'dataset']
  
   set.seed(8899)
   
    myClusterFUN <- function(x) {
      g <- bluster::makeSNNGraph(x, type = "jaccard")
      igraph::cluster_louvain(g, resolution = df_max[i, 'k'])$membership
    }
    
    
    DefaultAssay(py_based_next[[name]]) <- 'RNA'
    
     if(str_detect(string = name, pattern =  'scANVI')){
      pcs <- reducedDim(as.SingleCellExperiment(py_based_next[[name]]), "X_SCANVI")
    } else  pcs <- reducedDim(as.SingleCellExperiment(py_based_next[[name]]), "X_SCVI")
    
    originals <- myClusterFUN(pcs)
    
    py_based_next[[name]]$stab_clusters <- originals
 
}

# Aggregate results for r_based and py_based methods and save

all_datasets <- c(r_based_next, py_based_next)
saveRDS(all_datasets, file = 'r_py_all_datasets_processed.rds')
```

# Name clusters (using majority for a cluster)
(stability of clusters)

```{r}
# make new metadata variable 'smc_phenotypes_short'
for (name in names(all_datasets)) {
  Idents(all_datasets[[name]]) <- 'smc_phenotypes'
  
  new_names <-
    c("Term",
      "SMC1",
      "Trans_2",
      "Trans_1",
      "Term_2",
      'SMC2',
      "FB",
      "PC",
      "RERGL")
  names(new_names) <- levels(Idents(all_datasets[[name]]))
  
  all_datasets[[name]] <-
    Seurat::RenameIdents(object = all_datasets[[name]], new_names)
  new_names <- Idents(all_datasets[[name]])
  all_datasets[[name]]$smc_phenotypes_short <-
    new_names
  
  rm(new_names)
}

for (name in names(all_datasets)) {
  my_table <-
    table(all_datasets[[name]]$smc_phenotypes_short,
          all_datasets[[name]]$stab_clusters)
  
  # Function to get the row name with the max value for each column
  get_max_rowname <- function(column) {
    max_index <- which.max(column)
    rownames(my_table)[max_index]
  }
  
  # Apply this function to each column of the table
  max_rownames <- apply(my_table, 2, get_max_rowname)
  # Get the cluster numbers as they appear in the table
  cluster_numbers <- colnames(my_table)
  
  # Create a named vector where names are cluster numbers and values are the phenotypes
  named_max_rownames <- setNames(max_rownames, cluster_numbers)
  
  # Create the 'specific_clusters' metadata variable
  all_datasets[[name]]$specific_clusters <-
    sapply(all_datasets[[name]]$stab_clusters, function(cluster) {
      paste(cluster, named_max_rownames[as.character(cluster)], sep = "_")
    })
  
}
```

# Plots



```{r}
text_sizes <- theme(axis.text.x=element_text(size=9,colour="black"),
                      axis.text.y=element_text(size=9,colour="black"),
                      axis.title.y=element_text(size=9,colour="black", margin = margin(t = 2, l = 1, r = 1, b =2, unit = "pt")),
                      axis.title.x=element_text(size=9,colour="black", margin = margin(t = 1, l = 2, r = 2, b = 1, unit = "pt")),
                      legend.text = element_text(size=9,colour="black"),
                      legend.title = element_text(size=9,colour="black", margin = margin(t = 5, l = 0, r = 0, b = 5, unit = "pt")),
                      legend.key = element_rect(colour="transparent", fill = "transparent"),
                      strip.text.x = element_text(size=6,color = 'black',face="bold", angle=0),
                      strip.text.y = element_text(size=6,color = 'black', face="bold", angle=0, vjust=0.5, hjust=0),
                      axis.ticks= element_line(color = 'black', size=0.2),
                      axis.line = element_line(colour = "gray50", size = 0.2, linetype = "solid"),
                      plot.margin=unit(c(0,0,0,0),"pt"),
                      plot.title=element_text(size=7, face='plain', colour="black"))


common_minimal <- text_sizes + theme(
        plot.background = element_rect(fill = NA,colour = NA),
        strip.background = element_rect(fill = NA,colour = NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) + theme(
    legend.spacing = unit(0.15, 'cm'), 
    legend.key.size = unit(0.2, "cm"))


# commonly used, x axis text 45 degree
common_0x <- common_minimal + theme(axis.text.x = element_text(angle=0))

# pal_species <- pal_npg("nrc", alpha = 0.7)(9)[c(1,4,3)]
#   # MetBrewer::met.brewer(name = 'Thomas')[c(1,5,7)]
# names(pal_species) <- c('sscrofa', 'hsapiens', 'mmusculus')
# saveRDS(pal_species, file = 'pal_species.rds')
# 
# pal_celltypes <- c(pal_nejm(palette = 'default')(8)[-6], pal_jama(palette = 'default')(7) )
#   # sample(c(MetBrewer::met.brewer(name = 'Thomas'), MetBrewer::met.brewer(name = 'Cross')))
# saveRDS(pal_celltypes, file = 'pal_celltypes.rds')
# 
# show_col(pal_jama(palette = 'default')(9))
```

```{r}
pal_celltypes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_celltypes.rds")
pal_species <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_species.rds")
```


__Stacked barplots for celltype composition in species for the stable named clusters__

__First for the purpose of illustration, the contribution of species to the final dataset should be balanced. I will take 1500 cell from each species and exclude GFP and TDT mouse sorted cells__

```{r}

# creating balanced datasets
set.seed(8899)

for (name in names(all_datasets)) {
  all_datasets[[name]]$species_2 <- all_datasets[[name]]$species
  all_datasets[[name]]$species_2[all_datasets[[name]]$species_2 == 'sscrofa'] <- 'pig'
  # all_datasets[[name]]$species_2[all_datasets[[name]]$species_2 == 'mmusculus'] <- 'mouse'
  all_datasets[[name]]$species_2[all_datasets[[name]]$species_2 == 'hsapiens'] <- 'human'
}



my_seu_balanced <- list()
for (name in names(all_datasets)) {
  my_data <- all_datasets[[name]]
  my_data <- UpdateSeuratObject(my_data)
  my_data[["RNA3"]] <- as(object = my_data[["RNA"]], Class = "Assay")
  human <-
    my_data[, sample(colnames(my_data[, my_data$species %in% c('hsapiens')]), size =
                       3600, replace = F)]
  pig <-
    my_data[, sample(colnames(my_data[, my_data$species %in% c('sscrofa')]), size =
                       3600, replace = F)]
  
  balanced <-
    Merge_Seurat_List(list_seurat = list(human, pig), merge.data = T)
  my_seu_balanced[[name]] <- balanced
  rm(human, pig, balanced)
}
```

```{r}
new_names <- c("HE_CCA","HE_RPCA","HHC_CCA",  "HHC_RPCA","O2O_CCA", "O2O_RPCA", "HE_scANVI","HE_scVI",   "HHC_scANVI" ,    "HHC_scVI", "O2O_scANVI"  ,"O2O_scVI" )
names(new_names) <- names(all_datasets)

new_names <- new_names[c(2,4,6,
                         1,3,5,
                         7,9,11,
                         8,10,12)]

names(pal_species) <- c('pig', 'human', 'mouse')
```


```{r}
cluster_prop <- list()
for (name in names(new_names)) {
  balanced <- my_seu_balanced[[name]]
  # res <- colnames(balanced@meta.data)[grep(pattern = '0.6', x = colnames(balanced@meta.data))]
  # contribution
  df <-
    as.data.frame.array(table(balanced@meta.data[, 'specific_clusters'], balanced$species))
  df$sum <- rowSums(df)
  df$pct_sum <- round(df$sum / sum(df$sum) * 100, 1)
  
  #  creating a table with proportions
  # Assuming df is your original dataframe
  df_prop <- df
  
  # Calculate proportions for each column except 'sum'
  df_prop$hsapiens <- df$hsapiens / df$sum
  # df_prop$mmusculus <- df$mmusculus / df$sum
  df_prop$sscrofa <- df$sscrofa / df$sum
  
  # The 'sum' column remains unchanged or can be removed if not needed
  df_prop$sum <- NULL  # Uncomment to remove the 'sum' column
  
  # View the new dataframe with proportions
  df_prop[, 1:2] <- df_prop[, 1:2] * 100
  df_prop$cluster <- rownames(df_prop)
  # df_prop$cluster <- as.numeric(df_prop$cluster)
  df_prop <- df_prop[order(x = df_prop$pct_sum, decreasing = F), ]
  colnames(df_prop)[colnames(df_prop) == 'hsapiens'] <- 'human'
  colnames(df_prop)[colnames(df_prop) == 'sscrofa'] <- 'pig'
  # Transform data to long format
  df_long <- gather(df_prop, species, percent,-cluster,-pct_sum)
  
  
  
  # Calculate cumulative proportions for text placement correctly within each cluster
  df_long <- df_long %>%
    group_by(cluster) %>%
    mutate(cumulative = 100 - (cumsum(percent) - percent / 2))
  
  df_long$cluster <-
    factor(df_long$cluster, levels = unique(df_long$cluster))
  
  # p1 <-
  #   ggplot(df_long, aes(x = cluster, y = percent, fill = species)) +
  #   geom_bar(
  #     # aes(width = rescale(pct_sum, c(0.1, 1))),
  #     stat = "identity", position = "stack") +
  #   geom_text(aes(label = ifelse(percent >= 5, round(percent, 1), ""),
  #                 y = cumulative),
  #             size = 3) +
  #   geom_text(aes(label = pct_sum), y = 108) +
  #   scale_y_continuous(limits = c(0, 110)) +
  #   scale_fill_manual(values = pal_species) +
  #   theme_bw() +
  #   ggtitle(new_names[name]) +
  #   coord_flip()
  #   # rotate_x_text(angle = 45)
  
  p1 <-
    ggplot(df_long, aes(x = cluster, y = percent, fill = species)) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(aes(label = ifelse(percent >= 5, round(percent, 1), ""),
                  y = cumulative),
              size = 3,
              family = "Helvetica") +
    geom_text(aes(label = pct_sum), y = 108, family = "Helvetica") +
    scale_y_continuous(limits = c(0, 110)) +
    scale_fill_manual(values = pal_species) +
    theme_bw() +
    ggtitle(new_names[name]) +
    coord_flip() +
    theme(
      text = element_text(family = "Helvetica"),
      # Set global text family to Arial
      plot.title = element_text(family = "Helvetica"),
      # Ensure title is also Arial
      axis.text = element_text(family = "Helvetica"),
      # Axis texts
      axis.title = element_text(family = "Helvetica")
    )  # Axis titles
  
  
  cluster_prop[[name]] <- p1
}
saveRDS(cluster_prop, file = 'cluster_proportions_list_plots.rds')
n <- length(cluster_prop)
nCol <- floor(sqrt(n))

p1 <- do.call("grid.arrange", c(cluster_prop, ncol = nCol))
p1
# p1 <- plot_grid(plotlist = cluster_prop, nrow = 3, ncol = 4, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Helvetica', label_x = -0.1)
# 
# ggsave(filename = 'proportions_stable_cluster.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 2.5)

```


```{r}
pdf(file = 'proportions_stable_cluster.pdf', width = 16, height = 14, family = 'Helvetica')
plot(p1)
dev.off()
```

__UMAPs with the stable clusters + species__


```{r}
# for purpose of visualisation, subset human part to 2500 cells randomly

myplots <- list()


for(name in names(new_names)) {
  my_vec <- sample(names(all_datasets[[name]]$species[all_datasets[[name]]$species == 'hsapiens']), size = 3600, replace = F) 
  my_species = DimPlot(
  object = all_datasets[[name]][,all_datasets[[name]]$species %in% c('sscrofa')| colnames(all_datasets[[name]]) %in% my_vec],
    reduction = 'umap',
    group.by = c('species_2'),
    pt.size = 0.5,
    shuffle = TRUE
  ) + theme(legend.position = 'none') + text_sizes + theme(plot.title = element_blank()) + 
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
    scale_color_manual(values = pal_species) + theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  my_cell_type = DimPlot_scCustom(seurat_object = all_datasets[[name]][,all_datasets[[name]]$species %in% c('sscrofa')| colnames(all_datasets[[name]]) %in% my_vec],
    reduction = 'umap',
    group.by = 'specific_clusters',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes, repel = T, label.size = 2.3
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  
fig = ggdraw() +
  draw_plot(my_species, x = 0.0, y = .54, width = .99, height = .45) +
  draw_plot(my_cell_type, x = 0.0, y = 0.00005, width = .99, height = .55)  +
  draw_plot_label(label = new_names[name], size = 12,
                  x = c(0), y = c(1.005), family = 'Helvetica')

  
  # fig = plot_grid(my_species,
  #                 my_cell_type,
  #                 ncol = 1, rel_widths = c(0.5, 1),
  #                 rel_heights  = c(0.5, 1))
  
  myplots[[name]] <- fig
  
}

saveRDS(myplots, 'succesful_integration_specific_clusters_species.rds')

# library(gridExtra)
# 
# do.call("grid.arrange", c(myplots[c(1,2)], ncol=2))
```


```{r}

# p <- print(plot_grid(myplots[1][[1]], myplots[2][[1]], myplots[3][[1]], myplots[4][[1]], myplots[5][[1]], myplots[6][[1]], 
#                 myplots[7][[1]], myplots[8][[1]], myplots[9][[1]], myplots[10][[1]], myplots[11][[1]], myplots[12][[1]]
#                 , ncol = 6, axis = 'lb'))

p <- plot_grid(plotlist = myplots, nrow = 2, ncol = 6, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Helvetica', label_x = -0.1)

ggsave(filename = 'UMAP_succesful_integration_specific_clusters_species.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 2)

# save_plot(filename = 'test.pdf',  plot = p, ncol = 6, nrow = 4, )

```

# To find consistent species-specific and shared clusters, we assumed that a cluster is species-specific if: either one species contributes more than 90% of its composition, or one species contributes less than 10% of a cluster composition. In contrast, a clusters was considered as shared if all species contribute to it from 10.1 to 89.9%. 

Here I also introduce a limitation for a size of clusters - it is worth considering only big enough clusters, let's say that their contribution should be more than 2 percent out of all integrated dataset cells.

Therefore, here we work with balanced datasets, where each species contribute 3600 cells. The total number of cells is 7200, and the threshold is 100 cells 

```{r}
# Adjusted function to categorize each cluster
categorize_cluster <- function(human,  pig) {
  contributions <- c(human = human, pig = pig)
  high_contributions <- contributions > 90
  low_contributions <- contributions < 10
  
  if (sum(high_contributions) == 1) {
    # Single species contributes more than 90%
    return(paste(names(contributions)[high_contributions], "specific", sep = "-"))
  } else if (all(contributions >= 10 & contributions <= 90)) {
    # All species contribute between 10% and 90%
    return("shared")
  } else if (sum(low_contributions) == 1) {
    # Exactly one species contributes less than 10%, making the cluster specific for the other two species
    # Filter out the species with less than 10% contribution
    specific_species <- names(contributions)[!low_contributions]
    # Create a label indicating the cluster is specific to the remaining species
    specific_label <- paste(specific_species, collapse = "-")
    return(paste0(specific_label, "-specific"))
  } else {
    return("mixed")
  }
}

my_spec_clusters <- data.frame(
  Method = character(),
  
  Human_specific = character(),
  Pig_specific = character(),
  
  
  Shared = character()
)

list_dfs <- list()
for (name in names(new_names)) {
  balanced <- my_seu_balanced[[name]]
  # res <- colnames(balanced@meta.data)[grep(pattern = '0.6', x = colnames(balanced@meta.data))]
  # contribution
  df <-
    as.data.frame.array(table(balanced@meta.data[, 'specific_clusters'], balanced$species_2))
  df$sum <- rowSums(df)
  df$pct_sum <- round(df$sum / sum(df$sum) * 100, 1)
  
  #  creating a table with proportions
  # Assuming df is your original dataframe
  df_prop <- df
  
  # Calculate proportions for each column except 'sum'
  df_prop$human <- df$human / df$sum
  
  df_prop$pig <- df$pig / df$sum
  
  # The 'sum' column remains unchanged or can be removed if not needed
  # df_prop$sum <- NULL  # Uncomment to remove the 'sum' column
  
  # View the new dataframe with proportions
  df_prop[, 1:2] <- df_prop[, 1:2] * 100
  df_prop$cluster <- rownames(df_prop)
  # df_prop$cluster <- as.numeric(df_prop$cluster)
  df_prop <- df_prop[order(x = df_prop$pct_sum, decreasing = T), ]
  
  # filter out small clusters, that are less than 0.6%
  df_prop$Big <- df_prop$sum > 100
  
   # df_prop <- df_prop[df_prop$sum > 100,]
  
   
  df_prop$category <- apply(df_prop[, c("human",  "pig")], 1, function(x) categorize_cluster(x['human'],  x['pig']))
  
  df_prop$category <- gsub(pattern = 'human.human', replacement = 'human', x = df_prop$category)
  df_prop$category <- gsub(pattern = 'pig.pig', replacement = 'pig', x = df_prop$category)
  list_dfs[[name]] <- df_prop
  
  my_spec_clusters[name, 'Method'] <-  new_names[name]
  
  
  my_spec_clusters[name, 'Human_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'human-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  my_spec_clusters[name, 'Pig_specific'] <- paste(c(unlist(df_prop[df_prop$category == 'pig-specific'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  
  
  
  my_spec_clusters[name, 'Shared'] <- paste(c(unlist(df_prop[df_prop$category == 'shared'& df_prop$Big == TRUE, 'cluster'])), collapse = ', ')
  my_spec_clusters[name, 'Too_small_cluster'] <- paste(c(unlist(df_prop[df_prop$Big == FALSE, 'cluster'])), collapse = ', ')

}

saveRDS(my_spec_clusters, file = 'specific_clusters_df_90_10.rds')
saveRDS(list_dfs, file = 'species_specific_clusters_categoryproportions_df_cor.rds')

print(my_spec_clusters)

# specific_clusters_df_90_10 <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/specific_clusters_df_90_10.rds")
# specific_clusters_df_90_10$tool_name <- rownames(specific_clusters_df_90_10)
# openxlsx::write.xlsx(specific_clusters_df_90_10, file = 'specific_clusters_df_90_10.xlsx')
# 
# specific_clusters_df_10_80 <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/specific_clusters_df_10_80.rds")
# specific_clusters_df_10_80$tool_name <- rownames(specific_clusters_df_10_80)
# openxlsx::write.xlsx(specific_clusters_df_10_80, file = 'elastic_specific_clusters_df_10_80.xlsx')
```

```{r}
set.seed(8899)
# pal_cat <- sample(pal_jco(palette = 'default')(8))
# saveRDS(pal_cat, file = 'pal_cat.rds')
pal_cat <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/pal_cat.rds")
pal_celltypes_all <- c("#0072B5FF",  "#79AF97FF",  "#374E55FF","#80796BFF", "#FFDC91FF","#EE4C97FF", "#7876B1FF","#20854EFF", "#BC3C29FF",  "#E18727FF")
```


```{r}
my_prop <- list()
pal_cat <- c("#868686FF", "#8F7700FF", "#0073C2FF" ,"#EFC000FF")
names(pal_cat) <- c('shared','too_small', 'pig-specific', 'human-specific')
myplots <- list()
for (name in names(new_names)) {
  my_data <- my_seu_balanced[[name]]
  # res <- colnames(balanced@meta.data)[grep(pattern = '0.6', x = colnames(balanced@meta.data))]
  # contribution
  df <-
    as.data.frame.array(table(my_data@meta.data[, 'specific_clusters'], my_data$species_2))
  df$sum <- rowSums(df)
  df$pct_sum <- round(df$sum / sum(df$sum) * 100, 1)
  
  #  creating a table with proportions
  # Assuming df is your original dataframe
  df_prop <- df
  
  # Calculate proportions for each column except 'sum'
  df_prop$human <- df$human / df$sum
 
  df_prop$pig <- df$pig / df$sum
  
  # The 'sum' column remains unchanged or can be removed if not needed
  # df_prop$sum <- NULL  # Uncomment to remove the 'sum' column
  
  # View the new dataframe with proportions
  df_prop[, 1:2] <- df_prop[, 1:2] * 100
  df_prop$cluster <- rownames(df_prop)
  # df_prop$cluster <- as.numeric(df_prop$cluster)
  df_prop <- df_prop[order(x = df_prop$pct_sum, decreasing = T), ]
  
  # filter out small clusters, that are less than 0.6%
  df_prop$Big <- df_prop$sum > 100
   # df_prop <- df_prop[df_prop$sum > 100,]
  
  df_prop$category <- apply(df_prop[, c("human",  "pig")], 1, function(x) categorize_cluster(x['human'],  x['pig']))
  
  df_prop$category <- gsub(pattern = 'human.human', replacement = 'human', x = df_prop$category)
  df_prop$category <- gsub(pattern = 'pig.pig', replacement = 'pig', x = df_prop$category)
  
   df_prop[df_prop$Big == FALSE, 'category'] <- 'too_small'
  my_vec <- df_prop$category
  names(my_vec) <- df_prop$cluster
  my_prop[[name]] <- df_prop
  
  my_data <- all_datasets[[name]]
  Idents(my_data) <- 'specific_clusters'
  my_data$specific_clusters <- factor(my_data$specific_clusters)
  
  my_data$specific_clusters_2 <- my_data$specific_clusters
    
  non_pres <- levels(my_data$specific_clusters)[!levels(my_data$specific_clusters) %in% names(my_vec)]  
  
  my_data$specific_clusters_2[my_data$specific_clusters_2 %in% non_pres] <- NA
  
  Idents(my_data) <- 'specific_clusters_2'
  
  my_data <- RenameIdents(my_data, my_vec)
  my_vec_2 <- Idents(my_data)
  my_data$specificity_category <- my_vec_2
  
  all_datasets[[name]]$specificity_category <- my_vec_2
  
  my_data$smc_phenotypes <- factor(my_data$smc_phenotypes)
  
  my_species = DimPlot_scCustom(
  my_data,
    reduction = 'umap',
    group.by = c('specificity_category'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_cat, repel = T, label.size = 2.7
  ) + 
    scale_color_manual(values = pal_cat) +
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
    scale_color_manual(values = pal_cat) + theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  my_cell_type = DimPlot_scCustom(my_data,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_all, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_all) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
# fig = ggdraw() +
#   draw_plot(my_cell_type, x = 0.0, y = .5001, width = .6, height = .55) +
#   draw_plot(my_species , x = 0.0, y = 0.00005, width = .6, height = .55)  +
#   draw_plot_label(label = new_names[name], size = 12,
#                   x = c(-0.01), y = c(1.005), family = 'Helvetica')
# fig = ggdraw() +
#   draw_plot(my_cell_type, x = 0.0, y = .50, width = .99, height = .55) +
#   draw_plot(my_species, x = 0.0, y = 0.00005, width = .99, height = .55)  +
#   draw_plot_label(label = new_names[name], size = 12,
#                   x = c(0), y = c(1.005), family = 'Helvetica')

fig = ggdraw() +
  draw_plot(my_cell_type, x = 0.0, y = .54, width = .99, height = .45) +
  draw_plot(my_species, x = 0.0, y = 0.00005, width = .99, height = .55)  +
  draw_plot_label(label = new_names[name], size = 12,
                  x = c(0), y = c(1.005), family = 'Helvetica')
  
  # fig = plot_grid(my_species,
  #                 my_cell_type,
  #                 ncol = 1, rel_widths = c(0.5, 1),
  #                 rel_heights  = c(0.5, 1))
  
  myplots[[name]] <- fig
  
}

saveRDS(myplots, file = 'species_specific_clusters_category_9010.rds')

p <- plot_grid(plotlist = myplots, nrow = 2, ncol = 6, align = 'hv',  axis = 'b', greedy = T, label_fontfamily = 'Arial', label_x = -0.1)

ggsave(filename = 'UMAP_specific_clusters_species_categories_9010.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 16, height = 14, scale = 3)


```



# Compute mean and sd for composition of shared and specific clusters
```{r}
my_prop <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/species_specific_clusters_categoryproportions_df_cor.rds")
 

# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% filter(sum > 100) %>% 
    filter(category == "shared") %>%
    summarise(sum_pct = sum(pct_sum, na.rm = TRUE)) %>%
    pull(sum_pct)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df, na.rm = TRUE)
sd_sums   <- sd(sums_each_df, na.rm = TRUE)

cat("Mean of sums for shared category:", mean_sums, "\n") # 78.31667
cat("SD of sums for shared category:  ", sd_sums,   "\n") # 16.74975  

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 78.75 
cat("1st Quartile (Q1): ", Q1, "\n") # 70.975
cat("3rd Quartile (Q3): ", Q3, "\n") # 89.65
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 18.675


# Human-specific RERGL
# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 100) %>% 
    filter(category == "human-specific") %>%
    filter(grepl( 'RERGL', cluster)) %>%
    summarise(human = sum(human, na.rm = TRUE)) %>%
    pull(human)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  97.99235
cat("SD of sums for the category:  ", sd_sums,   "\n") # 1.425579

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 97.65458 
cat("1st Quartile (Q1): ", Q1, "\n") # 96.89737
cat("3rd Quartile (Q3): ", Q3, "\n") # 98.92704
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 2.029664



# Human-specific PC
# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 100) %>% 
    filter(category == "human-specific") %>%
    filter(grepl( 'PC', cluster)) %>%
    summarise(human = sum(human, na.rm = TRUE)) %>%
    pull(human)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0][-2], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0][-2], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  93.44728
cat("SD of sums for the category:  ", sd_sums,   "\n") # 1.860414 

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 94.37739
cat("1st Quartile (Q1): ", Q1, "\n") # 92.17523
cat("3rd Quartile (Q3): ", Q3, "\n") # 95.05678
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 2.881545



# Pig-specific transitional_2
# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 100) %>% 
     filter(category == "pig-specific") %>%
    filter(grepl( 'Trans_2', cluster)) %>%
    summarise(pig = sum(pig, na.rm = TRUE)) %>%
    pull(pig)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  96.84344
cat("SD of sums for the category:  ", sd_sums,   "\n") # 2.831492

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 96.53649
cat("1st Quartile (Q1): ", Q1, "\n") # 95.38626
cat("3rd Quartile (Q3): ", Q3, "\n") # 99.59423 
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 4.207971


# Pig-specific terminal_2
# For each data frame, sum pct_sum where category == "shared"
sums_each_df <- sapply(my_prop, function(df) {
  df %>% 
    filter(sum > 100) %>% 
    filter(category == "pig-specific") %>%
    filter(grepl( 'Term_2', cluster)) %>%
    summarise(pig = sum(pig, na.rm = TRUE)) %>%
    pull(pig)
})

#  Now sums_each_df is a numeric vector (one sum per data frame)
# Calculate mean and sd of those sums across the data frames
mean_sums <- mean(sums_each_df[sums_each_df != 0], na.rm = TRUE)
sd_sums   <- sd(sums_each_df[sums_each_df != 0], na.rm = TRUE)

cat("Mean of sums for the category:", mean_sums, "\n") #  95.72187 
cat("SD of sums for the category:  ", sd_sums,   "\n") # 0.9081213

median_sums <- median(sums_each_df[sums_each_df != 0], na.rm = TRUE)
IQR_value <- IQR(sums_each_df[sums_each_df != 0], na.rm = TRUE, type = 7)
Q1 <- quantile(sums_each_df[sums_each_df != 0], 0.25)  # 25th percentile (1st quartile)
Q3 <- quantile(sums_each_df[sums_each_df != 0], 0.75)  # 75th percentile (3rd quartile)

cat("Median: ", median_sums, "\n") # 95.40707
cat("1st Quartile (Q1): ", Q1, "\n") # 95.0102
cat("3rd Quartile (Q3): ", Q3, "\n") # 96.3186 
cat("Interquartile Range (IQR): ", IQR_value, "\n") # 1.308398 
```




```{r}
new_names <- c("HE_CCA","HE_RPCA","HHC_CCA",  "HHC_RPCA","O2O_CCA", "O2O_RPCA", "HE_scANVI","HE_scVI",   "HHC_scANVI" ,    "HHC_scVI", "O2O_scANVI"  ,"O2O_scVI" )
names(new_names) <- names(all_datasets)

new_names <- new_names[c(2,4,6,
                         1,3,5,
                         7,9,11,
                         8,10,12)]

specific_clusters_df <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/specific_clusters_df_90_10.rds")

my_data <- all_datasets$many_higher_expr_seuratRPCA

my_vec <- c(rep(NA, length(colnames(my_data))))
names(my_vec) <- colnames(my_data)
```

```{r}
### 1) Precompute cluster membership sets

# Let's say we build a list to store membership for each (name, k):
# E.g.: membership_list[[name]][[k]] = character vector of cell names
membership_list <- list()

for (name in names(new_names)) {
  membership_list[[name]] <- list()
  for (k in colnames(specific_clusters_df)[-1]) {
    # gather all cell names that are in specific_clusters_df[name, k]
    target_cluster <- strsplit(specific_clusters_df[name, k], ", ")[[1]]
    # subset the $specific_clusters from all_datasets[[name]]
    # fetch the names of those cells
    # We can store as a character vector
    membership_list[[name]][[k]] <-
      names(all_datasets[[name]]$specific_clusters[
        all_datasets[[name]]$specific_clusters %in% target_cluster
      ])
  }
}

### 2) For each k, combine membership across 'names' if that’s needed
# Because you do: "all_spec <- c(all_spec, ...)" across all "name in new_names"
# we can do that once per k:

combined_membership <- list()  # combined_membership[[k]] = all_spec for that k

for (k in colnames(specific_clusters_df)[-1]) {
  # We'll do a union or concatenation across all name
  all_spec_for_k <- c()
  for (name in names(new_names)) {
    all_spec_for_k <- c(all_spec_for_k, membership_list[[name]][[k]])
  }
  # store it
  combined_membership[[k]] <- all_spec_for_k
}

### 3) Now, loop over i in names(my_vec) and check membership count
# We can define a function is i in all_spec_for_k more than 5 times?

for (i in names(my_vec)) {
  # optional: track if we found a match
  assigned <- FALSE

  for (k in colnames(specific_clusters_df)[-1]) {
    # we have combined_membership[[k]] which is a vector of cell names
    # We do sum(...) or we can do table etc.

    count_i <- sum(combined_membership[[k]] == i)
    if (count_i > 3) {
      my_vec[i] <- k
      assigned <- TRUE
      break  # we found a match, so let's stop checking further k
    }
  }

  if (!assigned) {
    my_vec[i] <- "NS"
  }
}
table(my_vec) # Mouse_human_specific 2 cells, Too_small_cluster 128 cells
my_vec[my_vec %in% c('NS')] <- 'Too_small_cluster'

my_data$consensus <- my_vec[colnames(my_data)]
```

```{r}
p <- DimPlot_scCustom(my_data, group.by = 'consensus', reduction = 'umap') + DimPlot_scCustom(my_data, group.by = 'smc_phenotypes', reduction = 'umap') & theme(aspect.ratio = 1)
ggsave(filename = 'UMAP_HERPCA_consensus_at_least_4int.pdf',  plot = p, device = 'pdf', units = 'cm', dpi = 400, width = 8, height = 3, scale = 3)
```


```{r}
for (name in names(all_datasets)) {
  all_datasets[[name]]$consensus <-  my_vec[colnames(all_datasets[[name]])]
  all_datasets[[name]]$specificity_category_original_2 <-
    all_datasets[[name]]$specificity_category_original
  all_datasets[[name]]$specificity_category <- all_datasets[[name]]$consensus
  all_datasets[[name]]$specificity_category <- factor(all_datasets[[name]]$specificity_category)
}

```

```{r}
saveRDS(all_datasets, file = 'r_py_all_datasets_processed_with_specific_clusters_coronary_2.rds')
```

```{r}
# saveRDS(all_datasets, file = 'r_py_all_datasets_processed_with_specific_clusters.rds')
```

```{r}
all_datasets <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/r_py_all_datasets_processed_with_specific_clusters_coronary_2.rds")
```


# To find out markers for species-specific populations, I will go back to individual datasets. Integrated datasets do not have all genes, but only those that are expressed in all 3 species, while it can be important to have genes expressed solely in some species

```{r}
Idents(my_data) <- 'specificity_category'

tmp <- FindAllMarkers(object = my_data, assay = 'RNA_genes', only.pos = T)
tmp <- tmp[tmp$p_val_adj < 0.05,]

View(tmp[tmp$cluster == 'sscrofa-specific', c(2,3,4,6,7)]) ## IBSP, PTH1R, MIA, COMP, BMP2

table( my_data$smc_phenotypes ,my_data$specificity_category) 
```

# UMAPs of the individual datasets with highlighted species-specific and shared clusters (based on barcodes)


```{r}
 human <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/human_coronaries_seurat.rds")
 pig <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/pig_coronaries_seurat.rds")
```

# Use HE RPCA integareted datasets 

```{r}

human$specificity <- rep(NA, length(human$orig.ident))
pig$specificity <- rep(NA, length(pig$orig.ident))



# Transfer **shared** cluster to all individual species datasets


my_vec <-
      names(all_datasets$many_higher_expr_seuratRPCA$specificity_category[all_datasets$many_higher_expr_seuratRPCA$specificity_category == 'Shared']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Shared', human$specificity)

my_vec <-
      names(all_datasets$many_higher_expr_seuratRPCA$specificity_category[all_datasets$many_higher_expr_seuratRPCA$specificity_category == 'Shared']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Shared', pig$specificity)
        
    
# sscrofa-specific
my_vec <-
      names(all_datasets$many_higher_expr_seuratRPCA$specificity_category[all_datasets$many_higher_expr_seuratRPCA$specificity_category == 'Pig_specific']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Pig-specific', pig$specificity)
    
my_vec <-
      names(all_datasets$many_higher_expr_seuratRPCA$specificity_category[all_datasets$many_higher_expr_seuratRPCA$specificity_category == 'Pig_specific']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Pig-specific', human$specificity)
    
    
# hsapiens-specific
my_vec <-
      names(all_datasets$many_higher_expr_seuratRPCA$specificity_category[all_datasets$many_higher_expr_seuratRPCA$specificity_category == 'Human_specific']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Human-specific', human$specificity)
my_vec <-
      names(all_datasets$many_higher_expr_seuratRPCA$specificity_category[all_datasets$many_higher_expr_seuratRPCA$specificity_category == 'Human_specific']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Human-specific', pig$specificity)
    
### Small cluster    
my_vec <-
      names(all_datasets$many_higher_expr_scVI$specificity_category[all_datasets$many_higher_expr_scVI$specificity_category == 'Too_small_cluster']) %>% na.omit()
    my_vec <- gsub('-hsapiens', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    human$specificity <-
      ifelse(colnames(human) %in% names(my_vec), 'Small_contribution', human$specificity)

    
my_vec <-
      names(all_datasets$many_higher_expr_scVI$specificity_category[all_datasets$many_higher_expr_scVI$specificity_category == 'Too_small_cluster']) %>% na.omit()
    my_vec <- gsub('-sscrofa', replacement = '', x = my_vec)
    names(my_vec) <- my_vec
    pig$specificity <-
      ifelse(colnames(pig) %in% names(my_vec), 'Small_contribution', pig$specificity)

```


```{r}
# plots smc_phenotypes + specificIty on individual datasets
# pal_specificity <- met.brewer(name = 'Cross', n = 9, direction = -1)
# pal_specificity <- pal_specificity[c(3,9,5, 2, 7)]

pal_specificity <- c('#C1C8D9', '#8491B4df',  '#4f576c',pal_npg("nrc")(10)[c(1,2,3,10,5,7,8,9)])
names(pal_specificity) <- c('Contractile', 'Transitional', 'Terminal',  'Pericytes_Human-specific', 'LBH_SMC_Human-specific', 'Fibroblasts',   'Terminal_IGFBP2_Pig_specific', 'IFN_SMC', 'Transitional_2_Pig-specific')

# names(pal_specificity) <- c('Contractile', 'Transitional', 'Terminal',  'Pericytes_Human_pig_specific', 'DLX5_SMC_Human_specific','Terminal_Mouse_specific', 'Proliferating', 'Contractile_Human_pig_specific',   'Too_small_cluster')

levels(Idents(pig))
pig$smc_phenotypes <- factor(pig$smc_phenotypes, levels = levels(pig$smc_phenotypes)[c(3,4,2,1,5,8,7,6)])
Idents(pig) <- 'smc_phenotypes'
pal_celltypes_pig <-  c("#0072B5FF","#E18727FF", "coral1",  "#BC3C29FF","#7876B1FF" ,"#FFDC91FF", "#EE4C97FF" ,"#20854EFF")

pal_celltypes_human <- c( "#FFDC91FF","#20854EFF","#E18727FF","#0072B5FF",   "#374E55FF" ,"#BC3C29FF" )




human$smc_phenotypes_specificity <- paste(human$smc_phenotypes, human$specificity, sep = '_')
pig$smc_phenotypes_specificity <- paste(pig$smc_phenotypes, pig$specificity, sep = '_')

my_vec <- c('LBH_SMC_Human-specific', 'Pericytes_Human-specific',  'Terminal_IGFBP2_Pig-specific', 'Transitional_2_Pig-specific')

pig$for_heatmap <- ifelse(pig$smc_phenotypes_specificity %in% my_vec, pig$smc_phenotypes_specificity, 'Shared')
human$for_heatmap <- ifelse(human$smc_phenotypes_specificity %in% my_vec, human$smc_phenotypes_specificity, 'Shared')

# Pig
my_data <- readRDS("~/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/shared_Cor_HE_RPCA_cell_categorization.rds")


xxx <- my_data[, my_data$species == 'sscrofa']$shared_clusters 
names(xxx) <-  substr(names(xxx), start = 1, stop = 16)
xxx <- xxx[substr(colnames(pig), start = 1, stop = 16)]
names(xxx) <- colnames(pig)
pig$shared_clusters <- xxx
pig$shared_clusters <- as.character(pig$shared_clusters)
pig$shared_clusters[is.na(pig$shared_clusters)] <- 'specific'
table(pig$shared_clusters, pig$smc_phenotypes)

pig$shared_clusters <- ifelse(pig$shared_clusters == 'specific', pig$specificity, no = pig$shared_clusters)
pig$shared_clusters[pig$shared_clusters == 'Pig-specific' & pig$smc_phenotypes == 'Transitional_2'] <- 'Transitional_2_Pig-specific' 
pig$shared_clusters[pig$shared_clusters == 'Pig-specific' & pig$smc_phenotypes == 'Terminal_IGFBP2'] <- 'Terminal_IGFBP2_Pig_specific' 
pig$shared_clusters[pig$shared_clusters == 'Human-specific' & pig$smc_phenotypes == 'Pericytes'] <- 'Pericytes_Human-specific'

pig$shared_clusters[pig$smc_phenotypes == 'IFN_SMC'] <- 'IFN_SMC' 

table(pig$shared_clusters)
table(pig$shared_clusters, pig$smc_phenotypes)

pig <- pig[,pig$shared_clusters %in% c('Contractile', 'Fibroblasts', 'IFN_SMC', 'Terminal', 'Terminal_IGFBP2_Pig_specific', 'Pericytes_Human-specific', 'Transitional', 'Transitional_2_Pig-specific')]

# Human
xxx <- my_data[, my_data$species == 'hsapiens']$shared_clusters 
names(xxx) <-  substr(names(xxx), start = 12, stop = 27)
xxx <- xxx[substr(colnames(human), start = 12, stop = 27)]
names(xxx) <- colnames(human)
human$shared_clusters <- xxx
human$shared_clusters <- as.character(human$shared_clusters)
human$shared_clusters[is.na(human$shared_clusters)] <- 'specific'
human$shared_clusters <- ifelse(human$shared_clusters == 'specific', human$specificity, no = human$shared_clusters)
table(human$shared_clusters, human$smc_phenotypes)

human$shared_clusters[human$shared_clusters == 'Human-specific'& human$smc_phenotypes == 'Pericytes'] <- 'Pericytes_Human-specific' 
human$shared_clusters[human$shared_clusters == 'Human-specific'& human$smc_phenotypes == 'LBH_SMC'] <- 'LBH_SMC_Human-specific' 

table(human$shared_clusters)
table(human$shared_clusters, human$smc_phenotypes)

human <- human[,human$shared_clusters %in% c('Contractile', 'Fibroblasts', 'IFN_SMC', 'Terminal', 'LBH_SMC_Human-specific', 'Pericytes_Human-specific', 'Transitional')]

##### Pig
spec_pig = DimPlot_scCustom(
  pig,
    reduction = 'umap',
    group.by = c('shared_clusters'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_specificity, repel = T, label.size = 2.7
  ) + 
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
      scale_color_manual(values = pal_specificity) + 
  theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  smc_pig = DimPlot_scCustom(pig,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_pig, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_pig) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
###### Human
spec_human = DimPlot_scCustom(
  human,
    reduction = 'umap',
    group.by = c('shared_clusters'),
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T,  colors_use = pal_specificity, repel = T, label.size = 2.7
  ) + 
    theme(legend.position = 'none') +
   text_sizes + theme(plot.title = element_blank()) + 
    # guides(color = guide_legend(nrow = 3)) +
  #    plot_annotation(title =  new_names[name]) +  
  # theme(text = element_text(size = 10)) +
      scale_color_manual(values = pal_specificity) + 
  theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  
  
  smc_human = DimPlot_scCustom(human,
    reduction = 'umap',
    group.by = 'smc_phenotypes',
    pt.size = 0.5,
    shuffle = TRUE, label = T, label.box = T, colors_use = pal_celltypes_human, repel = T, label.size = 2.7
  ) + NoLegend()+
    # guides(color = guide_legend(nrow = 6)) +
    # theme(legend.position = 'none', legend.text = element_text(size = 9)) +  
    theme(plot.title = element_blank()) + 
    scale_color_manual(values = pal_celltypes_human) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
#### Combine

fig = ggdraw() +
  draw_plot(spec_pig, x = 0, y = 0, width = .4, height = .5) +
  draw_plot(smc_pig , x = 0, y = 0.5, width = .4 , height =  .5)  +
  draw_plot_label(label = 'Pig coronary SMCs', size = 12,
                  x = c(0), y = c(1.005), family = 'Helvetica')  +
  draw_plot(spec_human, x = 0.5, y = 0, width = .4, height = .5) +
  draw_plot(smc_human , x = 0.5, y = .5 , width = .4 , height = .5 )  +
  draw_plot_label(label = 'Human coronary SMCs', size = 12,
                  x = c(0.5), y = c(1.005), family = 'Helvetica')  

fig 

saveRDS(fig, file = 'UMAPS_individual_ds_species_specific_clusters_category.rds')


ggsave(filename = 'UMAPS_individual_ds_species_specific_clusters_category.pdf',  plot = fig, device = 'pdf', units = 'cm', dpi = 400, width = 6, height = 4, scale = 3)  


```

```{r}
# Marker identification for cross-species SMC phenotypes mapping
Idents(pig) <- 'shared_clusters'
tmp_p <- FindAllMarkers(pig, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_p, file = 'pig_coronary_unified_markers.xlsx')

Idents(human) <- 'shared_clusters'
tmp_h <- FindAllMarkers(human, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_h, file = 'human_coronary_unified_markers.xlsx')

Idents(pig) <- 'smc_phenotypes'
tmp_p <- FindAllMarkers(pig, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_p, file = 'pig_coronary_ind_markers.xlsx')

Idents(human) <- 'smc_phenotypes'
tmp_h <- FindAllMarkers(human, assay = 'RNA', only.pos = T)
openxlsx::write.xlsx(x = tmp_h, file = 'human_coronary_ind_markers.xlsx')

```


# Marker identification for species-specific SMC phenotypes

```{r}
```

# Human LBH_SMCs. Majority of the cluster cells are placed specifically as unique human cells, so I will find markers for LBH_SMCs cluster

```{r}
table(human$smc_phenotypes, human$specificity)

Idents(human) <- 'smc_phenotypes'

# LBH_SMC_markers <- multiMarkerStats(
#     wilcox=findMarkers(as.SingleCellExperiment(human), test="wilcox", direction="up",group = human$smc_phenotypes, add.summary = TRUE),
#     binom=findMarkers(as.SingleCellExperiment(human), test="binom", direction="up",group = human$smc_phenotypes, add.summary = TRUE), sorted = TRUE
# )
# 
# LBH_SMC_markers <- LBH_SMC_markers@listData[["LBH_SMC"]] %>% as.matrix() %>% as.data.frame()
# LBH_SMC_markers <- LBH_SMC_markers[LBH_SMC_markers$wilcox.summary.AUC > 0.7,]
# LBH_SMC_markers <- LBH_SMC_markers[order(LBH_SMC_markers$binom.summary.logFC, decreasing = T),] 
# LBH_SMC_markers$gene <- rownames(LBH_SMC_markers)


LBH_SMC_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'MAST', only.pos = T)
LBH_SMC_markers <- LBH_SMC_markers[LBH_SMC_markers$cluster == 'LBH_SMC',]
```

# Pig-specific Terminal_IGFBP2 cells

```{r}
table(pig$smc_phenotypes)
Idents(pig) <- 'smc_phenotypes'
# T_IGFBP2 <- multiMarkerStats(
#     wilcox=findMarkers(as.SingleCellExperiment(pig), test="wilcox", direction="up",group = pig$smc_phenotypes, add.summary = TRUE),
#     binom=findMarkers(as.SingleCellExperiment(pig), test="binom", direction="up",group = pig$smc_phenotypes, add.summary = TRUE), sorted = TRUE
# )
# 
# T_IGFBP2 <- T_IGFBP2@listData[["Terminal_IGFBP2"]] %>% as.matrix() %>% as.data.frame()
# T_IGFBP2 <- T_IGFBP2[T_IGFBP2$wilcox.summary.AUC > 0.7,]
# T_IGFBP2 <- T_IGFBP2[order(T_IGFBP2$binom.summary.logFC, decreasing = T),] 
# colnames(T_IGFBP2) <- paste('Pig', colnames(T_IGFBP2), sep = '_')
# T_IGFBP2$gene <- rownames(T_IGFBP2)


T_IGFBP2 <- FindAllMarkers(object = pig, assay = 'RNA', test.use = 'MAST', only.pos = T)
T_IGFBP2 <- T_IGFBP2[T_IGFBP2$cluster == 'Terminal_IGFBP2',]


```

# Pig-specific Transitional_2 cells

```{r}
table(pig$smc_phenotypes)
Idents(pig) <- 'smc_phenotypes'
# T_CD74 <- multiMarkerStats(
#     wilcox=findMarkers(as.SingleCellExperiment(pig), test="wilcox", direction="up",group = pig$smc_phenotypes, add.summary = TRUE),
#     binom=findMarkers(as.SingleCellExperiment(pig), test="binom", direction="up",group = pig$smc_phenotypes, add.summary = TRUE), sorted = TRUE
# )
# 
# T_CD74 <- T_CD74@listData[["Transitional_2"]] %>% as.matrix() %>% as.data.frame()
# T_CD74 <- T_CD74[T_CD74$wilcox.summary.AUC > 0.7,]
# T_CD74 <- T_CD74[order(T_CD74$binom.summary.logFC, decreasing = T),] 
# colnames(T_CD74) <- paste('Pig', colnames(T_CD74), sep = '_')
# T_CD74$gene <- rownames(T_CD74)


T_CD74 <- FindAllMarkers(object = pig, assay = 'RNA', test.use = 'MAST', only.pos = T)
T_CD74 <- T_CD74[T_CD74$cluster == 'Transitional_2',]


```


```{r}
species_markers <- list('Pig Terminal_IGFBP2' = T_IGFBP2, 
                        'Human LBH_pericytes' = LBH_SMC_markers,
                        'Pig Transitional_2' = T_CD74)

saveRDS(species_markers, file = 'df_markers_species_specific_clusters_scran_coronaries.rds')
openxlsx::write.xlsx(species_markers, file = "Markers_species_specific_clusters_scran_coronaries.xlsx",
                     asTable = TRUE, overwrite = TRUE)

# species_markers <- list('Pig Terminal_IGFBP2' = T_IGFBP2, 
#                         'Human LBH_pericytes' = LBH_SMC_markers)
# 
# saveRDS(species_markers, file = 'df_markers_species_specific_clusters_MAST.rds')
# openxlsx::write.xlsx(species_markers, file = "Markers_species_specific_clusters_MAST.xlsx",
#                      asTable = TRUE, overwrite = TRUE)
```

# Visualization of markers for species-specific clusters. 

Make one-to-one orthologs, but keep features that are not expressed in other species

```{r}
species_markers <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/after_bengal/cluster_specificity/df_markers_species_specific_clusters.rds") # or _scran with binom mode

DefaultAssay(pig) <- 'RNA'
DefaultAssay(human) <- 'RNA'

# mouse$specificity <- ifelse(is.na(mouse$specificity) == T, 'Other', mouse$specificity)
# pig$specificity <- ifelse(is.na(pig$specificity) == T, 'Other', pig$specificity)
# human$specificity <- ifelse(is.na(human$specificity) == T, 'Other', human$specificity)
# 
# mouse$specificity <- ifelse(mouse$specificity == 'Shared', 'shared', mouse$specificity)
# pig$specificity <- ifelse(pig$specificity == 'Shared', 'shared', pig$specificity)
# human$specificity <- ifelse(human$specificity == 'Shared', 'shared', human$specificity)



human$smc_phenotypes_specificity <- paste(human$smc_phenotypes, human$specificity, sep = '_')
pig$smc_phenotypes_specificity <- paste(pig$smc_phenotypes, pig$specificity, sep = '_')

my_vec <- c('LBH_SMC_Human-specific', 'Pericytes_Human-specific',  'Terminal_IGFBP2_Pig-specific', 'Transitional_2_Pig-specific')


pig$for_heatmap <- ifelse(pig$smc_phenotypes_specificity %in% my_vec, pig$smc_phenotypes_specificity, 'Pig_rest')
human$for_heatmap <- ifelse(human$smc_phenotypes_specificity %in% my_vec, human$smc_phenotypes_specificity, 'Human_rest')
```

```{r}
final_species_genes <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/biomart_109_orth_new.rds")
final_species_genes$Human_gene_name_v98 <- ifelse(is.na(final_species_genes$Human_gene_name_v98) == TRUE, final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)
final_species_genes$Human_gene_name_v98 <- ifelse(final_species_genes$Human_gene_name_v98 == '', final_species_genes$`Human Ensembl ID`, final_species_genes$Human_gene_name_v98)

final_species_genes[nrow(final_species_genes)+1,'Human_gene_name_v98' ] <- 'CST3'
final_species_genes[nrow(final_species_genes),'Human Ensembl ID' ] <- 'CST3'
final_species_genes[nrow(final_species_genes),'mouse_origin' ] <- 'Cst3'
final_species_genes[nrow(final_species_genes),'pig_origin' ] <- 'CST3'

final_species_genes[nrow(final_species_genes)+1,'Human_gene_name_v98' ] <- 'BGLAP'
final_species_genes[nrow(final_species_genes),'Human Ensembl ID' ]  <- 'BGLAP'
final_species_genes[nrow(final_species_genes),'mouse_origin' ] <- 'Bglap'
final_species_genes[nrow(final_species_genes),'pig_origin' ] <- 'BGLAP'


final_species_genes[nrow(final_species_genes)+1,'Human_gene_name_v98' ] <- 'COL9A3'
final_species_genes[nrow(final_species_genes),'Human Ensembl ID' ] <-  'COL9A3'
final_species_genes[nrow(final_species_genes),'mouse_origin' ] <- 'Col9a3'
final_species_genes[nrow(final_species_genes),'pig_origin' ] <- 'COL9A3'
```


```{r}
DefaultAssay(pig) <- 'RNA' 
pig_2 <- pig

# Counts
DefaultAssay(pig_2) <- 'RNA' 
counts_all <- GetAssayData(object = pig_2, slot = "counts", assay = 'RNA')
my_df <- data.frame(old=rownames(pig_2), ##create a dataframe of correspondence
                    old_2=rownames(pig_2),
                    original = rownames(pig_2))
rownames(my_df) <- rownames(pig_2)
colnames(my_df)[2] <- 'pig_origin' # Pig gene name
final_species_genes$scRnaseq_pig_2 <- final_species_genes$pig_origin %in% my_df$pig_origin
my_df_2 <- merge(my_df,final_species_genes,by=c('pig_origin'))  ## merging data about mouse dataset genes and ortholog table, Pig gene name
my_df_2$human_fin <- ifelse(is.na(my_df_2$Human_gene_name_v98) == TRUE, my_df_2$`Human Ensembl ID`, my_df_2$Human_gene_name_v98)
rownames(my_df_2) <- my_df_2$original
my_df_2['CST3', 'human_fin'] <- 'CST3'
filtered_counts_h <- counts_all[my_df_2$original, ] ## keep only genes that have orthologous
rownames(filtered_counts_h) <- my_df_2$human_fin
pig_2@assays$RNA@counts <- filtered_counts_h

# Data
DefaultAssay(pig_2) <- 'RNA' 
counts_all <- GetAssayData(object = pig, slot = "data", assay = 'RNA')
my_df <- data.frame(old=rownames(pig_2), ##create a dataframe of correspondence
                    old_2=rownames(pig_2),
                    original = rownames(pig_2))
rownames(my_df) <- rownames(pig_2)
colnames(my_df)[2] <- 'pig_origin' # Pig gene name
final_species_genes$scRnaseq_pig_2 <- final_species_genes$pig_origin %in% my_df$pig_origin
my_df_2 <- merge(my_df,final_species_genes,by=c('pig_origin'))  ## merging data about mouse dataset genes and ortholog table, Pig gene name
my_df_2$human_fin <- ifelse(is.na(my_df_2$Human_gene_name_v98) == TRUE, my_df_2$`Human Ensembl ID`, my_df_2$Human_gene_name_v98)
rownames(my_df_2) <- my_df_2$original
my_df_2['CST3', 'human_fin'] <- 'CST3'
filtered_counts_h <- counts_all[my_df_2$original, ] ## keep only genes that have orthologous
rownames(filtered_counts_h) <- my_df_2$human_fin
pig_2@assays$RNA@data <- filtered_counts_h
```



```{r}
human_2 <-  human[, sample(colnames(human), size = 2500, replace = F)]
pig_2 <-  pig_2[, sample(colnames(pig_2), size = 2500, replace = F)]
merged <- Merge_Seurat_List(list_seurat = list( pig_2, human_2), merge.data = T)
saveRDS(merged, file = 'merged_pig_human_smcs_new.rds')

merged <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/merged_pig_human_smcs_new.rds")


my_data <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/shared_Cor_HE_RPCA_cell_categorization.rds") # reclustered shared clusters


# We need to devide 'rest categories' to shared phenotypes
merged$for_heatmap <- factor(merged$for_heatmap)
xxx <- my_data$shared_clusters 

names(xxx) <-  substr(names(xxx), start = 1, stop = 16)
xxx <- xxx[substr(colnames(merged), start = 2, stop = 17)]
names(xxx) <- colnames(merged)
merged$shared_clusters <- xxx
merged$shared_clusters <- as.character(merged$shared_clusters)
merged$shared_clusters[is.na(merged$shared_clusters)] <- 'specific'
table(merged$shared_clusters, merged$for_heatmap)

merged$for_heatmap <-
  factor(merged$for_heatmap, levels = 
    c(
      "LBH_SMC_Human-specific",
      "Pericytes_Human-specific",
      'Terminal_IGFBP2_Pig-specific',
       "Transitional_2_Pig-specific",
      'Human_rest', 'Pig_rest'
    
  ))
table(merged$shared_clusters, merged$for_heatmap)
merged$xxx <- paste(merged$shared_clusters, merged$for_heatmap, sep = '_')

merged$xxx <- ifelse(merged$xxx %in% c('Contractile_Human_rest', 'Contractile_Pig_rest'), 'Contractile_shared', merged$xxx) 
merged$xxx <- ifelse(merged$xxx %in% c('Transitional_Human_rest', 'Transitional_Pig_rest'), 'Transitional_shared', merged$xxx )
merged$xxx <- ifelse(merged$xxx %in% c('Terminal_Human_rest', 'Terminal_Pig_rest'), 'Terminal_shared', merged$xxx )
merged$xxx <- ifelse(merged$xxx %in% c('Fibroblasts_Human_rest', 'Fibroblasts_Pig_rest'), 'Fibroblasts_shared', merged$xxx )
merged$xxx <- ifelse(merged$xxx %in% c('IFN_SMC_Human_rest', 'IFN_SMC_Pig_rest'), 'IFN_SMC_shared', merged$xxx )
table(merged$xxx)

# names(merged$xxx[merged$xxx == 'Terminal_shared'])
merged$for_heatmap_2 <- merged$for_heatmap
merged$for_heatmap_2 <- as.character(merged$for_heatmap_2)
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Terminal_shared'])] <- 'Terminal_shared'
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Contractile_shared'])] <- 'Contractile_shared'
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Transitional_shared'])] <- 'Transitional_shared'

merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'Fibroblasts_shared'])] <- 'Fibroblasts_shared'
merged$for_heatmap_2[names(merged$xxx[merged$xxx == 'IFN_SMC_shared'])] <- 'IFN_SMC_shared'

table(merged$for_heatmap_2, merged$smc_phenotypes) # 
table(merged$for_heatmap_2)



merged <- merged[,merged$for_heatmap_2 %in%  c(
      "LBH_SMC_Human-specific",
      "Pericytes_Human-specific",
      'Terminal_IGFBP2_Pig-specific',
       "Transitional_2_Pig-specific",
      
      'Contractile_shared',
      'Transitional_shared',
      'Terminal_shared' ,
      'IFN_SMC_shared',
      'Fibroblasts_shared')]

#### Make heatmap
Idents(merged) <- 'for_heatmap_2'


merged$for_heatmap_2 <-
  factor(merged$for_heatmap_2, levels = 
    c(
      "LBH_SMC_Human-specific",
      "Pericytes_Human-specific",
      'Terminal_IGFBP2_Pig-specific',
       "Transitional_2_Pig-specific",
      
      'Contractile_shared',
      'Transitional_shared',
      'Terminal_shared' ,
      'IFN_SMC_shared',
      'Fibroblasts_shared' ))





#### Make heatmap
Idents(merged) <- 'for_heatmap_2'
# tmp_mouse <- FindAllMarkers(object = merged, assay = 'RNA', test.use = 'roc', min.pct = 0.2, only.pos = T, max.cells.per.ident = 3000, min.diff.pct = 0.1)

tmp_mouse <- FindAllMarkers(object = merged, assay = 'RNA',  min.pct = 0.3, only.pos = T, max.cells.per.ident = 3000, min.diff.pct = 0.1)

top_markers <- c()
for (i in unique(tmp_mouse$cluster)){
  top_markers <- c(top_markers, head((subset(tmp_mouse, subset = tmp_mouse$cluster == i)[order(subset(tmp_mouse, subset = tmp_mouse$cluster == i)$avg_log2FC, decreasing = T),])$gene, 6))
}
# top_markers[40] <- 'LUM'
# top_markers[12] <- 'GJA4'
# top_markers[20] <- 'CD74'
# top_markers[7] <- 'RGS5'
top_markers[12] <- 'GJA4'
top_markers[7] <- 'RGS5'
top_markers[26] <- 'ACTA2'
top_markers[37] <- 'LUM'
 top_markers[36] <- 'GAS6'
top_markers[39] <- 'CLU'

# species_markers <- readRDS("/faststorage/project/THOR/diana/integration_pigs_with_WT/integrated_pig_mice_hum/closer/more_closer/SMC/coronaries/after_bengal/df_markers_species_specific_clusters_scran_coronaries.rds")
# 
# top_markers <- c()
# for (i in names(species_markers)[c(2,1)]){
#   my_col <- colnames(species_markers[[i]])[ grep('Top', colnames(species_markers[[i]]))][1]
#   top_markers <- c(top_markers, species_markers[[i]][order(species_markers[[i]][,my_col], decreasing = F), 'gene'][1:10])
# }

p1 <- DotPlot(
  merged,
  features = toupper(unique(top_markers)),
  assay = 'RNA',
  scale = T,
  group.by = 'for_heatmap_2'
) + 
  # ggtitle('Top markers for SMC phenotypes pig aorta plaque') + 
  # coord_flip() +
  # RotatedAxis() + 
  scale_colour_gradientn(name = 'log2 (count + 1)',colours = rev(brewer.pal(n = 11, name = "RdBu")))  + 
  theme_bw() +
  theme(axis.line  = element_blank()) +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ylab('') +theme(legend.position = 'bottom') +
  theme(axis.ticks = element_blank()) + coord_fixed(ratio = 1)+ text_sizes
print(p1)


ggsave(filename = 'dotplot_merged_species-specific_clusters_original_9010_coronaries_new.3.pdf', plot = p1, device = 'pdf', width = 12, height = 3, units = 'cm', scale = 3)
```

```{r}
table(merged$for_heatmap_2)
Idents(merged) <- 'for_heatmap_2'
tmp <- FindAllMarkers(merged, assay = 'RNA', only.pos = T)
saveRDS(tmp, 'markers_for_enrichment_coonaries.rds')
```

```{r}
# terminal
feats <- species_markers$`Mouse terminal`[1:12, 'gene']

FeaturePlot_scCustom(mouse, features = feats) + plot_annotation(title = 'Mouse SMCs') 

FeaturePlot_scCustom(pig, features = str_to_upper(feats) )+ plot_annotation(title = 'Pig SMCs') 

FeaturePlot_scCustom(human, features = str_to_upper(feats)) + plot_annotation(title = 'Human SMCs') 

# DLX5 human specific
feats <- c(species_markers$`Human DLX5`[1:12, 'gene'], 'MECOM',  'ID3')

# Pericytes
feats <- c(species_markers$`Human-pig pericytes`[species_markers$`Human-pig pericytes`$Human_pct.2 < 0.5 & species_markers$`Human-pig pericytes`$Pig_pct.2 < 0.5, 'gene'], 'Gucy1b3', 'Filip1')

# cSMCs
feats <- c(species_markers$`Human-pig cSMCs`[, 'gene'], )

FeaturePlot_scCustom(mouse, features = str_to_title(feats)  ) + plot_annotation(title = 'Mouse SMCs') 

FeaturePlot_scCustom(pig, features = str_to_upper(feats) )+ plot_annotation(title = 'Pig SMCs') 

FeaturePlot_scCustom(human, features = str_to_upper(feats)) + plot_annotation(title = 'Human SMCs') 
```


```{r}
human$is_pericytes <- ifelse(human$smc_phenotypes == 'Pericytes', 'Pericytes', 'Non-pericytes')
Idents(human) <- 'is_pericytes'
PC_markers <- FindAllMarkers(object = human, assay = 'RNA', test.use = 'roc')
PC_markers <- PC_markers[PC_markers$cluster == 'Pericytes',]

sum(mouse@assays$RNA@data['Cd200',] > 0) / dim(mouse)[2] 

sum(human@assays$RNA@data['CD200',] > 0) / dim(human)[2] 

sum(pig@assays$RNA@data['CD200',] > 0) / dim(pig)[2] 

p1 <- FeaturePlot_scCustom(mouse, features = 'Cd200',order = F) 
p2 <- FeaturePlot_scCustom(pig, features = 'CD200',order = F) 
p3 <- FeaturePlot_scCustom(human, features = 'CD200',order = F) 

ggdraw() +
  draw_plot(p1, x = 0.0, y = 0, width = .33, height = .48) +
  draw_plot_label(label = 'Mouse SMCs, 86%', size = 12,
                  x = c(-0.01), y = c(0.5), family = 'Helvetica')  +
  
  draw_plot(p2, x = 0.34, y = 0, width = .34, height = .48) +
 
  draw_plot_label(label = 'Pig SMCs, 33%', size = 12,
                  x = c(0.3), y = c(0.5), family = 'Helvetica')  +
  draw_plot(p3, x = 0.65, y = 0, width = .34, height = .48) +
 
  draw_plot_label(label = 'Human SMCs, 28%', size = 12,
                  x = c(0.6), y = c(0.5), family = 'Helvetica')  
```